### 2020-08-12 individual-cohorts/meta_analysis/meta-analysis_diagnosis.Rmd
Add separate meta analysis folder

---
title: "Meta-analysis on diagnosis in MSSM-Penn-Pitt and HBCC ACC/DLPFC"
author: "KM and GH"
date: "`r date()`"
output: html_document
---

```{r knit2synapse, eval=FALSE}
# rmarkdown::render("meta-analysis_diagnosis.Rmd")

library(synapser)
library(knit2synapse)
synLogin()
knit2synapse::createAndKnitToFolderEntity(file = "individual-cohorts/meta-analysis.Rmd",
                                          parentId ="syn22276904",
                                          folderName = 'Meta-analysis on diagnosis')
```

```{r setup, include=FALSE}
library(synapser)

synLogin()

library(ggplot2)
library(dplyr)
library(ComplexHeatmap)
library(readxl)
library(gtools)
library(corrplot)
library(pvclust)
library(metafor)
library(cowplot)
library(githubr)
library(org.Hs.eg.db)
library(ggrepel)
library(metafor)
library(patchwork)

source("../../common_functions/common_functions.R")

knitr::opts_chunk$set(
  echo = FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=FALSE, eval=TRUE}
thisFileName <- 'meta-analysis.Rmd'

# Github link
thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/',thisFileName))
```

```{r input_data}
mpp_acc <- readr::read_tsv(synGet("syn22079117")$path) %>% 
  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
mpp_dlpfc <- readr::read_tsv(synGet("syn22079472")$path) %>% 
  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
hbcc_dsACC <- readr::read_tsv(synGet("syn22086829")$path) %>% 
  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) 
hbcc_dlpfc <- readr::read_tsv(synGet("syn22084205")$path) %>% 
  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)
```

# Control vs SCZ
```{r meta_analysis.SZ}
make_list <- function(mpp, hbcc, string) {
  list <- list(MPP = mpp %>%
                 unique() %>% 
                 dplyr::filter(Comparison == string) %>% 
                 as.data.frame,
               HBCC = hbcc %>% 
                 unique() %>% 
                 dplyr::filter(Comparison == string) %>% 
                 as.data.frame)
}

dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Control_vs_SCZ")
accList <- make_list(mpp_acc, hbcc_dsACC, string = "Control_vs_SCZ")

rownames(dlpfcList[['MPP']]) = dlpfcList[['MPP']]$gene_id
rownames(dlpfcList[['HBCC']]) = dlpfcList[['HBCC']]$gene_id

res_dlpfc = run_meta_analysis( dlpfcList, "FE")
res_dlpfc$Dataset = "Meta-analysis"

rownames(accList[['MPP']]) = accList[['MPP']]$gene_id
rownames(accList[['HBCC']]) = accList[['HBCC']]$gene_id

res_acc = run_meta_analysis( accList, "FE")
res_acc$Dataset = "Meta-analysis"
```

```{r plot.meta_analysisSCZ}
kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))

plotVolcano( res_dlpfc, 20) + ggtitle("Meta analysis - DLPFC")

plotVolcano( res_acc, 20) + ggtitle("Meta analysis - ACC")
```

# Control vs Other/SCZ
```{r meta_analysis.SCZOther}
make_list <- function(mpp, hbcc, string) {
  list <- list(MPP = mpp %>%
                 unique() %>% 
                 dplyr::filter(Comparison == string) %>% 
                 as.data.frame,
               HBCC = hbcc %>% 
                 unique() %>% 
                 dplyr::filter(Comparison == string) %>% 
                 as.data.frame)
}

dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Control_vs_OtherSCZ")
accList <- make_list(mpp_acc, hbcc_dsACC, string = "Control_vs_OtherSCZ")

rownames(dlpfcList[['MPP']]) = dlpfcList[['MPP']]$gene_id
rownames(dlpfcList[['HBCC']]) = dlpfcList[['HBCC']]$gene_id

res_dlpfc = run_meta_analysis( dlpfcList, "FE")
res_dlpfc$Dataset = "Meta-analysis"

rownames(accList[['MPP']]) = accList[['MPP']]$gene_id
rownames(accList[['HBCC']]) = accList[['HBCC']]$gene_id

res_acc = run_meta_analysis( accList, "FE")
res_acc$Dataset = "Meta-analysis"
```

```{r plot.meta_analysisSCZOther}
kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))

plotVolcano( res_dlpfc, 20) + ggtitle("Meta analysis - DLPFC")

plotVolcano( res_acc, 20) + ggtitle("Meta analysis - ACC")
```

# Control vs Other
```{r meta_analysis.CTOther}
make_list <- function(mpp, hbcc, string) {
  list <- list(MPP = mpp %>%
                 unique() %>% 
                 dplyr::filter(Comparison == string) %>% 
                 as.data.frame,
               HBCC = hbcc %>% 
                 unique() %>% 
                 dplyr::filter(Comparison == string) %>% 
                 as.data.frame)
}

dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Control_vs_Other")
accList <- make_list(mpp_acc, hbcc_dsACC, string = "Control_vs_Other")

rownames(dlpfcList[['MPP']]) = dlpfcList[['MPP']]$gene_id
rownames(dlpfcList[['HBCC']]) = dlpfcList[['HBCC']]$gene_id

res_dlpfc = run_meta_analysis( dlpfcList, "FE")
res_dlpfc$Dataset = "Meta-analysis"

rownames(accList[['MPP']]) = accList[['MPP']]$gene_id
rownames(accList[['HBCC']]) = accList[['HBCC']]$gene_id

res_acc = run_meta_analysis( accList, "FE")
res_acc$Dataset = "Meta-analysis"
```

```{r plot.meta_analysisCTOther}
kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))

plotVolcano( res_dlpfc, 20) + ggtitle("Meta analysis - DLPFC")

plotVolcano( res_acc, 20) + ggtitle("Meta analysis - ACC")
```

# Other vs SCZ
```{r meta_analysis.OtherSCZ}
make_list <- function(mpp, hbcc, string) {
  list <- list(MPP = mpp %>%
                 unique() %>% 
                 dplyr::filter(Comparison == string) %>% 
                 as.data.frame,
               HBCC = hbcc %>% 
                 unique() %>% 
                 dplyr::filter(Comparison == string) %>% 
                 as.data.frame)
}

dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Other_vs_SCZ")
accList <- make_list(mpp_acc, hbcc_dsACC, string = "Other_vs_SCZ")

rownames(dlpfcList[['MPP']]) = dlpfcList[['MPP']]$gene_id
rownames(dlpfcList[['HBCC']]) = dlpfcList[['HBCC']]$gene_id

res_dlpfc = run_meta_analysis( dlpfcList, "FE")
res_dlpfc$Dataset = "Meta-analysis"

rownames(accList[['MPP']]) = accList[['MPP']]$gene_id
rownames(accList[['HBCC']]) = accList[['HBCC']]$gene_id

res_acc = run_meta_analysis( accList, "FE")
res_acc$Dataset = "Meta-analysis"
```

```{r plot.meta_analysisOtherSCZ}
kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))
kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))

plotVolcano( res_dlpfc, 20) + ggtitle("Meta analysis - DLPFC")

plotVolcano( res_acc, 20) + ggtitle("Meta analysis - ACC")
```

## RCSD1

Forest plot of example top gene.

```{r specific-gene}
# forrest plot of meta-analysis
ensID = 'ENSG00000198771.11'
kable(res_dlpfc[ensID,])
lapply(dlpfcList, function(x) x[ensID,])
ensID = rownames(res_dlpfc)[which.min(res_dlpfc$adj.P.Val)][1]
plot_forrest(dlpfcList, ensID)
```

### Source