### 2020-08-24 deconvolution/apply_dtangle_FANS4.Rmd
Initial commit

---
title: "Deconvolution to FANS4"
author: "Gabriel Hoffman"
date: "`r date()`"
output: html_document
---

```{r knit2synapse, eval=FALSE}
ml git pandoc
rmarkdown::render("deconvolution/apply_dtangle_FANS4.Rmd")

library(synapser)
library(knit2synapse)

synLogin()

knit2synapse::createAndKnitToFolderEntity(file = "deconvolution/apply_dtangle_FANS4.Rmd",
                                 parentId ="syn22321035",
                                 folderName = "Deconvolution results",
                                 knitmd=FALSE)
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
library(data.table)
library(dtangle)
library(edgeR)
library(variancePartition)
library(BiocParallel)

library(synapser)
library(future)
library(knitr)
library(githubr) # get the package from devtools::install_github('brian-bot/githubr')

source("../common_functions/common_functions.R")

synLogin()

# See dream documentation to customize parallel processing       http://bioconductor.org/packages/release/bioc/vignettes/variancePartition/inst/doc/dream.html#parallel-processing
param = SnowParam((availableCores()-2)/2, "SOCK", progressbar=TRUE)

register(param)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
parentId = 'syn22321035';
activityName = 'Deconvolution';
activityDescription = 'Deconvolution';

thisFileName <- 'apply_dtangle_FANS4.Rmd'

# Github link
thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('deconvolution/',thisFileName))
```

```{r read.data}

# Read reference panel
######################
geneExpr_FANS4_lcpm = read.csv(synGet('syn22321183')$path, header=TRUE)
rownames(geneExpr_FANS4_lcpm) = geneExpr_FANS4_lcpm[,1]
geneExpr_FANS4_lcpm = geneExpr_FANS4_lcpm[,-1]
metadata_FANS4 = read.csv(synGet("syn22321189")$path, header=TRUE, row.names=1)

# read Raw counts
#################
file_counts_synid = c('syn21888969', 'syn21862455', 'syn21887889', 'syn21825508')

# read one file at a time
geneCounts_list = lapply(file_counts_synid, function(id){
	read.table(synGet(id)$path, header=TRUE, check.names=FALSE)
	})

# lapply(geneCounts_list, function(x) x[1:3, 1:3])

# merge files based on gene name
geneCounts = Reduce(function(x,y) merge(x = x, y = y, by = "ensembl_gene_id"), geneCounts_list)
rownames(geneCounts) = trim_ensembl_ids(geneCounts[,1])
geneCounts = geneCounts[,-1]

# TMM normalization
geneExpr = DGEList( as.matrix(geneCounts))
geneExpr = calcNormFactors(geneExpr)
geneExpr_logCPM = cpm(geneExpr, log=TRUE)

# only keep genes measured in FANS4
keep = rownames(geneExpr_logCPM) %in% rownames(geneExpr_FANS4_lcpm)
geneExpr_logCPM = geneExpr_logCPM[keep,]

```

```{r process.counts}

idx = match(rownames(geneExpr_logCPM), rownames(geneExpr_FANS4_lcpm))

# a = cbind( rownames(geneExpr_logCPM), rownames(geneExpr_FANS4_lcpm[idx,]) )
# table(a[,1] == a[,2])
# which(a[,1] != a[,2])

cellFracEstimates = dtangle( Y = t(geneExpr_logCPM), reference = t(geneExpr_FANS4_lcpm[idx,]))

cellFracEstimates2 = dtangle2( Y = t(geneExpr_logCPM), reference = t(geneExpr_FANS4_lcpm[idx,]))

cor(cellFracEstimates$estimates, cellFracEstimates2$estimates, method="spearman")
```

### Data preprocessing
```{r preprocess.data}
# Get genotype ids from synapse
GENOTYPE_ID = 'syn16816490'
ALL_USED_IDs = c(ALL_USED_IDs, GENOTYPE_ID)
GENOTYPE = downloadFile(GENOTYPE_ID) %>% 
  dplyr::select(Individual_ID, `SNP_report:Genotyping_Sample_ID`, `SNP_report:Exclude?`) %>% 
  dplyr::inner_join(ANCESTRY) %>% 
  dplyr::filter(is.na(`SNP_report:Exclude?`))

# Merge metadata
METADATA = metadata %>%
  dplyr::left_join(GENOTYPE) %>% 
  dplyr::rename(Region = `rnaSeq_dissection:Brain_Region`,
                PMI = `PMI_(in_hours)`,
                RIN = `rnaSeq_isolation:RIN`, 
                ReportExclude = `rnaSeq_report:Exclude?`,
                GenotypeExclude = `SNP_report:Exclude?`,
                SampleID = Sample_RNA_ID,
                LibraryBatch = `rnaSeq_report:Library_Batch`,
                FlowcellBatch = `rnaSeq_report:Flowcell_Batch`,
                RibozeroBatch = `rnaSeq_report:Ribozero_Batch`,
                MappedReads = `rnaSeq_report:Mapped_Reads`,
                IntragenicRate = `rnaSeq_report:Intragenic_Rate`, IntronicRate = `rnaSeq_report:Intronic_Rate`, 
                IntergenicRate = `rnaSeq_report:Intergenic_Rate`, GenesDetected = `rnaSeq_report:Genes_Detected`,
                ExpProfEfficiency = `rnaSeq_report:Expression_Profiling_Efficiency`, rRNARate = `rnaSeq_report:rRNA_Rate`,
                TotalReads = `rnaSeq_report:Total_Reads`, AlignmentRate = `rnaSeq_report:Percent_Aligned`, TranscriptsDetected = `rnaSeq_report:Transcripts_Detected`) %>%
  dplyr::select(SampleID, Individual_ID, Institution, Cohort, Reported_Gender,Sex, Ethnicity, ageOfDeath, PMI, Dx, 
                RIN, ReportExclude, GenotypeExclude, EV.1, EV.2, EV.3, EV.4, EV.5,
                LibraryBatch, FlowcellBatch, RibozeroBatch, MappedReads, TotalReads, GenesDetected, AlignmentRate,
                IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate, TranscriptsDetected) %>%
  dplyr::filter(Cohort %in% "NIMH-HBCC") %>% 
  dplyr::filter(SampleID %in% colnames(COUNT), !is.na(SampleID)) %>% 
  dplyr::mutate(Dx = forcats::fct_recode(Dx, Other = "BP", Other = "undetermined", Control = "Control", SCZ = "SCZ"))

ind = METADATA$SampleID[which(METADATA$ReportExclude == 1 | METADATA$GenotypeExclude)]
writeLines(paste('Following',length(ind),'samples are marked exclude'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$Ethnicity) | is.na(METADATA$Institution) | is.na(METADATA$Dx)]
writeLines(paste('Following', length(ind), 'counts are missing any metadata'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$PMI)]
writeLines(paste('Following', length(ind), 'counts are missing PMI'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$Reported_Gender)]
writeLines(paste('Following', length(ind), 'counts are missing gender'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$ageOfDeath)]
writeLines(paste('Following', length(ind), 'counts are missing age of death'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$EV.1)]
writeLines(paste('Following', length(ind), 'counts are missing ancestry information'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))
```

```{r preprocess.data1, results='asis'}
# Match covariates to expression data
indToRetain = intersect(METADATA$SampleID, colnames(COUNT))

rownames(COUNT) <- COUNT$Geneid
COUNT = COUNT[,indToRetain]

rownames(METADATA) = METADATA$SampleID
METADATA = METADATA[indToRetain,]

METADATA %>% 
  group_by(Dx) %>% 
  dplyr::summarise(count = n()) %>% 
  kable()

FactorCovariates <- c('Individual_ID', "Reported_Gender", "LibraryBatch", "Dx", "FlowcellBatch", "RibozeroBatch")
ContCovariates <- c("ageOfDeath", "PMI", "RIN", "MappedReads", "IntragenicRate", "IntronicRate", "IntergenicRate",
                    "GenesDetected", "ExpProfEfficiency", "rRNARate", "TotalReads", "AlignmentRate",
                    "EV.1", "EV.2", "EV.3", "EV.4", "EV.5")

# Find inter relation between factor covariates
COVARIATES = METADATA[,c(FactorCovariates,ContCovariates),drop=F]
COVARIATES[,FactorCovariates] <- data.frame(lapply(COVARIATES[,FactorCovariates],function(x){
  x <- sapply(x,function(y){str_replace_all(as.character(y),'[^[:alnum:]]','_')})}))
rownames(COVARIATES) <- METADATA$SampleID

# Convert factor covariates to factors
COVARIATES[,FactorCovariates] = lapply(COVARIATES[,FactorCovariates], factor)
COVARIATES[,ContCovariates] = lapply(COVARIATES[,ContCovariates], function(x){
  x = as.numeric(as.character(gsub('[\\,\\%]','',x)))
})

```{r store, eval=FALSE}
# Code
CODE <- Folder(name = "Meta-analysis on sex", parentId = "syn22315403") # Make sure these match the folder name and parent Id in the first chunk.
CODE <- synStore(CODE)

# Store meta-analysis
file = "cellFracEstimates.RDS"
saveRDS(cellFracEstimates, file=file)
nEXP_OBJ = File(file, 
                name = 'Result of dtangle', 
                parentId = CODE$properties$id)
nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
                    executed = thisFile)
res_acc %>%
  tibble::rownames_to_column(var = "gene_id") %>% 
  write.table(file = 'CMC_acc_Sex_meta_analysis.tsv', sep = '\t', row.names=F, quote=F)
nEXP_OBJ = File('CMC_acc_Sex_meta_analysis.tsv', 
                name = 'Sex Normalised, meta-analysis (ACC)', 
                parentId = CODE$properties$id)
nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
                    executed = thisFile)
```

### Source
[markdown](`r thisFile`)