### 2020-08-26 deconvolution/apply_dtangle_FANS4.Rmd
add upload

---
title: "Deconvolution to FANS4"
author: "Gabriel Hoffman"
date: "`r date()`"
output: html_document
---

```{r knit2synapse, eval=FALSE}
ml git pandoc
rmarkdown::render("deconvolution/apply_dtangle_FANS4.Rmd")

library(synapser)
library(knit2synapse)

synLogin()

knit2synapse::createAndKnitToFolderEntity(file = "deconvolution/apply_dtangle_FANS4.Rmd",
                                 parentId ="syn22321035",
                                 folderName = "Deconvolution results",
                                 knitmd=TRUE)
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
suppressPackageStartupMessages({
library(data.table)
library(dtangle)
library(edgeR)
library(variancePartition)
library(BiocParallel)
library(RNOmni)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)
library(synapser)
library(future)
library(knitr)
library(githubr) # get the package from devtools::install_github('brian-bot/githubr')
})
source("../common_functions/common_functions.R")

synLogin()

# See dream documentation to customize parallel processing       http://bioconductor.org/packages/release/bioc/vignettes/variancePartition/inst/doc/dream.html#parallel-processing
param = SnowParam((availableCores()-2)/2, "SOCK", progressbar=TRUE)

register(param)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
parentId = 'syn22321035';
activityName = 'Deconvolution';
activityDescription = 'Deconvolution';

thisFileName <- 'apply_dtangle_FANS4.Rmd'

# Github link
thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('deconvolution/',thisFileName))
```

```{r read.data}
# Read reference panel
######################
geneExpr_FANS4_lcpm = read.csv(synGet('syn22321183')$path, header=TRUE)
rownames(geneExpr_FANS4_lcpm) = geneExpr_FANS4_lcpm[,1]
geneExpr_FANS4_lcpm = geneExpr_FANS4_lcpm[,-1]
metadata_FANS4 = read.csv(synGet("syn22321189")$path, header=TRUE, row.names=1)

# read Raw counts
#################
file_counts_synid = c('syn21888969', 'syn21862455', 'syn21887889', 'syn21825508')
ALL_USED_IDs = c('syn22321183', file_counts_synid)

# read one file at a time
geneCounts_list = lapply(file_counts_synid, function(id){
	read.table(synGet(id)$path, header=TRUE, check.names=FALSE)
	})

# lapply(geneCounts_list, function(x) x[1:3, 1:3])

# merge files based on gene name
geneCounts = Reduce(function(x,y) merge(x = x, y = y, by = "ensembl_gene_id"), geneCounts_list)
rownames(geneCounts) = trim_ensembl_ids(geneCounts[,1])
geneCounts = geneCounts[,-1]

# TMM normalization
geneExpr = DGEList( as.matrix(geneCounts))
geneExpr = calcNormFactors(geneExpr)
geneExpr_logCPM = cpm(geneExpr, log=TRUE)

# only keep genes measured in FANS4
keep = rownames(geneExpr_logCPM) %in% rownames(geneExpr_FANS4_lcpm)
geneExpr_logCPM = geneExpr_logCPM[keep,]

```

```{r process.counts}

idx = match(rownames(geneExpr_logCPM), rownames(geneExpr_FANS4_lcpm))

# a = cbind( rownames(geneExpr_logCPM), rownames(geneExpr_FANS4_lcpm[idx,]) )
# table(a[,1] == a[,2])
# which(a[,1] != a[,2])

cellFracEstimates = dtangle( Y = t(geneExpr_logCPM), reference = t(geneExpr_FANS4_lcpm[idx,]))

# cellFracEstimates2 = dtangle2( Y = t(geneExpr_logCPM), reference = t(geneExpr_FANS4_lcpm[idx,]))

# cor(cellFracEstimates$estimates, cellFracEstimates2$estimates, method="spearman")
```

### Data preprocessing
```{r preprocess.data}
synId_covariates = c('syn21862454', 'syn21887888', 'syn21825507') # 'syn21888968'

COV_list = lapply(synId_covariates, function(id){
	df = read.table( synGet(id)$path, header=TRUE, stringsAsFactors=FALSE)
	df$SampleID = as.character(df$SampleID)
	if( any(colnames(df) == "Dx.Tissue") ){
		df$Dx = gsub("^(\\S+)_(\\S+)$", "\\1", df$Dx.Tissue)
		df$Tissue = gsub("^(\\S+)_(\\S+)$", "\\2", df$Dx.Tissue)
	}	
	if( !any("Institution" == colnames(df)) ){
		df$Institution = "HBCC"

	}
	if( id == 'syn21862454' ){
		df$Tissue = "DLPFC"
	}
	if( id == 'syn21887888' ){
		df$Tissue = "sgACC"
	}
	if( !any('Tissue' == colnames(df))){
		stop(id)
	}
	df
	})

COVARIATES = gtools::smartbind( list = COV_list)

df = with(COVARIATES, data.frame(SampleID, Study = paste(Institution, Tissue)))
df$Study = gsub("Penn|Pitt", "MSSM", df$Study)

ids = unique(df$Study)
sampleByStudy = lapply( ids, function(stdy){
	df[df$Study == stdy, "SampleID"]
	})
names(sampleByStudy) = ids

# only keep individuals older than 17
COVARIATES = COVARIATES[COVARIATES$ageOfDeath > 17,]

```

```{r deconv.compare.covariates}

deconvEstimates = data.frame(cellFracEstimates$estimates)

f = log
# f = rankNorm

df_deconv_test = lapply( names(sampleByStudy) , function(stdy){

	smplids = sampleByStudy[[stdy]]

	deconvEst = deconvEstimates[smplids,]

	df = merge(deconvEst, COVARIATES[COVARIATES$SampleID %in% smplids,], by.x="row.names", by.y="SampleID")

	form = cbind(GABA 	= f(GABA), 
				GLU 	= f(GLU), 
				Olig 	= f(Olig),
				MgAs 	= f(MgAs)) ~ ageOfDeath + Reported_Gender + PMI + RIN + RIN2 + Dx 
	fit = lm(form, df)

	res = lapply( summary(fit), function(x) data.frame(p = -log10(coef(x)[-1,4]) * sign(coef(x)[-1,1]) ))

	df_return = do.call(cbind, res)
	colnames(df_return) = gsub("Response ", "", names(res))
	data.frame(Study = stdy, coef = rownames(df_return), df_return)

	})

df_deconv_test = do.call(rbind, df_deconv_test)

df_deconv_test_melt = reshape2::melt(df_deconv_test, id.vars=c("Study", "coef"))
n_tests = nrow(df_deconv_test_melt)
df_deconv_test_melt$isSignif = with(df_deconv_test_melt, abs(value) > -log10(0.05 / n_tests))

lim = max(abs(df_deconv_test_melt$value))

ratio = with(df_deconv_test_melt, length(unique(coef)) / length(unique(Study)))

ggplot(df_deconv_test_melt, aes(Study, coef, fill=value, label=ifelse(isSignif, '*', ''))) + geom_tile() + theme_bw() + theme(aspect.ratio = ratio, panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.text.x=element_text(angle=45, hjust=1)) + scale_fill_gradient2(high = "red", mid = "white", low="blue", limits=c(-lim, lim), name="-log10 p") + facet_wrap( ~ variable) + geom_text()
```

```{r combine.samples}

df = merge(deconvEstimates, COVARIATES, by.x="row.names", by.y="SampleID")

form = ~ (1|Individual_ID) + ageOfDeath + (1|Reported_Gender) + PMI + RIN + (1|Dx) + (1|Tissue)
vp = fitExtractVarPartModel( t(log(df[,colnames(deconvEstimates)])), form, df, BPPARAM=SerialParam())

plotPercentBars( vp )
```

```{r cca, fig.width=5, fig.height=5}
form = ~ Individual_ID + ageOfDeath + Reported_Gender + PMI + RIN + Dx + Tissue
C = canCorPairs(form, df)
plotCorrMatrix(C)
```

```{r store, eval=TRUE}
# Code
CODE <- Folder(name = "Deconvolution results", parentId = "syn22321035") # Make sure these match the folder name and parent Id in the first chunk.
CODE <- synStore(CODE)

# Store meta-analysis
file = "cellFracEstimates_dtangle_FANS4.tsv"
# saveRDS(data.frame(cellFracEstimates$estimates), file=file)
write.table(cellFracEstimates$estimates, quote=FALSE, sep='\t', file=file)

nEXP_OBJ = File(file, 
                name = 'Estimates from dtangle', 
                parentId = CODE$properties$id)
nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
                    executed = thisFile)
```

### Source
[markdown](`r thisFile`)