### 2020-09-01 individual-cohorts/meta_analysis/meta-analysis_sex.Rmd
Freeze analysis

---
title: "Meta-analysis on sex in MSSM-Penn-Pitt and HBCC ACC/DLPFC"
author: "KM and GH"
date: "`r date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

<!---

# individual-cohorts/meta_analysis/

# rmarkdown::render("individual-cohorts/meta_analysis/meta-analysis_sex.Rmd")

--->

```{r knit2synapse, eval=FALSE, echo=FALSE}

createAndKnitToFolderEntity <- function(file, parentId, folderName,
                                        wikiName=NULL, overwrite=FALSE, knitmd=TRUE, ...) {

  entity <- synapser::Folder(parentId=parentId, name=folderName)
  entity <- synapser::synStore(entity, ...)

  knitfile2synapse(file=file, owner=entity, wikiName=wikiName, 
                   overwrite=overwrite, knitmd=knitmd, parentWikiId=NULL)
}

library(synapser)
library(knit2synapse)
synLogin()
# knit2synapse::
createAndKnitToFolderEntity(file = "individual-cohorts/meta_analysis/meta-analysis_sex.Rmd",
                                          parentId ="syn22315403",
                                          folderName = 'Meta-analysis on sex',
                                          overwrite=TRUE, 
                                          knitmd=TRUE)
```

```{r setup, include=FALSE}
library(synapser)

synLogin()

library(ggplot2)
library(dplyr)
library(ComplexHeatmap)
library(readxl)
library(gtools)
library(corrplot)
library(pvclust)
library(metafor)
library(cowplot)
library(kableExtra)
library(githubr)
library(org.Hs.eg.db)
library(ggrepel)
library(metafor)
library(knitr)
library(patchwork)

source("../../common_functions/common_functions.R")

knitr::opts_chunk$set(
  echo = FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE, eval=TRUE}
thisFileName <- 'meta-analysis_sex.Rmd'
# thisFile = thisFileName
# Github link
thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/meta_analysis/',thisFileName))
```

```{r input_data}
mpp_acc <- readr::read_tsv(synGet("syn22278494")$path) %>% 
  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% getGeneSymbol( 'gene_id' )
mpp_dlpfc <- readr::read_tsv(synGet("syn22278597")$path) %>% 
  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% getGeneSymbol( 'gene_id' )
hbcc_dsACC <- readr::read_tsv(synGet("syn22280959")$path) %>% 
  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% getGeneSymbol( 'gene_id' )
hbcc_dlpfc <- readr::read_tsv(synGet("syn22280588")$path) %>% 
  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% getGeneSymbol( 'gene_id' )
Ids <- c("syn22278494", "syn22278597", "syn22280959", "syn22280588")
```

```{r plotma}

f = function(x){
  pmin(1, pmax(-1, x))
}

ggplot(mpp_acc, aes(AveExpr, f(logFC))) + geom_point() + theme_bw() + geom_smooth(method="loess")  + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + ggtitle("mpp_acc")

ggplot(mpp_dlpfc, aes(AveExpr, f(logFC))) + geom_point() + theme_bw() + geom_smooth(method="loess")  + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + ggtitle("mpp_dlpfc")

ggplot(hbcc_dsACC, aes(AveExpr, f(logFC))) + geom_point() + theme_bw() + geom_smooth(method="loess")  + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + ggtitle("hbcc_dsACC")

ggplot(hbcc_dlpfc, aes(AveExpr, f(logFC))) + geom_point() + theme_bw() + geom_smooth(method="loess")  + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + ggtitle("hbcc_dlpfc")
```

# Female vs Male
```{r meta_analysis.Sex}
make_list <- function(mpp, hbcc, string) {
  list <- list(MPP = mpp %>%
                 unique() %>% 
                 dplyr::filter(Comparison == string) %>% 
                 as.data.frame,
               HBCC = hbcc %>% 
                 unique() %>% 
                 dplyr::filter(Comparison == string) %>% 
                 as.data.frame)
}

dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Female_vs_Male")
accList <- make_list(mpp_acc, hbcc_dsACC, string = "Female_vs_Male")

rownames(dlpfcList[['MPP']]) = dlpfcList[['MPP']]$gene_id
rownames(dlpfcList[['HBCC']]) = dlpfcList[['HBCC']]$gene_id

res_dlpfc = run_meta_analysis( dlpfcList, "FE") %>% getGeneSymbol
res_dlpfc$Dataset = "Meta-analysis"

rownames(accList[['MPP']]) = accList[['MPP']]$gene_id
rownames(accList[['HBCC']]) = accList[['HBCC']]$gene_id

res_acc = run_meta_analysis( accList, "FE") %>% getGeneSymbol
res_acc$Dataset = "Meta-analysis"
```

## DLPFC
```{r plot.meta_analysis.dlpfc.Sex}
kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05")) %>%
  kable_styling(full_width = FALSE)
```

```{r plot.concordance.dlpfc.Sex, fig.height=20}

# get genes to highlight
idx = order(res_dlpfc$adj.P.Val)
showGenes = res_dlpfc$Symbol[idx[1:20]]
showGenes2 = res_dlpfc$Symbol[!(res_dlpfc$Chrom %in% c("X", "Y"))][idx[1:30]]

dlpfcList_autosome = lapply(dlpfcList, function(x){
  x[!(x$Chrom %in% c("X", "Y")),]
  })

resConcord_dlpfc = plot_concordance( dlpfcList_autosome, showGenes )

plot_grid(resConcord_dlpfc$fig_logFC, 
          resConcord_dlpfc$fig_tstat, 
          plotVolcano( res_dlpfc, showGenes),   
          plotVolcano( res_dlpfc[!(res_dlpfc$Chrom %in% c("X", "Y")),], showGenes2),
          ncol=1)
```

```{r plot.concordance.table.dlpfc.Sex, fig.width=6}
kable( resConcord_dlpfc$tab_pi ) %>%
  kable_styling(full_width = FALSE)

resConcord_dlpfc$tab_corr$P = format(resConcord_dlpfc$tab_corr$P, scientific=TRUE, digits=3)
kable( resConcord_dlpfc$tab_corr )  %>%
  kable_styling(full_width = FALSE)
```

```{r plot.forrest.dlpfc.Sex, fig.width=5, fig.height=20}  
figList = lapply(showGenes2[1:10], function(geneSymbol){
  plot_forrest( c(dlpfcList, list("Meta-analysis" = res_dlpfc)), geneSymbol)
})

plot_grid(plotlist=figList, ncol=1)
```

```{r plot.logFC_by_chr.dlpfc}
plot_circos_logFC( res_dlpfc ) + ggtitle("DLPFC sex differential expression")
```

### MPP
```{r DE.results.dlpfc.mpp, fig.width=10, fig.height=50}
figList = plot_sex_compare(mpp_dlpfc)
plot_grid(plotlist=figList, ncol=2) 
```

### HBCC
```{r DE.results.dlpfc.hbcc, fig.width=10, fig.height=50}
figList = plot_sex_compare(hbcc_dlpfc)
plot_grid(plotlist=figList, ncol=2) 
```

### Combine cohorts
```{r DE.results.dlpfc.combine.MA}
res = merge(res_dlpfc, mpp_dlpfc[,c("gene_id", "AveExpr")], by.x="row.names", by.y="gene_id")

ggplot(res, aes(AveExpr, logFC)) + geom_point() + theme_bw() + geom_smooth(method="loess")  + theme(aspect.ratio=1)
```

```{r DE.results.dlpfc.hbcc.plot_sex_compare, fig.width=10, fig.height=50}
figList = plot_sex_compare(res)
plot_grid(plotlist=figList, ncol=2) 
```

## ACC
```{r plot.meta_analysis.acc.Sex}
kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05")) %>%
  kable_styling(full_width = FALSE)
```

```{r plot.concordance.acc.Sex, fig.height=20}

# get genes to highlight
idx = order(res_acc$adj.P.Val)
showGenes = res_acc$Symbol[idx[1:20]]
showGenes2 = res_acc$Symbol[!(res_acc$Chrom %in% c("X", "Y"))][idx[1:30]]

accList_autosome = lapply(accList, function(x){
  x[!(x$Chrom %in% c("X", "Y")),]
  })

resConcord_acc = plot_concordance( accList_autosome, showGenes )

plot_grid(resConcord_acc$fig_logFC, 
          resConcord_acc$fig_tstat, 
          plotVolcano( res_acc, showGenes),   
          plotVolcano( res_acc[!(res_acc$Chrom %in% c("X", "Y")),], showGenes2),
          ncol=1)
```

```{r plot.concordance.table.acc.Sex, fig.width=6}
kable( resConcord_acc$tab_pi ) %>%
  kable_styling(full_width = FALSE)

resConcord_acc$tab_corr$P = format(resConcord_acc$tab_corr$P, scientific=TRUE, digits=3)
kable( resConcord_acc$tab_corr )  %>%
  kable_styling(full_width = FALSE)
```

```{r plot.forrest.acc.Sex, fig.width=5, fig.height=20}  
figList = lapply(showGenes2[1:10], function(geneSymbol){
  plot_forrest( c(accList, list("Meta-analysis" = res_acc)), geneSymbol)
})

plot_grid(plotlist=figList, ncol=1)
```

```{r plot.logFC_by_chr.acc}
plot_circos_logFC( res_acc) + ggtitle("ACC sex differential expression")
```

### MPP
```{r DE.results.acc.mpp, fig.width=10, fig.height=50}
figList = plot_sex_compare(mpp_acc)
plot_grid(plotlist=figList, ncol=2) 
```

### HBCC
```{r DE.results.acc.hbcc, fig.width=10, fig.height=50}
figList = plot_sex_compare(hbcc_dsACC)
plot_grid(plotlist=figList, ncol=2) 
```

### Combine cohorts
```{r DE.results.acc.combine, fig.width=10, fig.height=50}
res = merge(res_acc, mpp_acc[,c("gene_id", "AveExpr")], by.x="row.names", by.y="gene_id")

figList = plot_sex_compare(res)
plot_grid(plotlist=figList, ncol=2)

```

## Compare across brain regions
```{r plot.concordance.combine, fig.height=15}

resCombine = merge(res_dlpfc, res_acc, by="row.names")

res_dlpfc$t = with(res_dlpfc, logFC / se)
res_acc$t = with(res_acc, logFC / se)

resConcord = plot_concordance( list(
        DLPFC = res_dlpfc[!(res_dlpfc$Chrom %in% c("X", "Y")),], 
        ACC = res_acc[!(res_acc$Chrom %in% c("X", "Y")),]), showGenes2 )

plot_grid(resConcord$fig_logFC, 
          resConcord$fig_tstat,     
          nrow=2) 
```

```{r plot.concordance.table.combine, fig.width=6}
kable( resConcord$tab_pi ) %>%
  kable_styling(full_width = FALSE)

resConcord$tab_corr$P = format(resConcord$tab_corr$P, scientific=TRUE, digits=3)
kable( resConcord$tab_corr )  %>%
  kable_styling(full_width = FALSE)
```

```{r store, eval=TRUE}
# Code
CODE <- Folder(name = "Meta-analysis on sex", parentId = "syn22315403") # Make sure these match the folder name and parent Id in the first chunk.
CODE <- synStore(CODE)

# Store meta-analysis
res_dlpfc %>%
  tibble::rownames_to_column(var = "gene_id") %>% 
  write.table(file = 'CMC_dlpfc_Sex_meta_analysis.tsv', sep = '\t', row.names=F, quote=F)
nEXP_OBJ = File('CMC_dlpfc_Sex_meta_analysis.tsv', 
                name = 'Sex Normalised, meta-analysis (DLPFC)', 
                parentId = CODE$properties$id)
nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
                    executed = thisFile)
res_acc %>%
  tibble::rownames_to_column(var = "gene_id") %>% 
  write.table(file = 'CMC_acc_Sex_meta_analysis.tsv', sep = '\t', row.names=F, quote=F)
nEXP_OBJ = File('CMC_acc_Sex_meta_analysis.tsv', 
                name = 'Sex Normalised, meta-analysis (ACC)', 
                parentId = CODE$properties$id)
nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
                    executed = thisFile)
```

### Source
[markdown](`r thisFile`)