### 2020-09-02 individual-cohorts/meta_analysis/meta-analysis_interaction.Rmd
Initial commit

---
title: "Meta-analysis on Sex*Dianosis interaction in MSSM-Penn-Pitt and HBCC ACC/DLPFC"
author: "KM and GH"
date: "`r date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r knit2synapse, eval=FALSE, echo=FALSE}
# rmarkdown::render("individual-cohorts/meta_analysis/meta-analysis_interaction.Rmd")

library(synapser)
library(knit2synapse)
synLogin()
# knit2synapse::

createAndKnitToFolderEntity <- function(file, parentId, folderName,
                                        wikiName=NULL, overwrite=FALSE, knitmd=TRUE, ...) {

  entity <- synapser::Folder(parentId=parentId, name=folderName)
  entity <- synapser::synStore(entity, ...)

  knitfile2synapse(file=file, owner=entity, wikiName=wikiName, 
                   overwrite=overwrite, knitmd=knitmd, parentWikiId=NULL)
}

createAndKnitToFolderEntity(file = "individual-cohorts/meta_analysis/meta-analysis_interaction.Rmd",
                                          parentId ="syn22315403",
                                          folderName = 'Meta-analysis on Sex*Diagnosis interaction',
                                          overwrite=TRUE, 
                                          knitmd=TRUE)
```

```{r setup, include=FALSE}
library(synapser)

synLogin()
suppressPackageStartupMessages({
library(ggplot2)
library(dplyr)
library(ComplexHeatmap)
library(readxl)
library(gtools)
library(corrplot)
library(pvclust)
library(metafor)
library(cowplot)
library(kableExtra)
library(githubr)
library(org.Hs.eg.db)
library(ggrepel)
library(metafor)
library(knitr)
library(patchwork)
})

source("../../common_functions/common_functions.R")

knitr::opts_chunk$set(
  echo = FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=FALSE, eval=TRUE}
thisFileName <- 'meta-analysis_interaction.Rmd'

# Github link
thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('individual-cohorts/meta_analysis/',thisFileName))
```

```{r input_data}
mpp_acc <- readr::read_tsv(synGet("syn22276990")$path) %>% 
  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% getGeneSymbol( 'gene_id' )

mpp_dlpfc <- readr::read_tsv(synGet("syn22277179")$path) %>% 
  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% getGeneSymbol( 'gene_id' )

hbcc_dsACC <- readr::read_tsv(synGet("syn22086829")$path) %>% 
  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val)  %>% getGeneSymbol( 'gene_id' )

hbcc_dlpfc <- readr::read_tsv(synGet("syn22280364")$path) %>% 
  dplyr::select(Comparison, gene_id, AveExpr, t, logFC, P.Value, adj.P.Val) %>% getGeneSymbol( 'gene_id' )

Ids <- c("syn22276990", "syn22277179", "syn22086829", "syn22280364")
```

# Control vs SCZ
```{r meta_analysis.SZ}
make_list <- function(mpp, hbcc, string) {
  list <- list(MPP = mpp %>%
                 unique() %>% 
                 dplyr::filter(Comparison == string) %>% 
                 as.data.frame,
               HBCC = hbcc %>% 
                 unique() %>% 
                 dplyr::filter(Comparison == string) %>% 
                 as.data.frame)
}

dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Control_vs_SCZ")
accList <- make_list(mpp_acc, hbcc_dsACC, string = "Control_vs_SCZ")

rownames(dlpfcList[['MPP']]) = dlpfcList[['MPP']]$gene_id
rownames(dlpfcList[['HBCC']]) = dlpfcList[['HBCC']]$gene_id

res_dlpfc = run_meta_analysis( dlpfcList, "FE") %>% getGeneSymbol
res_dlpfc$Dataset = "Meta-analysis"

rownames(accList[['MPP']]) = accList[['MPP']]$gene_id
rownames(accList[['HBCC']]) = accList[['HBCC']]$gene_id

res_acc = run_meta_analysis( accList, "FE") %>% getGeneSymbol
res_acc$Dataset = "Meta-analysis"
```

## DLPFC

```{r plot.meta_analysisSCZ}
kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05")) %>%
  kable_styling(full_width = FALSE)
```

```{r plot.concordance.dlpfc, fig.height=20}

# get genes to highlight
idx = order(res_dlpfc$adj.P.Val)
showGenes = res_dlpfc$Symbol[idx[1:20]]

resConcord_dlpfc = plot_concordance( dlpfcList, showGenes )

plot_grid(resConcord_dlpfc$fig_logFC, 
          resConcord_dlpfc$fig_tstat, 
          plotVolcano( res_dlpfc, showGenes),   
          nrow=3)
```

```{r plot.concordance.table.dlpfc, fig.width=6}
kable( resConcord_dlpfc$tab_pi ) %>%
  kable_styling(full_width = FALSE)

resConcord_dlpfc$tab_corr$P = format(resConcord_dlpfc$tab_corr$P, scientific=TRUE, digits=3)
kable( resConcord_dlpfc$tab_corr )  %>%
  kable_styling(full_width = FALSE)
```

```{r test, eval=FALSE}
# # forrest plot of meta-analysis
# geneSymbol = 'RCSD1'

# idx = with(res_dlpfc, (Symbol == geneSymbol) & !is.na(Symbol) )
# kable(res_dlpfc[idx,] )
```

```{r plot.forrest.dlpfc, fig.width=5, fig.height=20}  
figList = lapply(showGenes[1:10], function(geneSymbol){
  plot_forrest( c(dlpfcList, list("Meta-analysis" = res_dlpfc)), geneSymbol)
})

plot_grid(plotlist=figList, ncol=1)
```

## ACC

```{r plot.meta_analysisSCZ.acc}
kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))%>%
  kable_styling(full_width = FALSE)
```
```{r plot.concordance.acc, fig.height=20}

# get genes to highlight
idx = order(res_acc$adj.P.Val)
showGenes = res_acc$Symbol[idx[1:20]]

resConcord_acc = plot_concordance( accList, showGenes )

plot_grid(resConcord_acc$fig_logFC, 
          resConcord_acc$fig_tstat, 
          plotVolcano( res_acc, showGenes),   
          nrow=3)
```

```{r plot.concordance.table.acc, fig.width=6}
kable( resConcord_acc$tab_pi ) %>%
  kable_styling(full_width = FALSE)

resConcord_acc$tab_corr$P = format(resConcord_acc$tab_corr$P, scientific=TRUE, digits=3)
kable( resConcord_acc$tab_corr )  %>%
  kable_styling(full_width = FALSE)
```

```{r plot.forrest.acc, fig.width=5, fig.height=20}  
figList = lapply(showGenes[1:10], function(geneSymbol){
  plot_forrest( c(accList, list("Meta-analysis" = res_acc)), geneSymbol)
})

plot_grid(plotlist=figList, ncol=1)
```

## Compare across brain regions
```{r plot.concordance.combine, fig.height=15}

resCombine = merge(res_dlpfc, res_acc, by="row.names")

res_dlpfc$t = with(res_dlpfc, logFC / se)
res_acc$t = with(res_acc, logFC / se)

resConcord = plot_concordance( list(DLPFC = res_dlpfc, ACC = res_acc), showGenes )

plot_grid(resConcord$fig_logFC, 
          resConcord$fig_tstat,     
          nrow=2) 
```

```{r plot.concordance.table.combine, fig.width=6}
kable( resConcord$tab_pi ) %>%
  kable_styling(full_width = FALSE)

resConcord$tab_corr$P = format(resConcord$tab_corr$P, scientific=TRUE, digits=3)
kable( resConcord$tab_corr )  %>%
  kable_styling(full_width = FALSE)
```

# Control vs Other
```{r meta_analysis.Other}
dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Control_vs_Other")
accList <- make_list(mpp_acc, hbcc_dsACC, string = "Control_vs_Other")

rownames(dlpfcList[['MPP']]) = dlpfcList[['MPP']]$gene_id
rownames(dlpfcList[['HBCC']]) = dlpfcList[['HBCC']]$gene_id

res_dlpfc = run_meta_analysis( dlpfcList, "FE") %>% getGeneSymbol
res_dlpfc$Dataset = "Meta-analysis"

rownames(accList[['MPP']]) = accList[['MPP']]$gene_id
rownames(accList[['HBCC']]) = accList[['HBCC']]$gene_id

res_acc = run_meta_analysis( accList, "FE") %>% getGeneSymbol
res_acc$Dataset = "Meta-analysis"
```

## DLPFC
```{r plot.meta_analysisOther}
kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05")) %>%
  kable_styling(full_width = FALSE)
```

```{r plot.concordance.dlpfc.Other, fig.height=20}

# get genes to highlight
idx = order(res_dlpfc$adj.P.Val)
showGenes = res_dlpfc$Symbol[idx[1:20]]

resConcord_dlpfc = plot_concordance( dlpfcList, showGenes )

plot_grid(resConcord_dlpfc$fig_logFC, 
          resConcord_dlpfc$fig_tstat, 
          plotVolcano( res_dlpfc, showGenes),   
          nrow=3)
```

```{r plot.concordance.table.dlpfc.Other, fig.width=6}
kable( resConcord_dlpfc$tab_pi ) %>%
  kable_styling(full_width = FALSE)

resConcord_dlpfc$tab_corr$P = format(resConcord_dlpfc$tab_corr$P, scientific=TRUE, digits=3)
kable( resConcord_dlpfc$tab_corr )  %>%
  kable_styling(full_width = FALSE)
```

```{r plot.forrest.dlpfc.Other, fig.width=5, fig.height=20}  
figList = lapply(showGenes[1:10], function(geneSymbol){
  plot_forrest( c(dlpfcList, list("Meta-analysis" = res_dlpfc)), geneSymbol)
})

plot_grid(plotlist=figList, ncol=1)
```

## ACC

```{r plot.meta_analysisSCZ.acc.Other}
kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))%>%
  kable_styling(full_width = FALSE)
```
```{r plot.concordance.acc.Other, fig.height=20}

# get genes to highlight
idx = order(res_acc$adj.P.Val)
showGenes = res_acc$Symbol[idx[1:20]]

resConcord_acc = plot_concordance( accList, showGenes )

plot_grid(resConcord_acc$fig_logFC, 
          resConcord_acc$fig_tstat, 
          plotVolcano( res_acc, showGenes),   
          nrow=3)
```

```{r plot.concordance.table.acc.Other, fig.width=6}
kable( resConcord_acc$tab_pi ) %>%
  kable_styling(full_width = FALSE)

resConcord_acc$tab_corr$P = format(resConcord_acc$tab_corr$P, scientific=TRUE, digits=3)
kable( resConcord_acc$tab_corr )  %>%
  kable_styling(full_width = FALSE)
```

```{r plot.forrest.acc.Other, fig.width=5, fig.height=20}  
figList = lapply(showGenes[1:10], function(geneSymbol){
  plot_forrest( c(accList, list("Meta-analysis" = res_acc)), geneSymbol)
})

plot_grid(plotlist=figList, ncol=1)
```

## Compare across brain regions
```{r plot.concordance.combine.Other, fig.height=15}

resCombine = merge(res_dlpfc, res_acc, by="row.names")

res_dlpfc$t = with(res_dlpfc, logFC / se)
res_acc$t = with(res_acc, logFC / se)

resConcord = plot_concordance( list(DLPFC = res_dlpfc, ACC = res_acc), showGenes )

plot_grid(resConcord$fig_logFC, 
          resConcord$fig_tstat,     
          nrow=2) 
```

```{r plot.concordance.table.combine.Other, fig.width=6}
kable( resConcord$tab_pi ) %>%
  kable_styling(full_width = FALSE)

resConcord$tab_corr$P = format(resConcord$tab_corr$P, scientific=TRUE, digits=3)
kable( resConcord$tab_corr )  %>%
  kable_styling(full_width = FALSE)
```

# Other vs SCZ
```{r meta_analysis.OtherSz}
dlpfcList <- make_list(mpp_dlpfc, hbcc_dlpfc, string = "Other_vs_SCZ")
accList <- make_list(mpp_acc, hbcc_dsACC, string = "Other_vs_SCZ")

rownames(dlpfcList[['MPP']]) = dlpfcList[['MPP']]$gene_id
rownames(dlpfcList[['HBCC']]) = dlpfcList[['HBCC']]$gene_id

res_dlpfc = run_meta_analysis( dlpfcList, "FE") %>% getGeneSymbol
res_dlpfc$Dataset = "Meta-analysis"

rownames(accList[['MPP']]) = accList[['MPP']]$gene_id
rownames(accList[['HBCC']]) = accList[['HBCC']]$gene_id

res_acc = run_meta_analysis( accList, "FE") %>% getGeneSymbol
res_acc$Dataset = "Meta-analysis"
```

## DLPFC
```{r plot.meta_analysis.OtherSz}
kable(table(res_dlpfc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05")) %>%
  kable_styling(full_width = FALSE)
```

```{r plot.concordance.dlpfc..OtherSz, fig.height=20}

# get genes to highlight
idx = order(res_dlpfc$adj.P.Val)
showGenes = res_dlpfc$Symbol[idx[1:20]]

resConcord_dlpfc = plot_concordance( dlpfcList, showGenes )

plot_grid(resConcord_dlpfc$fig_logFC, 
          resConcord_dlpfc$fig_tstat, 
          plotVolcano( res_dlpfc, showGenes),   
          nrow=3)
```

```{r plot.concordance.table.dlpfc.OtherSz, fig.width=6}
kable( resConcord_dlpfc$tab_pi ) %>%
  kable_styling(full_width = FALSE)

resConcord_dlpfc$tab_corr$P = format(resConcord_dlpfc$tab_corr$P, scientific=TRUE, digits=3)
kable( resConcord_dlpfc$tab_corr )  %>%
  kable_styling(full_width = FALSE)
```

```{r plot.forrest.dlpfc..OtherSz, fig.width=5, fig.height=20}  
figList = lapply(showGenes[1:10], function(geneSymbol){
  plot_forrest( c(dlpfcList, list("Meta-analysis" = res_dlpfc)), geneSymbol)
})

plot_grid(plotlist=figList, ncol=1)
```

## ACC

```{r plot.meta_analysisSCZ.acc.OtherSz}
kable(table(res_acc$adj.P.Val < 0.05), col.names = c("", "Num. genes with p-value < 0.05"))%>%
  kable_styling(full_width = FALSE)
```
```{r plot.concordance.acc.OtherSz, fig.height=20}

# get genes to highlight
idx = order(res_acc$adj.P.Val)
showGenes = res_acc$Symbol[idx[1:20]]

resConcord_acc = plot_concordance( accList, showGenes )

plot_grid(resConcord_acc$fig_logFC, 
          resConcord_acc$fig_tstat, 
          plotVolcano( res_acc, showGenes),   
          nrow=3)
```

```{r plot.concordance.table.acc.OtherSz, fig.width=6}
kable( resConcord_acc$tab_pi ) %>%
  kable_styling(full_width = FALSE)

resConcord_acc$tab_corr$P = format(resConcord_acc$tab_corr$P, scientific=TRUE, digits=3)
kable( resConcord_acc$tab_corr )  %>%
  kable_styling(full_width = FALSE)
```

```{r plot.forrest.acc.OtherSz, fig.width=5, fig.height=20}  
figList = lapply(showGenes[1:10], function(geneSymbol){
  plot_forrest( c(accList, list("Meta-analysis" = res_acc)), geneSymbol)
})

plot_grid(plotlist=figList, ncol=1)
```

## Compare across brain regions
```{r plot.concordance.combine.OtherSz, fig.height=15}

resCombine = merge(res_dlpfc, res_acc, by="row.names")

res_dlpfc$t = with(res_dlpfc, logFC / se)
res_acc$t = with(res_acc, logFC / se)

resConcord = plot_concordance( list(DLPFC = res_dlpfc, ACC = res_acc), showGenes )

plot_grid(resConcord$fig_logFC, 
          resConcord$fig_tstat,     
          nrow=2) 
```

```{r plot.concordance.table.combine.OtherSz, fig.width=6}
kable( resConcord$tab_pi ) %>%
  kable_styling(full_width = FALSE)

resConcord$tab_corr$P = format(resConcord$tab_corr$P, scientific=TRUE, digits=3)
kable( resConcord$tab_corr )  %>%
  kable_styling(full_width = FALSE)
```

```{r store, eval=FALSE}
# Code
CODE <- Folder(name = "Meta-analysis on Sex*Diagnosis interaction", parentId = "syn22315403") # Make sure these match the folder name and parent Id in the first chunk.
CODE <- synStore(CODE)

# Store meta-analysis
res_dlpfc %>%
  tibble::rownames_to_column(var = "gene_id") %>% 
  write.table(file = 'CMC_dlpfc_interaction_meta_analysis.tsv', sep = '\t', row.names=F, quote=F)
nEXP_OBJ = File('CMC_dlpfc_interaction_meta_analysis.tsv', 
                name = 'Interaction meta-analysis (DLPFC)', 
                parentId = CODE$properties$id)
nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
                    executed = thisFile)
res_acc %>%
  tibble::rownames_to_column(var = "gene_id") %>% 
  write.table(file = 'CMC_acc_Interaction_meta_analysis.tsv', sep = '\t', row.names=F, quote=F)
nEXP_OBJ = File('CMC_acc_Interaction_meta_analysis.tsv', 
                name = 'Interaction meta-analysis (ACC)', 
                parentId = CODE$properties$id)
nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
                    executed = thisFile)
```

### Source
[markdown](`r thisFile`)