### 2020-09-04 combined_analysis/combine_all_data.Rmd
Fix Dx

---
title: "Joint analysis of read count data"
subtitle: ''
author: "Developed by [Gabriel Hoffman](http://gabrielhoffman.github.io/)"
date: "Run on `r Sys.time()`"
documentclass: article
output: 
  html_document:
  toc: true
  smart: false
---

<!---

cd /hpc/users/hoffmg01/work/covarr-de
# rm -rf molec_profiling_cache/
ml git pandoc
git pull
R

system("git pull")

rmarkdown::render("combined_analysis/combine_all_data.Rmd")

--->

--->

```{r knit2synapse, eval=FALSE}
library(synapser)
library(knit2synapse)
synLogin()
knit2synapse::createAndKnitToFolderEntity(file = "combined_analysis/combine_all_data.Rmd",
                                          parentId ="syn22231629",
                                          folderName = 'Joint analysis of count data',
                                          overwrite=TRUE, 
                                          knitmd=TRUE)
```

GSEA and dream from raw counts

```{r setup, include=FALSE}
suppressPackageStartupMessages({
library(synapser)

synLogin()

library(ggplot2)
library(dplyr)
library(ComplexHeatmap)
library(readxl)
library(gtools)
library(corrplot)
library(pvclust)
library(metafor)
library(data.table)
library(cowplot)
library(kableExtra)
library(githubr)
library(ggrepel)
library(knitr)
library(patchwork)
library(plyr)
library(BiocParallel)

library(limma)
library(edgeR)
library(variancePartition)

})
source("../common_functions/common_functions.R")

knitr::opts_chunk$set(
  echo = FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r read.counts}
# Read raw counts from MPP and HBCC
counts.MPP = fread(synGet('syn21867938')$path, data.table=FALSE)
rownames(counts.MPP) = counts.MPP$Geneid
counts.MPP = counts.MPP[,-c(1:6)]

# HBCC - DLFPC
counts.HBCC = fread(synGet('syn21886235')$path, data.table=FALSE)
rownames(counts.HBCC) = counts.HBCC$Geneid
counts.HBCC = counts.HBCC[,-c(1:6)]

# HBCC - sgACC
counts.HBCC4 = fread(synGet('syn21867866')$path, data.table=FALSE)
rownames(counts.HBCC4) = counts.HBCC4$Geneid
counts.HBCC4 = counts.HBCC4[,-c(1:6)]

geneCountsMerged = cbind(counts.MPP, counts.HBCC, counts.HBCC4)
```

### Data download
#### Obtain effective count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
downloadFile <- function(id){
  fread(synGet(id)$path, data.table = F)
}
downloadFile_version <- function(id , version){
  fread(synGet(id, version = version)$path, data.table = F)
}

# Get RNASeq QCmetadata
METADATA_QC_DLPFC_ID = 'syn16816488' 
ALL_USED_IDs = METADATA_QC_DLPFC_ID
metadata = downloadFile_version(METADATA_QC_DLPFC_ID, version = 19) %>%
  dplyr::select(`Individual_ID`, Institution, Cohort, `Reported_Gender`, Sex, Ethnicity, ageOfDeath, `PMI_(in_hours)`, Dx, 
                Sample_RNA_ID, one_of('rnaSeq_isolation:RIN',
                                      'rnaSeq_dissection:Brain_Region', 
                                      'rnaSeq_isolation:RNA_Isolation_Batch',
                                      'rnaSeq_report:Ribozero_Batch', 'rnaSeq_report:Library_Batch', 
                                      'rnaSeq_report:Flowcell_Batch', 'rnaSeq_report:Exclude?',
                                      'rnaSeq_report:Mapped_Reads',
                                      'rnaSeq_report:Intragenic_Rate','rnaSeq_report:Intronic_Rate',
                                      'rnaSeq_report:Intergenic_Rate','rnaSeq_report:Genes_Detected',
                                      'rnaSeq_report:Expression_Profiling_Efficiency',
                                      'rnaSeq_report:rRNA_Rate',
                                      'rnaSeq_report:Total_Reads', 'rnaSeq_report:Percent_Aligned','rnaSeq_report:Transcripts_Detected'))

# Get ancestry vector calculated using gemtools
ANCESTRY_ID = 'syn9922992'
ALL_USED_IDs = c(ALL_USED_IDs, ANCESTRY_ID)
ANCESTRY.HBCC = downloadFile(ANCESTRY_ID) %>% 
  plyr::rename(c(ID = 'SNP_report:Genotyping_Sample_ID'))

ANCESTRY_ID = 'syn2511399'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = ANCESTRY_ID
ANCESTRY.MPP = downloadFile(ANCESTRY_ID) %>% 
  plyr::rename(c('DNA_report..Genotyping.Sample_ID' = 'SNP_report:Genotyping_Sample_ID'))

ANCESTRY = rbind(ANCESTRY.HBCC[,colnames(ANCESTRY.MPP)[-2]], ANCESTRY.MPP[,colnames(ANCESTRY.MPP)[-2]])

# Get genotype ids from synapse. 
GENOTYPE_ID = 'syn16816490'
ALL_USED_IDs = c(ALL_USED_IDs, GENOTYPE_ID)
GENOTYPE = downloadFile(GENOTYPE_ID) %>% 
  dplyr::select(Individual_ID, `SNP_report:Genotyping_Sample_ID`, `SNP_report:Exclude?`) %>% 
  dplyr::inner_join(ANCESTRY) %>% 
  dplyr::filter(is.na(`SNP_report:Exclude?`))
```

### Data preprocessing
```{r preprocess.data}
# Merge metadata
METADATA = metadata %>%
  dplyr::left_join(GENOTYPE) %>%      
  dplyr::rename(Region = `rnaSeq_dissection:Brain_Region`,
                PMI = `PMI_(in_hours)`,
                RIN = `rnaSeq_isolation:RIN`, 
                ReportExclude = `rnaSeq_report:Exclude?`,
                GenotypeExclude = `SNP_report:Exclude?`,
                SampleID = Sample_RNA_ID,
                LibraryBatch = `rnaSeq_report:Library_Batch`,
                FlowcellBatch = `rnaSeq_report:Flowcell_Batch`,
                IsolationBatch = `rnaSeq_isolation:RNA_Isolation_Batch`, #isolation batch to replace ribozero batch
                RibozeroBatch = `rnaSeq_report:Ribozero_Batch`,
                MappedReads = `rnaSeq_report:Mapped_Reads`,
                IntragenicRate = `rnaSeq_report:Intragenic_Rate`, IntronicRate = `rnaSeq_report:Intronic_Rate`, 
                IntergenicRate = `rnaSeq_report:Intergenic_Rate`, GenesDetected = `rnaSeq_report:Genes_Detected`,
                ExpProfEfficiency = `rnaSeq_report:Expression_Profiling_Efficiency`, rRNARate = `rnaSeq_report:rRNA_Rate`,
                TotalReads = `rnaSeq_report:Total_Reads`, AlignmentRate = `rnaSeq_report:Percent_Aligned`, TranscriptsDetected = `rnaSeq_report:Transcripts_Detected`) %>%
  dplyr::select(SampleID, Individual_ID, Institution, Cohort, Reported_Gender,Sex, Ethnicity, ageOfDeath, PMI, Dx, Region,
                RIN, ReportExclude, GenotypeExclude, EV.1, EV.2, EV.3, EV.4, EV.5,
                LibraryBatch, FlowcellBatch, IsolationBatch, RibozeroBatch, MappedReads, TotalReads, GenesDetected, AlignmentRate,
                IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate, TranscriptsDetected) %>%
  		 		 dplyr::filter(SampleID %in% colnames(geneCountsMerged), !is.na(SampleID)) %>%
  		 		 dplyr::filter(Region %in% c('DLPFC', 'ACC')) %>%
			 	 dplyr::mutate(Cohort = forcats::fct_recode(Cohort, `MSSM-Penn-Pitt` = 'MSSM-Penn-Pitt', `NIMH-HBCC`='NIMH-HBCC', `NIMH-HBCC`='NIMH-HBCC-4')) %>%
			 	 dplyr::filter(Dx %in% c('Control', 'BP', "SCZ"))
			 	 #%>%
			 	 # dplyr::mutate(Dx_crude = forcats::fct_recode(Dx, Other = "AFF", Other = "BP", Other = "undetermined", Control = "Control", SCZ = "SCZ"))

ind = METADATA$SampleID[which(METADATA$ReportExclude == 1 | METADATA$GenotypeExclude)]
writeLines(paste('Following',length(ind),'samples are marked exclude'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$Ethnicity) | is.na(METADATA$Institution) | is.na(METADATA$Dx)]
writeLines(paste('Following', length(ind), 'counts are missing any metadata'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$PMI)]
writeLines(paste('Following', length(ind), 'counts are missing PMI'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$Reported_Gender)]
writeLines(paste('Following', length(ind), 'counts are missing gender'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$ageOfDeath)]
writeLines(paste('Following', length(ind), 'counts are missing age of death'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$EV.1)]
writeLines(paste('Following', length(ind), 'counts are missing ancestry information'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))
```

```{r preprocess.data1, results='asis'}
# Match covariates to expression data
indToRetain = intersect(METADATA$SampleID, colnames(geneCountsMerged))

geneCountsMerged = geneCountsMerged[,indToRetain]

rownames(METADATA) = METADATA$SampleID
METADATA = METADATA[indToRetain,]

METADATA %>% 
  group_by(Dx) %>%  
  dplyr::summarise(count = n()) %>% 
  kable() 
```

```{r process.counts}

# keep genes that are express in at least one combination of cohort and gender
isexpr.lst = dlply(METADATA, .(Cohort, Reported_Gender), .fun = function(mtd){
	rowSums(cpm(geneCountsMerged[,mtd$SampleID])>1) >= 0.5*ncol(geneCountsMerged[,mtd$SampleID])
	})

isexpr = do.call(rbind, isexpr.lst)
isexpr = colSums(isexpr) > 0
table(isexpr)

# normalize counts
dge = DGEList( geneCountsMerged[isexpr,] )
dge = calcNormFactors(dge) # TMM

xtabs(~ Region + Cohort, METADATA)
xtabs(~ Institution + Cohort, METADATA)

# FORMULAs

# HBCC - DLPFC: 
# "Dx + scale(RIN) + ageOfDeath + (1 | Reported_Gender) + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate)"

# HBCC dsACC: 
# Dx + scale(RIN) + (1 | Reported_Gender) + ageOfDeath + (1 | IsolationBatch) + scale(IntragenicRate) + scale(IntronicRate) + scale(IntergenicRate)"

# MPP ACC
 # Dx + scale(RIN) + scale(ageOfDeath) + (1 | Institution) + (1 | Reported_Gender) + scale(RIN2) + scale(IntronicRate) + scale(rRNARate) + scale(IntragenicRate) + scale(IntergenicRate)

# MPP DLPFC
# Dx + scale(RIN2) + (1 | Institution) + scale(ageOfDeath) + scale(RIN) + (1 | Reported_Gender) + scale(IntronicRate) + scale(IntragenicRate)

form = ~ (1|Dx) + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)

vobj = voomWithDreamWeights(dge, form, METADATA, save.plot=TRUE, BPPARAM = SnowParam(18))

plot_voom(vobj)
```

### Test marginals
```{r differential.expression.simple}

form = ~ 0 + Dx + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)

# get contrasts
Lbase = getContrast( vobj, form, METADATA, "RIN")
Lbase[] = 0

# SCZ
L_SCZ = Lbase
L_SCZ['DxControl'] = -1 
L_SCZ['DxSCZ'] = 1

# BP
L_BP = Lbase
L_BP['DxControl'] = -1 
L_BP['DxBP'] = 1

# BP
L_sex = Lbase
L_sex['Reported_GenderMale'] = 1

L = cbind(	Scz = L_SCZ, 
			BP 	= L_BP, 
			Sex = L_sex)

plotContrasts(L) + theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust=1))

fit_marginal = dream(vobj, form, METADATA, L=L,BPPARAM = SnowParam(18))
```

```{r plots.marginal}
# Disease
fig1 = plotVolcano( topTable(fit_marginal, "Scz", number=Inf) ) + ggtitle("SCZ")
fig2 = plotVolcano( topTable(fit_marginal, "BP", number=Inf) ) + ggtitle("BP")

plot_grid( fig1, fig2, ncol=2)

# Sex
tab = topTable(fit_marginal, "Reported_GenderMale", number=Inf) %>% getGeneSymbol

fig3 = plotVolcano( tab ) + ggtitle("Sex")
fig4 = plotVolcano( tab[tab$Chrom %in% 1:22,] ) + ggtitle("Sex")

plot_grid( fig3, fig4, ncol=2)
```

```{r exit, cache=FALSE}
knitr::knit_exit()
```

### Test Diagnosis * Sex
```{r differential.expression.interaction.DxSex}

form = ~ 0 + Dx*Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)

# get contrasts
Lbase = getContrast( vobj, form, METADATA, "RIN")
Lbase[] = 0

# SCZ by sex
L_SCZ = Lbase
L_SCZ['DxSCZ:Reported_GenderMale'] = 1
L_SCZ['DxControl:Reported_GenderMale'] = -1
L_SCZ['DxSCZ'] = -1
L_SCZ['DxControl'] = 1 
# (A - B) - (C - D) = A - B - C + D

L = cbind(	SczSex = L_SCZ)

plotContrasts(L) + theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust=1))

fit_interaction_DxSex = dream(vobj, form, METADATA, L=L,BPPARAM = SnowParam(18))
```

### Test Region * Sex
```{r differential.expression.interaction.RegionSex, eval=FALSE}

form = ~ 0 + Dx + Region*Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Cohort) + (1|Individual_ID)

# get contrasts
Lbase = getContrast( vobj, form, METADATA, "RIN")
Lbase[] = 0

# SCZ by sex
L_SCZ = Lbase
L_SCZ['DxSCZ:Reported_GenderMale'] = 1
L_SCZ['DxControl:Reported_GenderMale'] = -1
L_SCZ['DxSCZ'] = -1
L_SCZ['DxControl'] = 1 
# (A - B) - (C - D) = A - B - C + D

L = cbind(	SczSex = L_SCZ)

plotContrasts(L) + theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust=1))

fit_interaction_RegionSex = dream(vobj, form, METADATA, L=L,BPPARAM = SnowParam(18))
```

### Test Region * Disease
```{r differential.expression.interaction.RegionDx, eval=FALSE}

form = ~ 0 + Region*Dx + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Cohort) + (1|Individual_ID)

# get contrasts
Lbase = getContrast( vobj, form, METADATA, "RIN")
Lbase[] = 0

# SCZ by sex
L_SCZ = Lbase
L_SCZ['DxSCZ:Reported_GenderMale'] = 1
L_SCZ['DxControl:Reported_GenderMale'] = -1
L_SCZ['DxSCZ'] = -1
L_SCZ['DxControl'] = 1 
# (A - B) - (C - D) = A - B - C + D

L = cbind(	SczSex = L_SCZ)

plotContrasts(L) + theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust=1))

fit_interaction_RegionSex = dream(vobj, form, METADATA, L=L,BPPARAM = SnowParam(18))
```