### 2020-09-16 combined_analysis/combine_all_DLPFC.Rmd
Inital commit

---
title: "CommonMind Consortium: DLPFC"
subtitle: "Analysis of read count data by cohort"
author:
- name: "[Gabriel Hoffman](http://gabrielhoffman.github.io)$^{1}$ and Kelsey Montgomery$^{2}$"
  affiliation: | 
    [1] Icahn School of Medicine at Mount Sinai, New York
    [2] Sage Bionetworks, Seattle, WA
date: "Run on `r Sys.time()`"
documentclass: article
output:
  html_document:
    toc: true
    toc_float: true
---

<!---

cd /hpc/users/hoffmg01/work/covarr-de
# rm -rf molec_profiling_cache/
ml git pandoc
git pull
R

system("git pull")

rmarkdown::render("combined_analysis/combine_all_DLPFC.Rmd")

--->

```{r knit2synapse, eval= FALSE}
library(synapser)
library(knit2synapse)
synLogin()
knit2synapse::createAndKnitToFolderEntity(file = "combined_analysis/combine_all_DLPFC.Rmd",
                                          parentId ="syn22231629",
                                          folderName = 'Joint analysis of DLPFC count data',
                                          overwrite=TRUE, 
                                          knitmd=TRUE)
```

TODO

zenith 1 sided

zenith plots
replication or correlation as a function of expression magnitude

difference networks
Number of DE genes

<!---
DONE

Effect of ilr should depend on Instaition
suffix 1.
zenith for Sex with autosomes

--->

```{r setup, include=FALSE}
suppressPackageStartupMessages({
library(synapser)

synLogin()

library(ggplot2)
library(dplyr)
library(ComplexHeatmap)
library(readxl)
library(gtools)
library(corrplot)
library(pvclust)
library(metafor)
library(data.table)
library(cowplot)
library(kableExtra)
library(githubr)
library(ggrepel)
library(corrplot)
library(zenith)
library(compositions)
library(knitr)
library(patchwork)
library(plyr) 
library(BiocParallel)

library(limma)
library(edgeR)
library(variancePartition)

}) 
source("../common_functions/common_functions.R")

knitr::opts_chunk$set(
  echo = FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r read.counts}
# Read raw counts from MPP and HBCC
counts.MPP = fread(synGet('syn21867938')$path, data.table=FALSE)
rownames(counts.MPP) = counts.MPP$Geneid
counts.MPP = counts.MPP[,-c(1:6)]

# HBCC - DLFPC
counts.HBCC = fread(synGet('syn21886235')$path, data.table=FALSE)
rownames(counts.HBCC) = counts.HBCC$Geneid
counts.HBCC = counts.HBCC[,-c(1:6)]

# # HBCC - sgACC
counts.HBCC4 = fread(synGet('syn21867866')$path, data.table=FALSE)
rownames(counts.HBCC4) = counts.HBCC4$Geneid
counts.HBCC4 = counts.HBCC4[,-c(1:6)]

geneCountsMerged = cbind(counts.MPP, counts.HBCC, counts.HBCC4)
```

### Data download
#### Obtain effective count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
downloadFile <- function(id){
  fread(synGet(id)$path, data.table = F)
}
downloadFile_version <- function(id , version){
  fread(synGet(id, version = version)$path, data.table = F)
}

# Get RNASeq QCmetadata
METADATA_QC_DLPFC_ID = 'syn16816488' 
ALL_USED_IDs = METADATA_QC_DLPFC_ID
metadata = downloadFile_version(METADATA_QC_DLPFC_ID, version = 19) %>%
  dplyr::select(`Individual_ID`, Institution, Cohort, `Reported_Gender`, Sex, Ethnicity, ageOfDeath, `PMI_(in_hours)`, Dx, 
                Sample_RNA_ID, one_of('rnaSeq_isolation:RIN',
                                      'rnaSeq_dissection:Brain_Region', 
                                      'rnaSeq_isolation:RNA_Isolation_Batch',
                                      'rnaSeq_report:Ribozero_Batch', 'rnaSeq_report:Library_Batch', 
                                      'rnaSeq_report:Flowcell_Batch', 'rnaSeq_report:Exclude?',
                                      'rnaSeq_report:Mapped_Reads',
                                      'rnaSeq_report:Intragenic_Rate','rnaSeq_report:Intronic_Rate',
                                      'rnaSeq_report:Intergenic_Rate','rnaSeq_report:Genes_Detected',
                                      'rnaSeq_report:Expression_Profiling_Efficiency',
                                      'rnaSeq_report:rRNA_Rate',
                                      'rnaSeq_report:Total_Reads', 'rnaSeq_report:Percent_Aligned','rnaSeq_report:Transcripts_Detected'))

# Get ancestry vector calculated using gemtools
ANCESTRY_ID = 'syn9922992'
ALL_USED_IDs = c(ALL_USED_IDs, ANCESTRY_ID)
ANCESTRY.HBCC = downloadFile(ANCESTRY_ID) %>% 
  plyr::rename(c(ID = 'SNP_report:Genotyping_Sample_ID'))

ANCESTRY_ID = 'syn2511399'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = ANCESTRY_ID
ANCESTRY.MPP = downloadFile(ANCESTRY_ID) %>% 
  plyr::rename(c('DNA_report..Genotyping.Sample_ID' = 'SNP_report:Genotyping_Sample_ID'))

ANCESTRY = rbind(ANCESTRY.HBCC[,colnames(ANCESTRY.MPP)[-2]], ANCESTRY.MPP[,colnames(ANCESTRY.MPP)[-2]])

# Get genotype ids from synapse. 
GENOTYPE_ID = 'syn16816490'
ALL_USED_IDs = c(ALL_USED_IDs, GENOTYPE_ID)
GENOTYPE = downloadFile(GENOTYPE_ID) %>% 
  dplyr::select(Individual_ID, `SNP_report:Genotyping_Sample_ID`, `SNP_report:Exclude?`) %>% 
  dplyr::inner_join(ANCESTRY) %>% 
  dplyr::filter(is.na(`SNP_report:Exclude?`))
```

### Data preprocessing
```{r preprocess.data}
# Merge metadata
METADATA = metadata %>%
  dplyr::left_join(GENOTYPE) %>%      
  dplyr::rename(Region = `rnaSeq_dissection:Brain_Region`,
                PMI = `PMI_(in_hours)`,
                RIN = `rnaSeq_isolation:RIN`, 
                ReportExclude = `rnaSeq_report:Exclude?`,
                GenotypeExclude = `SNP_report:Exclude?`,
                SampleID = Sample_RNA_ID,
                LibraryBatch = `rnaSeq_report:Library_Batch`,
                FlowcellBatch = `rnaSeq_report:Flowcell_Batch`,
                IsolationBatch = `rnaSeq_isolation:RNA_Isolation_Batch`, #isolation batch to replace ribozero batch
                RibozeroBatch = `rnaSeq_report:Ribozero_Batch`,
                MappedReads = `rnaSeq_report:Mapped_Reads`,
                IntragenicRate = `rnaSeq_report:Intragenic_Rate`, IntronicRate = `rnaSeq_report:Intronic_Rate`, 
                IntergenicRate = `rnaSeq_report:Intergenic_Rate`, GenesDetected = `rnaSeq_report:Genes_Detected`,
                ExpProfEfficiency = `rnaSeq_report:Expression_Profiling_Efficiency`, rRNARate = `rnaSeq_report:rRNA_Rate`,
                TotalReads = `rnaSeq_report:Total_Reads`, AlignmentRate = `rnaSeq_report:Percent_Aligned`, TranscriptsDetected = `rnaSeq_report:Transcripts_Detected`) %>%
  dplyr::select(SampleID, Individual_ID, Institution, Cohort, Reported_Gender,Sex, Ethnicity, ageOfDeath, PMI, Dx, Region,
                RIN, ReportExclude, GenotypeExclude, EV.1, EV.2, EV.3, EV.4, EV.5,
                LibraryBatch, FlowcellBatch, IsolationBatch, RibozeroBatch, MappedReads, TotalReads, GenesDetected, AlignmentRate,
                IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate, TranscriptsDetected) %>%
  		 		 dplyr::filter(SampleID %in% colnames(geneCountsMerged), !is.na(SampleID)) %>%
  		 		 dplyr::filter(Region %in% c('DLPFC')) %>%
			 	 dplyr::mutate(Cohort = forcats::fct_recode(Cohort, `MSSM-Penn-Pitt` = 'MSSM-Penn-Pitt', `NIMH-HBCC`='NIMH-HBCC', `NIMH-HBCC`='NIMH-HBCC-4')) %>%
			 	 dplyr::filter(Dx %in% c('Control', "SCZ"))  %>%
         dplyr::filter(ageOfDeath > 17)

# Set Control to baseline
METADATA$Dx = factor(METADATA$Dx, c('Control', 'SCZ'))

ind = METADATA$SampleID[which(METADATA$ReportExclude == 1 | METADATA$GenotypeExclude)]
writeLines(paste('Following',length(ind),'samples are marked exclude'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$Ethnicity) | is.na(METADATA$Institution) | is.na(METADATA$Dx)]
writeLines(paste('Following', length(ind), 'counts are missing any metadata'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$PMI)]
writeLines(paste('Following', length(ind), 'counts are missing PMI'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$Reported_Gender)]
writeLines(paste('Following', length(ind), 'counts are missing gender'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$ageOfDeath)]
writeLines(paste('Following', length(ind), 'counts are missing age of death'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$EV.1)]
writeLines(paste('Following', length(ind), 'counts are missing ancestry information'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))
```

```{r cellFractionEstimates}
# include estimated cell fractions in METADATA
df_cellFractions = read.table(synGet('syn22333694')$path, row.names=1)

celFrac_ilr = ilr(df_cellFractions)
colnames(celFrac_ilr) = paste0("cellFrac_ilr_", 1:3)

METADATA = merge(METADATA, df_cellFractions, by.x = 'SampleID', by.y='row.names')
METADATA = merge(METADATA, celFrac_ilr, by.x = 'SampleID', by.y='row.names')
```

```{r preprocess.data1, results='asis'}
# Match covariates to expression data
indToRetain = intersect(METADATA$SampleID, colnames(geneCountsMerged))

geneCountsMerged = geneCountsMerged[,indToRetain]

rownames(METADATA) = METADATA$SampleID
```

## Summarize samples
```{r summary}
METADATA %>% group_by(Dx, Institution) %>% dplyr::summarise(count = n()) %>% 
  kable(caption="Diagnosis, Institution") %>% kable_styling(full_width=FALSE)

METADATA %>% group_by( Institution, Reported_Gender) %>% dplyr::summarise(count = n()) %>% 
  kable(caption="Institution, Sex") %>% kable_styling(full_width=FALSE)

METADATA %>% group_by(Dx, Institution, Reported_Gender) %>% dplyr::summarise(count = n()) %>% 
  kable(caption="Diagnosis, Institution, Sex") %>% kable_styling(full_width=FALSE)

ggplot(METADATA, aes(Reported_Gender, ageOfDeath)) + geom_boxplot() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1)
```

### Sex by cell fraction
```{r summary.sex, fig.width=8, fig.height=10}
figList = lapply(colnames(df_cellFractions), function(CT){
  ggplot(METADATA, aes_string('Reported_Gender', CT)) + geom_boxplot() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1)
})
plot_grid(plotlist=figList, ncol=1) 
```

### Age by cell fraction
```{r summary.Age, fig.width=8, fig.height=10}
figList = lapply(colnames(df_cellFractions), function(CT){
  ggplot(METADATA, aes_string('ageOfDeath', CT)) + geom_point() + geom_smooth(method="gam") + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1)
}) 
plot_grid(plotlist=figList, ncol=1) 
```

## Age by cell fraction log ratio
```{r summary.Age.log.ratio, fig.width=8, fig.height=8}
figList = lapply(colnames(celFrac_ilr), function(CT){
  ggplot(METADATA, aes_string('ageOfDeath', CT)) + geom_point() + geom_smooth(method="gam") + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1)
})
plot_grid(plotlist=figList, ncol=1) 
```

```{r process.counts}
# keep genes that are express in at least one combination of cohort and gender
isexpr.lst = dlply(METADATA, .(Cohort), .fun = function(mtd){
	rowSums(cpm(geneCountsMerged[,mtd$SampleID])>.5) >= .3*ncol(geneCountsMerged[,mtd$SampleID])
	})

isexpr = do.call(rbind, isexpr.lst)
isexpr = colSums(isexpr) == 2 # keep gene expressed in both cohorts 
table(isexpr)

# Normalize each cohort separately
dge.lst = lapply( levels(METADATA$Cohort), function(Cohort){
  # normalize counts
  dge = DGEList( geneCountsMerged[isexpr,METADATA$Cohort==Cohort] )
  calcNormFactors(dge, "none") # TMM
})
names(dge.lst) = levels(METADATA$Cohort)

# form = ~ (1|Dx) + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)

form.lst = list(
  `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
  `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)

vobj.lst = lapply(dge.lst, function(dge){

  i = match(colnames(dge), rownames(METADATA))
  info = METADATA[i,]

  form_use = form.lst[[info$Cohort[1]]]

  voomWithDreamWeights(dge, form_use, info, save.plot=TRUE, BPPARAM = SnowParam(18))
  })
names(vobj.lst) = names(dge.lst)
```

### Voom trend plots
```{r plot.voom, fig.height=5, fig.width=10}
fig1 = plot_voom(vobj.lst[[1]]) + ggtitle(names(vobj.lst)[1])
fig2 = plot_voom(vobj.lst[[2]]) + ggtitle(names(vobj.lst)[2])
plot_grid( fig1, fig2, nrow=1) 
```

<!---

# Examine genes on chrX and chrY

# dge = DGEList(geneCountsMerged)
# dge = calcNormFactors(dge, "none")
# geneExpr = cpm(dge, log=TRUE)

# df = merge( data.frame(Expression = geneExpr['ENSG00000229807.11',]), METADATA, by="row.names")
# ggplot(df, aes(Reported_Gender, Expression)) + geom_violin() + theme_bw() + facet_wrap(~Cohort)

# id = 'ENSG00000229807.11'
# isexpr[id]

# df = as.data.frame(vobj.lst[[2]]$voom.xy)
# with(df, plot(x,y, cex=.2, pch=20))
# with(df[id,], points(x,y, col="red", pch=20, cex=2))

--->

```{r Create.residuals}

### Create residuals
form.lst = list(
  `MSSM-Penn-Pitt`  = ~ Dx*Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
  `NIMH-HBCC`       = ~ Dx*Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)

# Compute residuals, but add back in Dx*Reported_Gender
f = function(fit){
  residuals(fit) + get_prediction( fit, ~ Dx*Reported_Gender)
}

residuals.lst = lapply( vobj.lst, function(vobj){

  i = match(colnames(vobj), rownames(METADATA))
  info = METADATA[i,]

  form_use = form.lst[[info$Cohort[1]]]

  # Compute residuals which adding back Diagnosis and Sex effects
  resid.lst = fitVarPartModel(vobj, form_use, info, fxn = f, BPPARAM = SnowParam(18))  
  do.call(rbind, resid.lst)
  })
names(residuals.lst) = names(vobj.lst)
```

```{r check.correlation}
ngenes = 300
C1 = cor(t(residuals.lst[[1]][1:ngenes,]))
C2 = cor(t(residuals.lst[[2]][1:ngenes,]))

df = data.table(A   = as.numeric(C1[lower.tri(C1)]),
                B   = as.numeric(C2[lower.tri(C2)]))

df$density <- get_density(df$A, df$B, n = 80)
ggplot(df, aes(A, B, color=density)) + geom_point() + geom_point(size=.4) + theme_bw(15) + theme(aspect.ratio=1, legend.position="bottom", plot.title = element_text(hjust = 0.5)) + geom_abline(color="red") + scale_color_viridis() + xlim(-1,1) + ylim(-1,1) + xlab(paste0('Correlation (', names(residuals.lst)[1], ')')) + ylab(paste0('Correlation (', names(residuals.lst)[2], ')'))
```

# Schizophrenia vs controls

### Test marginals for Scz vs controls
```{r differential.expression.simple}
form.lst = list(
  `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
  `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)

fit_marginal.lst = lapply( vobj.lst, function(vobj){

  i = match(colnames(vobj), rownames(METADATA))
  info = METADATA[i,]

  form_use = form.lst[[info$Cohort[1]]]

  fit = dream(vobj, form_use, info, BPPARAM = SnowParam(18))  
  eBayes( fit )
  })
names(fit_marginal.lst) = names(vobj.lst)

coefTest = c('DxSCZ', 'Reported_GenderMale', 'ageOfDeath')
```

```{r load.genesets, eval=TRUE}
# Load Gene Ontology database 
gs.GO = get_GeneOntology()
gs.MSigDB = get_MSigDB()

# convert from GeneSetCollection to list used by camera and zenith
gs.GO.collection = recodeToList( gs.GO )
gs.MSigDB.collection = recodeToList( gs.MSigDB )
```

```{r run.zenith, eval=TRUE}
res_zenith = lapply( names(fit_marginal.lst), function(key){

  fit = fit_marginal.lst[[key]]

  res.GO = run_zenith(  fit, 
                        coefTest,
                        gs.GO.collection)
  res.GO$Resource = "GO"   
  res.GO$Cohort = key

  res.MSigDB = run_zenith(fit, 
                          coefTest,
                          gs.MSigDB.collection ) 
  res.MSigDB$Resource = "MSigDB"
  res.MSigDB$Cohort = key

  rbind(res.GO, res.MSigDB)
})
names(res_zenith) = names(fit_marginal.lst)

# only autosomes
res_zenith_sex_autosomes = lapply( names(fit_marginal.lst), function(key){

  fit = fit_marginal.lst[[key]]

  # only keep autosomes, for Reported_GenderMale
  tab = coef(fit) %>% getGeneSymbol()
  geneids = rownames(tab)[tab$Chrom %in% 1:22]
  fit = fit[geneids,]

  res.GO = run_zenith(  fit, 
                        'Reported_GenderMale',
                        gs.GO.collection)
  res.GO$Resource = "GO"   
  res.GO$Cohort = key

  res.MSigDB = run_zenith(fit, 
                          'Reported_GenderMale',
                          gs.MSigDB.collection ) 
  res.MSigDB$Resource = "MSigDB"
  res.MSigDB$Cohort = key

  rbind(res.GO, res.MSigDB)
})
names(res_zenith_sex_autosomes) = names(fit_marginal.lst)
```

## Meta-analysis of zenith results
```{r zenith.meta, eval=TRUE}
res_meta_zenith = zenith_meta_analysis( res_zenith )                  
res_meta_zenith.sex_autosome = zenith_meta_analysis( res_zenith_sex_autosomes )
```

```{r get.signatures, fig.width=10, fig.height=5}
# Get DE signatures
df_DE_results = lapply(fit_marginal.lst, function(fit){
  res = lapply(coefTest, function(coef){
    res = topTable(fit, coef, number=Inf) 
    res$Coef = coef
    res %>% getGeneSymbol()  
    })
  names(res) = coefTest
  res
  })
names(df_DE_results) = names(fit_marginal.lst)
```

```{r test.scz, fig.height=5, fig.width=10}
# Schizophrenia testing
# extract from each cohort
coef = 'DxSCZ'
# ymin = min(sapply(df_DE_results, function(x) min(x[[coef]]$P.Value)))

figList.Scz = lapply( names(df_DE_results), function(key){
  df = df_DE_results[[key]][[coef]]

  main = paste(key, coef, sep = ': ' )
  plotVolcano( df, df$Symbol[1:10] ) + ggtitle( main ) #+ ylim(0, -log10(ymin))
})

# volcano plots
plot_grid(plotlist=figList.Scz, nrow=1)
```

## Meta-analysis across cohorts
```{r meta.analysis.DE}
# Store results for each coef and cohort
tab.de.lst = lapply(coefTest, function(coef){    
  tab.lst = lapply( df_DE_results, function(lst){
    lst[[coef]]
  })
  names(tab.lst) = names(df_DE_results)
  tab.lst
  })
names(tab.de.lst) = coefTest

# Create list preparing for meta-analysis for SCZ
df_meta = mclapply(coefTest, function(coef){
  # Run meta-analysis
  run_meta_analysis( tab.de.lst[[coef]], "FE") %>% getGeneSymbol
}, mc.cores=6)
names(df_meta) = coefTest
```

```{r concordance.scz, fig.height=6, fig.width=10}
# Concordance analysis
resConcord.SCZ = plot_concordance(tab.de.lst[[coef]], showGenes = df_meta[[coef]]$Symbol[1:20])

main = "Concordance of SCZ"
plot_grid(  resConcord.SCZ$fig_logFC + ggtitle(main),
            resConcord.SCZ$fig_tstat + ggtitle(main), nrow=1)

data.frame(Cohort = names(tab.de.lst[[coef]]),
          "# DE"  = resConcord.SCZ$n.de,
          pi1     = resConcord.SCZ$pi_single,
          check.names=FALSE) %>% 
          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)

resConcord.SCZ$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
resConcord.SCZ$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
```

```{r plot.meta.analysis.SCZ, fig.height=5, fig.width=10}
# volcano plot for meta-analysis
main = paste('Meta-analysis', coef, sep = ': ' )
plotVolcano( df_meta[[coef]], df_meta[[coef]]$Symbol[1:20] ) + ggtitle( main )

data.frame(pi1 = 1 - qvalue(df_meta[[coef]]$P)$pi0,
          `# DE` = sum(df_meta[[coef]]$adj.P.Val < 0.05),
          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for SCZ meta-analysis") %>% kable_styling(full_width=FALSE)

```

## Plot top examples
```{r plot.forrest.dlpfc, fig.width=8, fig.height=20}  
plot_gene = function(residuals.lst, df_meta, geneSymbol, variable, fill=c("#1E90FF", "#FF8D1E"), mode=""){

  ensGene = rownames(df_meta)[which(df_meta$Symbol == geneSymbol)]

  exprResid = lapply( names(residuals.lst), function(key){
      data.frame(Expression = residuals.lst[[key]][ensGene,])
    })
  exprResid = do.call(rbind, exprResid)

  df = merge(exprResid, METADATA, by="row.names")

  if( is.factor(df[[variable]]) || length(unique(df[[variable]])) < 5 ){
    fig = ggplot(df, aes_string(variable, 'Expression')) + geom_violin(aes_string(fill=variable)) + geom_boxplot(width=0.1, fill="grey80") + scale_fill_manual(values = fill)
  }else{    
    fig = ggplot(df, aes_string(variable, 'Expression')) + geom_point() + geom_smooth(method="lm") #+ scale_color_manual(values = fill)
  }

  fig + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + ylab('Expression residuals') + ggtitle(geneSymbol) + facet_wrap(~Cohort)
}

coef = 'DxSCZ'
figList = lapply(df_meta[[coef]]$Symbol[1:10], function(geneSymbol){
  fig1 = plot_forrest( c(tab.de.lst[[coef]], list("Meta-analysis" = df_meta[[coef]])), geneSymbol)

  fig2 = plot_gene(residuals.lst, df_meta[[coef]], geneSymbol, "Dx") + xlab("Diagnosis")

  plot_grid(fig1, fig2, nrow=1)
})

plot_grid(plotlist=figList, ncol=1)
```

### Concordance with other datasets
```{r CMC.concordance}
# CMC Analysis 1
CMC1 <- readr::read_tsv(synGet("syn5607603")$path) %>% 
  dplyr::rename(gene_id = genes,
                `CMC 1 t-statistic` = t) %>% 
  dplyr::select(gene_id, `CMC 1 t-statistic`)
# HBCC-DLPFC microarray
CMC_HBCC <- readr::read_tsv(synGet("syn4941553")$path) %>% 
  dplyr::rename(gene_id = genes,
                `HBCC microarray t-statistic` = t) %>% 
  dplyr::select(gene_id, `HBCC microarray t-statistic`)
# Gandal, Science 2018, Table S1
# Use Tab:DGE
# wget https://science.sciencemag.org/highwire/filestream/720110/field_highwire_adjunct_files/1/aat8127_Table_S1.xlsx
gandal <- readxl::read_xlsx( synGet("syn22255217")$path, sheet = "DGE") %>% 
  dplyr::rename(gene_id = ensembl_gene_id,
                `Gandal, SCZ t-statistic` = `SCZ.t.value`,
                `Gandal, BD t-statistic` = `BD.t.value`,
                `Gandal, ASD t-statistic` = `ASD.t.value`) %>%
  dplyr::select(gene_id, `Gandal, SCZ t-statistic`,
                `Gandal, BD t-statistic`, `Gandal, ASD t-statistic`)
# AMP-AD RNASeq reprocessing GRCh37 - https://www.synapse.org/#!Synapse:syn9702085
MSSM <- readr::read_tsv(synGet("syn10526259")$path) %>%
  dplyr::filter(Comparison %in% "AD-CONTROL",
                Model %in% "Diagnosis") %>%
  dplyr::select(ensembl_gene_id, Tissue, t) %>%
  dplyr::group_by(ensembl_gene_id, Tissue) %>% 
  dplyr::filter(row_number() == 1) %>% 
  tidyr::pivot_wider(names_from = Tissue,
                     values_from = t) %>%
  dplyr::rename(`AMP-AD MSSM FP AD t-statistic` = FP,
                `AMP-AD MSSM IGF AD t-statistic` = IFG,
                `AMP-AD MSSM PHG AD t-statistic` = PHG,
                `AMP-AD MSSM STG AD t-statistic` = STG,
                gene_id = ensembl_gene_id) %>% 
  dplyr::select(gene_id, `AMP-AD MSSM STG AD t-statistic`) 
MSSM <- MSSM[!is.na(MSSM$`AMP-AD MSSM STG AD t-statistic`),]

MAYO <- readr::read_tsv(synGet("syn8468023")$path) %>% 
  dplyr::filter(Comparison %in% "AD-CONTROL",
                Model %in% "Diagnosis") %>%
  dplyr::select(ensembl_gene_id, `Tissue.ref`, t) %>%
  dplyr::group_by(ensembl_gene_id, `Tissue.ref`) %>% 
  dplyr::filter(row_number() == 1) %>% 
  tidyr::pivot_wider(names_from = `Tissue.ref`,
                     values_from = t) %>% 
  dplyr::rename(`AMP-AD MAYO CBE AD t-statistic` = CBE,
                `AMP-AD MAYO TCX AD t-statistic` = TCX,
                gene_id = ensembl_gene_id) %>% 
  dplyr::select(-`AMP-AD MAYO TCX AD t-statistic`)

ROSMAP <- readr::read_tsv(synGet("syn8456721")$path) %>%
    dplyr::filter(Comparison %in% "AD-CONTROL",
                Model %in% "Diagnosis") %>%
  dplyr::rename(gene_id = ensembl_gene_id, 
                `AMP-AD ROSMAP DLPFC AD t-statistic` = t) %>% 
  dplyr::select(gene_id, `AMP-AD ROSMAP DLPFC AD t-statistic`)

out_group <- inner_join(MSSM, MAYO) %>% 
  inner_join(., ROSMAP) %>% 
  inner_join(., gandal)

# merge with meta-analysis

coef = 'DxSCZ'
df = df_meta[[coef]]
df_two = data.frame(gene_id = trim_ensembl_ids(rownames(df)),
                `CMC - meta-analysis` = with(df, logFC/se), check.names=FALSE)

df = tab.de.lst[[coef]][[1]]
df.tmp = data.frame(gene_id = trim_ensembl_ids(rownames(df)),
                `t` = df$t, check.names=FALSE)
colnames(df.tmp)[2] = names(tab.de.lst[[coef]])[1]
df_two = df_two %>% inner_join(., df.tmp)

df = tab.de.lst[[coef]][[2]]
df.tmp = data.frame(gene_id = trim_ensembl_ids(rownames(df)),
                `t` = df$t, check.names=FALSE)
colnames(df.tmp)[2] = names(tab.de.lst[[coef]])[2]
df_two = df_two %>% inner_join(., df.tmp)

# Remove transcript Ids
df_two <- df_two %>% 
  tidyr::separate(gene_id, c("gene_id", "transcript")) %>% 
  dplyr::select(-one_of("transcript"))
df_two <- inner_join(df_two, CMC1, by = c("gene_id")) 
df_two <- inner_join(df_two, CMC_HBCC, by = c("gene_id"))
df_two <- inner_join(df_two, out_group, by = c("gene_id")) %>% 
  group_by(gene_id) %>% 
  dplyr::filter(row_number()==1) %>% 
  tibble::column_to_rownames(var = "gene_id")

colnames(df_two) <- gsub(" t-statistic", "", colnames(df_two))

col2 <- colorRampPalette(rev(c("#67001F", "#B2182B", "#D6604D", "#F4A582",
                           "#FDDBC7", "#FFFFFF", "#D1E5F0", "#92C5DE",
                           "#4393C3", "#2166AC", "#053061")))

cor <- cor(df_two)
# clustering based on distance
hcl = hclust(as.dist(1-cor), method='ward.D2')

results <- pvclust(cor, method.hclust='ward.D2')
# reorder based on clustering
reorder = cor[results$hclust$order, results$hclust$order]

# plot clustering
plot(results, hang=-1)
# plot(hcl, hang=-1)
corrplot(reorder, col = col2(200))
```

```{r test.multtest, eval=FALSE}

library(irlba)
library(pinnacle)
library(poolR)
library(matrixStats)

i = with(res_meta_zenith, (Resource == 'GO') & (Coef == "DxSCZ"))
df = res_meta_zenith[i,]
df$qvalue = qvalue(df$PValue)$qvalue
# df$p.adjust = p.adjust(df$PValue, method="bon" )
df$p.adjust = pmin(1, 2805 * df$PValue)

M = recodeToSparseMatrix(gs.GO)
M = M[rownames(M) %in% trim_ensembl_ids(rownames(df_meta_SCZ)),]
keep = (colSums(M) >= 10) & (colnames(M) %in% rownames(df) )
M = M[,keep]

m = colMeans(M)
s = apply(M, 2, var) * nrow(M) / (nrow(M) + 1)

dcmp = irlba( M, nv = 500, center=m, scale=sqrt(s), right_only=FALSE)
ev = dcmp$d^2 / nrow(M)

plot(cumsum(ev) / ncol(M))

df_unsort = df[colnames(M),]
edcmp = list(vectors = dcmp$v, values = dcmp$d^2 / nrow(M))

res = p.corr.adjust( df_unsort$PValue, dcmp = edcmp, method.sim="eigen", B = nrow(df_unsort)*4)

plot( cumsum(dcmp$d^2 / sum(dcmp$d^2)))

C = cor(as.matrix(M[,1:1000]))
evals = eigen(C)$values

dcmp = svd(scale(M) / sqrt(nrow(M)))
dcmp$d^2

plot(cumsum(dcmp$d^2 / sum(dcmp$d^2)))

meff(C, method='gao')

dcmp = svd(scale(M[,1:100]))

meff(eigen = dcmp$d^2 , method = "galwey")

```

```{r show.gene_set_test.marginal, eval=TRUE}
show_zenith_results = function( res, coef, Resource, FDR = 0.05){

  i = which((res$Resource == Resource) & (res$Coef == coef))
  df = res[i,]

	# filter by FDR
	df = df[df$FDR <= FDR,]
  df$Coef = c()
  df$se = c()
  df$Resource = c()

	df %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption=paste(Resource, coef), row.names=FALSE) %>% kable_styling(full_width=FALSE)
}

show_zenith_results( res_meta_zenith, "DxSCZ", "GO", FDR=0.05)
show_zenith_results( res_meta_zenith, "DxSCZ", "MSigDB", FDR=0.05)
```

# Sex Analysis
```{r test.Sex, fig.height=5, fig.width=10}
# Schizophrenia testing
# extract from each chort
coef = 'Reported_GenderMale'
figList.Sex = lapply( names(df_DE_results), function(key){
  df = df_DE_results[[key]][[coef]]
  main = paste(key, coef, sep = ': ' )
  plotVolcano( df, df$Symbol[1:20] ) + ggtitle( main )
})

# volcano plots
plot_grid(plotlist=figList.Sex, nrow=1)
```

### Autosomes
```{r test.Sex.autosome, fig.height=5, fig.width=10}
# Schizophrenia testing
# extract from each chort
coef = 'Reported_GenderMale'
figList.Sex = lapply( names(df_DE_results), function(key){
  df = df_DE_results[[key]][[coef]]
  main = paste(key, coef, sep = ': ' )
  df = df[df$Chrom %in% 1:22,]
  plotVolcano( df, df$Symbol[1:20] ) + ggtitle( main )
})

# volcano plots
plot_grid(plotlist=figList.Sex, nrow=1)
```

## Meta-analysis across cohorts
```{r concordance.Sex, fig.height=6, fig.width=10}
# Concordance analysis
resConcord.Sex = plot_concordance(tab.de.lst[[coef]], showGenes = df_meta[[coef]]$Symbol[1:20])

main = "Concordance of Sex"
plot_grid(  resConcord.Sex$fig_logFC + ggtitle(main),
            resConcord.Sex$fig_tstat + ggtitle(main), nrow=1)

data.frame(Cohort = names(tab.de.lst[[coef]]),
          "# DE"  = resConcord.Sex$n.de,
          pi1     = resConcord.Sex$pi_single,
          check.names=FALSE) %>% 
          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)

resConcord.Sex$tab_pi  %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
resConcord.Sex$tab_corr  %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
```

```{r concordance.Sex.autosome, fig.height=6, fig.width=10}

tab.de.lst.autosome = lapply(tab.de.lst[[coef]], function(x) x[x$Chrom %in% 1:22,])

# Concordance analysis
resConcord.Sex = plot_concordance(tab.de.lst.autosome, showGenes = tab.de.lst.autosome$Symbol[1:20])

main = "Concordance of Sex - autosomes"
plot_grid(  resConcord.Sex$fig_logFC + ggtitle(main),
            resConcord.Sex$fig_tstat + ggtitle(main), nrow=1)

data.frame(Cohort = names(resConcord.Sex$pi_single),
          "# DE"  = resConcord.Sex$n.de,
          pi1     = resConcord.Sex$pi_single,
          check.names=FALSE) %>% 
          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)

resConcord.Sex$tab_pi %>% mutate_if(is.numeric, format, digits=3)  %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
resConcord.Sex$tab_corr  %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
```

```{r plot.meta.analysis.Sex, fig.height=5, fig.width=10}
# volcano plot for meta-analysis
main = 'Meta-analysis: Sex'
fig1 = plotVolcano( df_meta[[coef]], df_meta[[coef]]$Symbol[1:20] ) + ggtitle( main )

df = df_meta[[coef]][df_meta[[coef]]$Chrom %in% 1:22,]
fig2 = plotVolcano( df, df$Symbol[1:20] ) + ggtitle( paste(main, "autosome") )
plot_grid(fig1, fig2, ncol=2)

data.frame(pi1 = 1 - qvalue(df_meta[[coef]]$P)$pi0,
          `# DE` = sum(df_meta[[coef]]$adj.P.Val < 0.05),
          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for Sex meta-analysis") %>% kable_styling(full_width=FALSE)
```

## Sex circos
```{r sex.circos}
coef = 'Reported_GenderMale'
sapply( names(df_DE_results), function(key){
  plot_circos_logFC( df_DE_results[[key]][[coef]] ) + ggtitle(key)
  })

plot_circos_logFC( df_meta[[coef]] ) + ggtitle("Meta analysis")
```

## Plot top examples
```{r plot.forrest.dlpfc.Sex, fig.width=8, fig.height=20}  
figList = lapply(df_meta[[coef]]$Symbol[1:10], function(geneSymbol){

  fig1 = plot_forrest( c(tab.de.lst[[coef]], list("Meta-analysis" = df_meta[[coef]])), geneSymbol)

  fig2 = plot_gene( residuals.lst, df_meta[[coef]], geneSymbol, "Reported_Gender", fill=c("red", "blue")) + xlab("Sex")

  plot_grid(fig1, fig2, nrow=1)
})

plot_grid(plotlist=figList, ncol=1)
```

## Meta-analysis of zenith results
```{r show.gene_set_test.marginal.Sex, eval=TRUE}
show_zenith_results( res_meta_zenith, "Reported_GenderMale", "GO", FDR=0.05)         
show_zenith_results( res_meta_zenith, "Reported_GenderMale", "MSigDB", FDR=0.05)        
```

### Only autosomes
```{r show.gene_set_test.marginal.Sex.autosome, eval=TRUE}
show_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "GO", FDR=0.05) 
show_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "MSigDB", FDR=0.05)   
```

# Age Analysis
```{r test.Age, fig.height=5, fig.width=10}
# Schizophrenia testing
# extract from each chort
coef = 'ageOfDeath'
figList.Age = lapply( names(df_DE_results), function(key){
  df = df_DE_results[[key]][[coef]]
  main = paste(key, coef, sep = ': ' )
  plotVolcano( df, df$Symbol[1:20] ) + ggtitle( main )
})

# volcano plots
plot_grid(plotlist=figList.Age, nrow=1)
```

## Meta-analysis across cohorts
```{r concordance.Age, fig.height=6, fig.width=10}
# Concordance analysis
resConcord.Age = plot_concordance(tab.de.lst[[coef]], showGenes = df_meta[[coef]]$Symbol[1:20])

main = "Concordance of Age"
plot_grid(  resConcord.Age$fig_logFC + ggtitle(main),
            resConcord.Age$fig_tstat + ggtitle(main), nrow=1)

data.frame(Cohort = names(tab.de.lst[[coef]]),
          "# DE"  = resConcord.Age$n.de,
          pi1     = resConcord.Age$pi_single,
          check.names=FALSE) %>% 
          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)

resConcord.Age$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
resConcord.Age$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
```

```{r plot.meta.analysis.Age, fig.height=5, fig.width=10}
# volcano plot for meta-analysis
main = 'Meta-analysis: Age'
plotVolcano( df_meta[[coef]], df_meta[[coef]]$Symbol[1:20] ) + ggtitle( main )

data.frame(pi1 = 1 - qvalue(df_meta[[coef]]$P)$pi0,
          `# DE` = sum(df_meta[[coef]]$adj.P.Val < 0.05),
          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for Age meta-analysis") %>% kable_styling(full_width=FALSE)
```

## Plot top examples
```{r plot.forrest.dlpfc.Age, fig.width=8, fig.height=20}  
figList = lapply(df_meta[[coef]]$Symbol[1:10], function(geneSymbol){

  fig1 = plot_forrest( c(tab.de.lst[[coef]], list("Meta-analysis" = df_meta[[coef]])), 
    geneSymbol)

  expr.list = lapply(vobj.lst, function(x)x$E)

  fig2 = plot_gene( expr.list, df_meta[[coef]], geneSymbol, "ageOfDeath", fill=c("red", "blue")) + xlab("Age")

  plot_grid(fig1, fig2, nrow=1)
})

plot_grid(plotlist=figList, ncol=1)
```

## Meta-analysis of zenith results
```{r show.gene_set_test.marginal.Age, eval=TRUE}
show_zenith_results( res_meta_zenith, "ageOfDeath", "GO", FDR=0.01)   
show_zenith_results( res_meta_zenith, "ageOfDeath", "MSigDB", FDR=0.01)    
```

# Sex by Disease interaction
```{r dream.sex.disease}
form.lst = list(
  `MSSM-Penn-Pitt`  = ~ Dx*Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
  `NIMH-HBCC`       = ~ Dx*Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)

fit_inter_sex.disease.lst = lapply( vobj.lst, function(vobj){

  i = match(colnames(vobj), rownames(METADATA))
  info = METADATA[i,]

  form_use = form.lst[[info$Cohort[1]]]

  fit = dream(vobj, form_use, info, BPPARAM = SnowParam(18))  
  eBayes( fit )
  })
names(fit_inter_sex.disease.lst) = names(vobj.lst)

coefTest = c('DxSCZ:Reported_GenderMale')
```

```{r run.zenith.inter_sex.disease, eval=TRUE}
res_zenith.inter_sex.disease = lapply( names(fit_inter_sex.disease.lst), function(key){

  fit = fit_inter_sex.disease.lst[[key]]

  res.GO = run_zenith(  fit, 
                        coefTest,
                        gs.GO.collection)
  res.GO$Resource = "GO"   
  res.GO$Cohort = key

  res.MSigDB = run_zenith(fit, 
                          coefTest,
                          gs.MSigDB.collection) 
  res.MSigDB$Resource = "MSigDB"
  res.MSigDB$Cohort = key

  rbind(res.GO, res.MSigDB)
})
names(res_zenith.inter_sex.disease) = names(fit_inter_sex.disease.lst)

# only autosomes
res_zenith.inter_sex.disease_autosome = lapply( names(fit_inter_sex.disease.lst), function(key){

  fit = fit_inter_sex.disease.lst[[key]]

  # only keep autosomes, for Reported_GenderMale
  tab = coef(fit) %>% getGeneSymbol()
  geneids = rownames(tab)[tab$Chrom %in% 1:22]
  fit = fit[geneids,]

  res.GO = run_zenith(  fit, 
                        coefTest,
                        gs.GO.collection)
  res.GO$Resource = "GO"   
  res.GO$Cohort = key

  res.MSigDB = run_zenith(fit, 
                          coefTest,
                          gs.MSigDB.collection) 
  res.MSigDB$Resource = "MSigDB"
  res.MSigDB$Cohort = key

  rbind(res.GO, res.MSigDB)
})
names(res_zenith.inter_sex.disease_autosome) = names(fit_inter_sex.disease.lst)
```

## Meta-analysis of zenith results
```{r zenith.meta.inter_sex.disease, eval=TRUE}
res_meta_zenith.inter_sex.disease = zenith_meta_analysis( res_zenith.inter_sex.disease )  
res_meta_zenith.inter_sex.disease_autosome = zenith_meta_analysis( res_zenith.inter_sex.disease_autosome )
```

```{r get.signatures.inter_sex.disease, fig.width=10, fig.height=5}
# Get DE signatures
df_DE_results.inter_sex.disease = lapply(fit_inter_sex.disease.lst, function(fit){
  res = lapply(coefTest, function(coef){
    res = topTable(fit, coef, number=Inf) 
    res$Coef = coef
    res %>% getGeneSymbol()  
    })
  names(res) = coefTest
  res
  })
names(df_DE_results.inter_sex.disease) = names(fit_inter_sex.disease.lst)
```

```{r test.scz.inter_sex.disease, fig.height=5, fig.width=10}
coef = 'DxSCZ:Reported_GenderMale'

figList.Scz = lapply( names(df_DE_results), function(key){
  df = df_DE_results.inter_sex.disease[[key]][[coef]]

  main = paste(key, coef, sep = ': ' )
  plotVolcano( df, df$Symbol[1:10] ) + ggtitle( main )
})

# volcano plots
plot_grid(plotlist=figList.Scz, nrow=1)
```

## Meta-analysis across cohorts
```{r meta.analysis.DE.inter_sex.disease}
# Store results for each coef and cohort
tab.de.inter_sex.disease.lst = lapply(coefTest, function(coef){    
  tab.de.inter_sex.disease.lst = lapply( df_DE_results.inter_sex.disease, function(lst){
    lst[[coef]]
  })
  names(tab.de.inter_sex.disease.lst) = names(df_DE_results.inter_sex.disease)
  tab.de.inter_sex.disease.lst
  })
names(tab.de.inter_sex.disease.lst) = coefTest

# Create list preparing for meta-analysis for SCZ
df_meta.inter_sex.disease = mclapply(coefTest, function(coef){
  # Run meta-analysis
  run_meta_analysis( tab.de.inter_sex.disease.lst[[coef]], "FE") %>% getGeneSymbol
}, mc.cores=6) 
names(df_meta.inter_sex.disease) = coefTest
```

```{r concordance.scz.inter_sex.disease, fig.height=6, fig.width=10}
# Concordance analysis.  
resConcord.inter_sex.disease = plot_concordance(tab.de.inter_sex.disease.lst[[coef]], showGenes = df_meta.inter_sex.disease[[coef]]$Symbol[1:20])

main = "Concordance of DxSCZ:Reported_GenderMale"
plot_grid(  resConcord.inter_sex.disease$fig_logFC + ggtitle(main),
            resConcord.inter_sex.disease$fig_tstat + ggtitle(main), nrow=1)

data.frame(Cohort = names(tab.de.inter_sex.disease.lst[[coef]]),
          "# DE"  = resConcord.inter_sex.disease$n.de,
          pi1     = resConcord.inter_sex.disease$pi_single,
          check.names=FALSE) %>% 
          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)

resConcord.inter_sex.disease$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
resConcord.inter_sex.disease$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
```

```{r plot.meta.analysis.SCZ.inter_sex.disease, fig.height=5, fig.width=10}
# volcano plot for meta-analysis.  
main = paste('Meta-analysis', coef, sep = ': ' )
plotVolcano( df_meta.inter_sex.disease[[coef]], df_meta.inter_sex.disease[[coef]]$Symbol[1:20] ) + ggtitle( main )

data.frame(pi1 = 1 - qvalue(df_meta.inter_sex.disease[[coef]]$P)$pi0,
          `# DE` = sum(df_meta.inter_sex.disease[[coef]]$adj.P.Val < 0.05),
          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for SCZ:Sex meta-analysis") %>% kable_styling(full_width=FALSE)
```

## Plot top examples
```{r plot.forrest.dlpfc.inter_sex.disease, fig.width=8, fig.height=40}

figList = lapply(df_meta.inter_sex.disease[[coef]]$Symbol[1:20], function(geneSymbol){
  fig1 = plot_forrest( c(tab.de.inter_sex.disease.lst[[coef]], list("Meta-analysis" = df_meta.inter_sex.disease[[coef]])), geneSymbol, plot2='p')

  fig2 = plot_gene(residuals.lst, df_meta.inter_sex.disease[[coef]], geneSymbol, "Dx") + xlab("Diagnosis") + facet_wrap(~Reported_Gender + Cohort, nrow=1)

  plot_grid(fig1, fig2, nrow=1)
})

plot_grid(plotlist=figList, ncol=1)
```

## Meta-analysis of zenith results
```{r show.gene_set_test.marginal.inter_sex.disease, eval=TRUE}
coef = 'DxSCZ:Reported_GenderMale'   
show_zenith_results( res_meta_zenith.inter_sex.disease, coef, "GO", FDR=0.1)  
show_zenith_results( res_meta_zenith.inter_sex.disease, coef, "MSigDB", FDR=0.1)

show_zenith_results( res_meta_zenith.inter_sex.disease_autosome, coef, "GO", FDR=0.1) 
show_zenith_results( res_meta_zenith.inter_sex.disease_autosome, coef, "MSigDB", FDR=0.1)    
```

```{r exit2, cache=FALSE}

# DIFFERENTIAL CORRLEATION NETWORKS
# CLUSTER on C_Scz - C_Control
# Like contrastive PCA

#   rmarkdown::render("combined_analysis/combine_all_DLPFC.Rmd")

knitr::knit_exit()
```