### 2020-09-19 combined_analysis/diff.networks.Rmd
Update diff networks

---
title: "CommonMind Consortium: DLPFC"
subtitle: "Analysis of read count data by cohort"
author:
- name: "[Gabriel Hoffman](http://gabrielhoffman.github.io)$^{1}$ and Kelsey Montgomery$^{2}$"
  affiliation: | 
    [1] Icahn School of Medicine at Mount Sinai, New York
    [2] Sage Bionetworks, Seattle, WA
date: "Run on `r Sys.time()`"
documentclass: article
output:
  html_document:
    toc: true
    toc_float: true
---

<!---

cd /hpc/users/hoffmg01/work/covarr-de
# rm -rf combined_analysis/diff.networks_* combined_analysis/diff.networks.html
ml git pandoc udunits proj gdal geos
git pull
R

system("git pull")

rmarkdown::render("combined_analysis/diff.networks.Rmd")

--->

Since we are generating networks across cohorts and sexes, identifiying conserved clusters using standard methods becomes non-trivial.  Instead of defining clusters for each subset and evaluating the difference, we first compute the differences in the correlation networks, then identify a single set of clusters based on this difference-network.  Then use hypothesis test of differential correlation in the second cohort to get p-value.

But if concensus models are constructed to be differential by definition, then these clusters could show differential correlation even under the null.  Therefore, use set A as discovery for M/F differential network, and replicated in set B.

```{r setup, include=FALSE}
suppressPackageStartupMessages({
library(synapser)

synLogin()

library(WGCNA)
library(variancePartition)
library(decorate)
library(ggrepel)
library(cowplot) 
library(zenith) 
library(knitr) 
library(plyr) 
library(qvalue)  
library(BiocParallel)  
}) 
source("../common_functions/common_functions.R")   
# source("common_functions/common_functions.R")

knitr::opts_chunk$set(
  echo = FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,  
  tidy = FALSE,
  cache = TRUE,
  cache.lazy = FALSE)
```

```{r load.data}
# Load data
resid.add = readRDS(synGet('syn22416489')$path)
vobj.lst = readRDS(synGet('syn22416446')$path)
METADATA = readRDS(synGet('syn22416411')$path)
METADATA$Reported_Gender = factor(METADATA$Reported_Gender, c("Male", "Female"))

# identify genes to keep: remove chrY genes
geneEnsID = lapply(resid.add, function(x){
	lapply(x, rownames)
	})
geneEnsID = unique(unlist(geneEnsID))

df_genes = data.frame(ENSEMBL = geneEnsID, stringsAsFactors=FALSE) %>% getGeneSymbol('ENSEMBL')
# table(df_genes$Chrom)

keep.ensID = df_genes$ENSEMBL[df_genes$Chrom %in% c(1:22, 'X', 'MT')]

minClusterSize = 20
```

```{r load.go}
# Load Gene Ontology database 
go.gs = get_GeneOntology()
```

```{r sex.confounding, fig.width=5, fig.height=5}
form = ~ Reported_Gender + GABA + GLU + Olig + MgAs + ageOfDeath 
par(mfrow=c(2,2))
res = lapply( unique(METADATA$Institution), function(INST){
	i = METADATA$Institution == INST
	C = canCorPairs(form, METADATA[i,]) 
	plotCorrMatrix(C, main=INST)
})
```

```{r load.dream}
fit_marginal.lst = readRDS(synGet('syn22416500')$path)
fit_inter_sex.disease.lst = readRDS(synGet('syn22416524')$path)
```

# Null network
```{r diff.networks.null}

# Compute difference networks
##############################

resVersion = "Sex"

METADATA$rnd = sample(c('A','B'), nrow(METADATA), replace=TRUE)
METADATA$rnd = factor(METADATA$rnd, c('A','B'))

variable = 'rnd'

diff.net = get_difference_network(resid.add[[resVersion]], METADATA, variable)

# Discover clusters in first cohort
#----------------------------------
C.diff.discovery = diff.net[[1]][keep.ensID,keep.ensID]

rm(diff.net) # saving this to disk is slow, so just delete it

hcl = hclust( as.dist(C.diff.discovery) )

dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize)

dynamicColors = labels2colors(dynamicMods, colorSeq = colors())
tab = table(dynamicColors)
hist(tab, xlab="Cluster size")
```

```{r testing.null}
# Hypothesis testing         
df_test = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors, METADATA, useSLED=FALSE)   #, BPPARAM=SnowParam(18)) 
```

```{r null.results, fig.width=5}
p = df_test$P.Value[df_test$Cohort=="MSSM-Penn-Pitt"]
# print(1 - qvalue(p)$pi0)
hist(p, main="P-values under the null in the discovery cohort", xlab="p-value")

p = df_test$P.Value[df_test$Cohort== "NIMH-HBCC"]
data.frame("# DE" = sum(p.adjust(p, 'fdr') < 0.05), check.names=FALSE) %>% kable
hist(p, main="P-values under the null in the replication cohort", xlab="p-value")
```

# Sex network
## Difference networks
```{r diff.networks}

# Compute difference networks
##############################

resVersion = "Sex"
variable = 'Reported_Gender'

diff.net.Sex = get_difference_network(resid.add[[resVersion]], METADATA, variable)

# Discover clusters in first cohort
#----------------------------------
C.diff.discovery = diff.net.Sex[[1]][keep.ensID,keep.ensID]

rm(diff.net.Sex) # saving this to disk is slow, so just delete it

hcl = hclust( as.dist(C.diff.discovery) )

dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize)

dynamicColors.Sex = labels2colors(dynamicMods, colorSeq = colors())
tab = table(dynamicColors.Sex)
hist(tab, xlab="Cluster size")
```

```{r testing}
# Hypothesis testing         
df_test.Sex = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Sex, METADATA)#, BPPARAM=SnowParam(18)) 
```

## Discovery cohort
```{r plot.discovery}
# plot discovery and testing.      
ggplot(df_test.Sex, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test.Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed") 
```

```{r modules.in.discovery, fig.width=8, fig.height=40}

figList = lapply(df_test.Sex$Module[1:5], function(mod){

	plot_module( mod, df_test.Sex, METADATA, dynamicColors.Sex, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt", 5)   
	})

plot_grid( plotlist = figList, ncol=1)
```

## Replication cohort
```{r second.cohort, fig.width=8, fig.height=12}
df_cohort2 = df_test.Sex[df_test.Sex$Cohort == "NIMH-HBCC",]
df_cohort2$Module = factor(df_cohort2$Module, unique(df_cohort2$Module))
table(df_cohort2$FDR < 0.05)

df = df_cohort2[df_cohort2$P.Value < 0.1,]

ggplot(df, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
```

```{r modules.in.second.cohort, fig.width=8, fig.height=25}

i = which(df_cohort2$FDR < 0.05)

figList = lapply(df_cohort2$Module[i], function(mod){

	plot_module( mod, df_test.Sex, METADATA, dynamicColors.Sex, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC")   
	})

plot_grid( plotlist = figList, ncol=1)
```

## Module enrichment
```{r top.enrich.Sex}
df_module.Sex = data.frame(	ENSEMBL = rownames(C.diff.discovery) %>% trim_ensembl_ids, 
						Module = dynamicColors.Sex)

df_module_enrich.Sex = enrich.test( df_module.Sex, go.gs, df_cohort2$Module[i])
```

```{r plot.enrich.Sex, fig.width=9}
figList = lapply( df_cohort2$Module[i], function(module){
	plot_enrich(df_module_enrich.Sex, module, 8) + ggtitle(module)
	}) 
plot_grid(plotlist = figList, ncol=1)
```

# Disease network
## Difference networks
```{r diff.networks.Dx}

# Compute difference networks
##############################

resVersion = "Dx"
variable = 'Dx'

diff.net.Dx = get_difference_network(resid.add[[resVersion]], METADATA, variable)

# Discover clusters in first cohort
#----------------------------------
C.diff.discovery = diff.net.Dx[[1]][keep.ensID,keep.ensID]

rm(diff.net.Dx) # saving this to disk is slow, so just delete it

hcl = hclust( as.dist(C.diff.discovery) )

dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize)

dynamicColors.Dx = labels2colors(dynamicMods, colorSeq = colors())
tab = table(dynamicColors)
hist(tab, xlab="Cluster size")
```

```{r testing.Dx}
# Hypothesis testing     
df_test.Dx = test_differential_correlation(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Dx, METADATA)  
```

## Discovery cohort
```{r plot.discovery.Dx}
# plot discovery and testing.  
ggplot(df_test.Dx, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test.Dx$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
```

```{r modules.in.discovery.Dx, fig.width=10, fig.height=50}

figList = lapply(df_test$Module[1:5], function(mod){

	plot_module( mod, df_test.Dx, METADATA, dynamicColors, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt")   
	})

plot_grid( plotlist = figList, ncol=1)
```

## Replication cohort
```{r second.cohort.Dx, fig.width=6, fig.height=12}
df_cohort2 = df_test.Dx[df_test.Dx$Cohort == "NIMH-HBCC",]
df_cohort2$Module = factor(df_cohort2$Module, unique(df_cohort2$Module))
table(df_cohort2$FDR < 0.05)

df = df_cohort2[df_cohort2$P.Value < 0.1,]

ggplot(df, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
```

```{r modules.in.second.cohort.Dx, fig.width=8, fig.height=100}

i = which(df_cohort2$FDR < 0.05)[1:10]

figList = lapply(df_cohort2$Module[i], function(mod){

	plot_module( mod, df_test.Dx, METADATA, dynamicColors.Dx, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC")   
	})

plot_grid( plotlist = figList, ncol=1)
```

## Module enrichment
```{r top.enrich.Dx}
df_module.Dx = data.frame(	ENSEMBL = rownames(C.diff.discovery) %>% trim_ensembl_ids, 
						Module = dynamicColors.Dx)

df_module_enrich.Dx = enrich.test( df_module.Dx, go.gs, df_cohort2$Module[i])
```

```{r plot.enrich.Dx, fig.width=9}
figList = lapply( df_cohort2$Module[i], function(module){
	plot_enrich(df_module_enrich.Dx, module, 8) + ggtitle(module)
	}) 
plot_grid(plotlist = figList, ncol=1)
```

# Interaction analysis

## Difference networks
```{r diff.networks.Dx_Sex}

# Compute difference networks
##############################

METADATA$Dx_Sex = with(METADATA, paste(Reported_Gender, Dx, sep='_'))
METADATA$Dx_Sex = factor(METADATA$Dx_Sex, c('Female_Control', 'Female_SCZ', 'Male_Control', 'Male_SCZ'))

resVersion = "Dx_Sex"
variable = 'Dx_Sex'

net.Dx_Sex = get_networks(resid.add[[resVersion]], METADATA, variable)

lvls = levels(METADATA$Dx_Sex)

print(paste0('(',lvls[4], ' - ',lvls[3], ') - (',lvls[2], ' - ',lvls[1], ')'))

# Discover clusters in first cohort
#----------------------------------
C = net.Dx_Sex[[1]]
C.diff.discovery = (C[[lvls[4]]] - C[[lvls[3]]]) - (C[[lvls[2]]] - C[[lvls[1]]])

C.diff.discovery = C.diff.discovery[keep.ensID,keep.ensID]

rm(net.Dx_Sex) # saving this to disk is slow, so just delete it

hcl = hclust( as.dist(C.diff.discovery) )

dynamicMods = cutreeDynamic(hcl, distM=C.diff.discovery, minClusterSize = minClusterSize)

dynamicColors.Dx_Sex = labels2colors(dynamicMods, colorSeq = colors())
tab = table(dynamicColors.Dx_Sex)
hist(tab, xlab="Cluster size")
```

```{r testing.Dx_Sex}
# Hypothesis testing     
df_test.Dx_Sex = test_differential_correlation_interaction(resid.add[[resVersion]], C.diff.discovery, dynamicColors.Dx_Sex, METADATA, nperm=1000)    
```

## Discovery cohort
```{r plot.discovery.Dx_Sex}
# plot discovery and testing.  
ggplot(df_test.Dx_Sex, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df_test.Dx_Sex$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
```

```{r modules.in.discovery.Dx_Sex, fig.width=10, fig.height=50}

figList = lapply(df_test$Module[1:5], function(mod){.

	plot_module( mod, df_test.Dx_Sex, METADATA, dynamicColors.Dx_Sex, C.diff.discovery, resid.add[[resVersion]], variable, "MSSM-Penn-Pitt")   
	})

plot_grid( plotlist = figList, ncol=1)
```

## Replication cohort
```{r second.cohort.Dx_Sex, fig.width=6, fig.height=12}
df_cohort2 = df_test.Dx_Sex[df_test.Dx_Sex$Cohort == "NIMH-HBCC",]
df_cohort2$Module = factor(df_cohort2$Module, unique(df_cohort2$Module))
table(df_cohort2$FDR < 0.05)

df = df_cohort2[df_cohort2$P.Value < 0.1,]

ggplot(df, aes(Module, -log10(FDR))) + geom_bar(stat='identity', fill=df$Module) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) + coord_flip() + facet_wrap(~Cohort) + ylab(bquote(-log[10]~FDR)) + scale_y_continuous(expand=c(0,0)) + geom_hline(yintercept = -log10(0.05), color="red", linetype="dashed")
```

```{r modules.in.second.cohort.Dx_Sex, fig.width=8, fig.height=80}

i = which(df_cohort2$FDR < 0.1)

figList = lapply(df_cohort2$Module[i], function(mod){

	fig1 = plot_module( mod, df_test.Dx_Sex, METADATA, dynamicColors.Dx_Sex, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC", lvlidx=1:2)

	fig2 = plot_module( mod, df_test.Dx_Sex, METADATA, dynamicColors.Dx_Sex, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC", lvlidx=3:4)

	fig3 = plot_module( mod, df_test.Dx_Sex, METADATA, dynamicColors.Dx_Sex, C.diff.discovery, resid.add[[resVersion]], variable, "NIMH-HBCC", lvlidx=1:4)

	plot_grid(fig1, fig2, fig3, ncol=1)
	})

plot_grid( plotlist = figList, ncol=1)
```

## Module enrichment
```{r top.enrich.Dx_Sex}
df_module.Dx_Sex = data.frame(	ENSEMBL = rownames(C.diff.discovery) %>% trim_ensembl_ids, 
						Module = dynamicColors.Dx_Sex)

df_module_enrich.Dx_Sex = enrich.test( df_module.Dx_Sex, go.gs, df_cohort2$Module[i])
```

```{r plot.enrich.Dx_Sex, fig.width=9}
figList = lapply( df_cohort2$Module[i], function(module){
	plot_enrich(df_module_enrich.Dx_Sex, module, 8) + ggtitle(module)
	}) 
plot_grid(plotlist = figList, ncol=1)
```

```{r exit, cache=FALSE}
knitr::knit_exit()
```

<!---

# need to generate correct residuals

GWAS hits, TWAS, SCHEMA
ChrX and ChrY genes with autism and Scz

DE within each sex, on chrX and chrY.
	drop chrY genes for computing FDR?

Can we predict sex given autosomal genes?

PMID: 26892004
https://science.sciencemag.org/content/369/6509/eaba3066
https://www.nature.com/articles/s41576-018-0083-1

test: how much sex differences are given by age, cell type and institution.

Apparent differences in cell type composition can be explained by age.
	After accounting for age, can cell type predict sex?

library(mvtnorm)
library(Rfast)

X = matrnorm(300, 20)
Y = matrnorm(300, 20)

library(BiocParallel)
res = decorate:::.sLED(X=X, Y=Y, npermute=100000, verbose=TRUE, BPPARAM = SerialParam())

res$pVal

C_x = cor(X)
C_y = cor(Y)

df_cor = data.frame( 	a = C_x[lower.tri(C_x)],
						b = C_y[lower.tri(C_y)])

with(df_cor, plot(a,b))

a = calc_sde(id=1,points=df_cor);

Sigma = cov(df_cor)

X = matrnorm(1000, 30)

C_x = cor(X)

C = C_x[1:5, 1:5]
colnames(C) = paste0('gene_', 1:nrow(C))
rownames(C) = colnames(C)

library(ggraph)
library(ggplot2)
library(tidygraph)
library(Rfast)

load("~/Downloads/C.test.RDATA")

C = C1 - C2

plot_corr_network( C1-C2 )

rstat_nodes <- data.frame(name = c("H

	adley", "David", "Romain", "Julia"))
rstat_edges <- data.frame(from = c(1, 1, 1, 2, 3, 3, 4, 4, 4),
                         to = c(2, 3, 4, 1, 1, 2, 1, 2, 3))
rstat_edges$weight = 1:9

rstat_nodes <- data.frame(name = c("Hadley", "David", "Romain", "Julia"))
rstat_edges <- data.frame(from = c(1, 1, 1, 2, 3, 3, 4, 4, 4),
                         to = c(2, 3, 4, 1, 1, 2, 1, 2, 3))
rstat_edges$weight = rnorm(9)

net = tbl_graph(nodes = rstat_nodes, edges = rstat_edges, directed=FALSE)

ggraph(net, layout = "stress") + 
  geom_node_point(size=10, color="blue") +
  geom_edge_link(aes(width = weight), alpha = 0.8) + 
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_text(aes(label = name), repel = TRUE) +
  labs(edge_width = "Correlation") +
  theme_graph()

net = ggnetwork(C)

set.edge.attribute
set.edge.value

ggplot(net, aes(x = x, y = y, xend = xend, yend = yend)) + theme_blank() + geom_nodes(aes(x,y)) + geom_edges()

ggplot(
ggnetwork(emon[[1]], layout = "kamadakawai", arrow.gap = 0.025),
aes(x, y, xend = xend, yend = yend) )+
geom_edges(aes(color = Frequency, size=Frequency^2),
arrow = arrow(length = unit(10, "pt"), type = "closed")
)+
geom_nodes(aes(size = Formalization)) + scale_color_gradient(low = "grey50", high = "tomato") + scale_size_area(breaks = 1:3) +
theme_blank()

net = network(C)

get.edges(net, "gene_1")

rstat_nodes <- data.frame(name = c("Hadley", "David", "Romain", "Julia"))
rstat_edges <- data.frame(from = c(1, 1, 1, 2, 3, 3, 4, 4, 4),
                         to = c(2, 3, 4, 1, 1, 2, 1, 2, 3))
rstat_edges$weight = 1:9

net = tbl_graph(nodes = rstat_nodes, edges = rstat_edges, directed=FALSE)

ggraph(net, layout = "auto") + 
  geom_node_point(size=10, color="blue") +
  geom_edge_link(aes(width = weight), alpha = 0.8) + 
  scale_edge_width(range = c(0.2, 2)) +
  geom_node_text(aes(label = name), repel = TRUE) +
  labs(edge_width = "Correlation") +
  theme_graph()

--->