### 2020-09-20 combined_analysis/combine_all_DLPFC.Rmd
filer

---
title: "CommonMind Consortium: DLPFC"
subtitle: "Analysis of read count data by cohort"
author:
- name: "[Gabriel Hoffman](http://gabrielhoffman.github.io)$^{1}$ and Kelsey Montgomery$^{2}$"
  affiliation: | 
    [1] Icahn School of Medicine at Mount Sinai, New York
    [2] Sage Bionetworks, Seattle, WA
date: "Run on `r Sys.time()`"
documentclass: article
output:
  html_document:
    toc: true
    toc_float: true
params:
  upload: TRUE
---

<!---

cd /hpc/users/hoffmg01/work/covarr-de
# rm -rf combined_analysis/combine_all_DLPFC_cache
ml git pandoc
git pull
R

system("git pull")

rmarkdown::render("combined_analysis/combine_all_DLPFC.Rmd")

--->

```{r knit2synapse, eval= FALSE}
library(synapser)
library(knit2synapse)
synLogin()
knit2synapse::createAndKnitToFolderEntity(file = "combined_analysis/combine_all_DLPFC.Rmd",
                                          parentId ="syn22416298",
                                          folderName = 'Joint analysis of DLPFC count data',
                                          overwrite=TRUE, 
                                          knitmd=TRUE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}

library(githubr)
parentId = 'syn22416298';
activityName = 'Joint analysis of DLPFC count data';
activityDescription = 'Joint analysis of DLPFC count data';

thisFileName <- 'combine_all_DLPFC.Rmd'

# Github link
thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('combined_analysis/',thisFileName))
```

<!---

TODO

zenith 1 sided

zenith plots
replication or correlation as a function of expression magnitude

difference networks
-------------------

DONE

Effect of ilr should depend on Instaition
suffix 1.
zenith for Sex with autosomes
check zenith squared test statistics

--->

```{r setup, include=FALSE}
suppressPackageStartupMessages({
library(synapser)

synLogin()

library(ggplot2)
library(dplyr)
library(ComplexHeatmap)
library(readxl)
library(gtools)
library(corrplot)
library(pvclust)
library(metafor)
library(data.table)
library(cowplot)
library(ggraph)
library(tidygraph)
library(kableExtra)
library(githubr)
library(ggrepel)
library(corrplot)
library(zenith)
library(compositions)
library(knitr)
library(patchwork)
library(plyr) 
library(BiocParallel)

library(limma)
library(edgeR)
library(variancePartition)

}) 
source("../common_functions/common_functions.R")

knitr::opts_chunk$set(
  echo = FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r read.counts}
# Read raw counts from MPP and HBCC
counts.MPP = fread(synGet('syn21867938')$path, data.table=FALSE)
rownames(counts.MPP) = counts.MPP$Geneid
counts.MPP = counts.MPP[,-c(1:6)]

# HBCC - DLFPC
counts.HBCC = fread(synGet('syn21886235')$path, data.table=FALSE)
rownames(counts.HBCC) = counts.HBCC$Geneid
counts.HBCC = counts.HBCC[,-c(1:6)]

# # HBCC - sgACC
counts.HBCC4 = fread(synGet('syn21867866')$path, data.table=FALSE)
rownames(counts.HBCC4) = counts.HBCC4$Geneid
counts.HBCC4 = counts.HBCC4[,-c(1:6)]

geneCountsMerged = cbind(counts.MPP, counts.HBCC, counts.HBCC4)
```

### Data download
#### Obtain effective count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
downloadFile <- function(id){
  fread(synGet(id)$path, data.table = F)
}
downloadFile_version <- function(id , version){
  fread(synGet(id, version = version)$path, data.table = F)
}

# Get RNASeq QCmetadata
METADATA_QC_DLPFC_ID = 'syn16816488' 
ALL_USED_IDs = METADATA_QC_DLPFC_ID
metadata = downloadFile_version(METADATA_QC_DLPFC_ID, version = 19) %>%
  dplyr::select(`Individual_ID`, Institution, Cohort, `Reported_Gender`, Sex, Ethnicity, ageOfDeath, `PMI_(in_hours)`, Dx, 
                Sample_RNA_ID, one_of('rnaSeq_isolation:RIN',
                                      'rnaSeq_dissection:Brain_Region', 
                                      'rnaSeq_isolation:RNA_Isolation_Batch',
                                      'rnaSeq_report:Ribozero_Batch', 'rnaSeq_report:Library_Batch', 
                                      'rnaSeq_report:Flowcell_Batch', 'rnaSeq_report:Exclude?',
                                      'rnaSeq_report:Mapped_Reads',
                                      'rnaSeq_report:Intragenic_Rate','rnaSeq_report:Intronic_Rate',
                                      'rnaSeq_report:Intergenic_Rate','rnaSeq_report:Genes_Detected',
                                      'rnaSeq_report:Expression_Profiling_Efficiency',
                                      'rnaSeq_report:rRNA_Rate',
                                      'rnaSeq_report:Total_Reads', 'rnaSeq_report:Percent_Aligned','rnaSeq_report:Transcripts_Detected'))

# Get ancestry vector calculated using gemtools
ANCESTRY_ID = 'syn9922992'
ALL_USED_IDs = c(ALL_USED_IDs, ANCESTRY_ID)
ANCESTRY.HBCC = downloadFile(ANCESTRY_ID) %>% 
  plyr::rename(c(ID = 'SNP_report:Genotyping_Sample_ID'))

ANCESTRY_ID = 'syn2511399'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = ANCESTRY_ID
ANCESTRY.MPP = downloadFile(ANCESTRY_ID) %>% 
  plyr::rename(c('DNA_report..Genotyping.Sample_ID' = 'SNP_report:Genotyping_Sample_ID'))

ANCESTRY = rbind(ANCESTRY.HBCC[,colnames(ANCESTRY.MPP)[-2]], ANCESTRY.MPP[,colnames(ANCESTRY.MPP)[-2]])

# Get genotype ids from synapse. 
GENOTYPE_ID = 'syn16816490'
ALL_USED_IDs = c(ALL_USED_IDs, GENOTYPE_ID)
GENOTYPE = downloadFile(GENOTYPE_ID) %>% 
  dplyr::select(Individual_ID, `SNP_report:Genotyping_Sample_ID`, `SNP_report:Exclude?`) %>% 
  dplyr::inner_join(ANCESTRY) %>% 
  dplyr::filter(is.na(`SNP_report:Exclude?`))
```

### Data preprocessing
```{r preprocess.data}
# Merge metadata
METADATA = metadata %>%
  dplyr::left_join(GENOTYPE) %>%      
  dplyr::rename(Region = `rnaSeq_dissection:Brain_Region`,
                PMI = `PMI_(in_hours)`,
                RIN = `rnaSeq_isolation:RIN`, 
                ReportExclude = `rnaSeq_report:Exclude?`,
                GenotypeExclude = `SNP_report:Exclude?`,
                SampleID = Sample_RNA_ID,
                LibraryBatch = `rnaSeq_report:Library_Batch`,
                FlowcellBatch = `rnaSeq_report:Flowcell_Batch`,
                IsolationBatch = `rnaSeq_isolation:RNA_Isolation_Batch`, #isolation batch to replace ribozero batch
                RibozeroBatch = `rnaSeq_report:Ribozero_Batch`,
                MappedReads = `rnaSeq_report:Mapped_Reads`,
                IntragenicRate = `rnaSeq_report:Intragenic_Rate`, IntronicRate = `rnaSeq_report:Intronic_Rate`, 
                IntergenicRate = `rnaSeq_report:Intergenic_Rate`, GenesDetected = `rnaSeq_report:Genes_Detected`,
                ExpProfEfficiency = `rnaSeq_report:Expression_Profiling_Efficiency`, rRNARate = `rnaSeq_report:rRNA_Rate`,
                TotalReads = `rnaSeq_report:Total_Reads`, AlignmentRate = `rnaSeq_report:Percent_Aligned`, TranscriptsDetected = `rnaSeq_report:Transcripts_Detected`) %>%
  dplyr::select(SampleID, Individual_ID, Institution, Cohort, Reported_Gender,Sex, Ethnicity, ageOfDeath, PMI, Dx, Region,
                RIN, ReportExclude, GenotypeExclude, EV.1, EV.2, EV.3, EV.4, EV.5,
                LibraryBatch, FlowcellBatch, IsolationBatch, RibozeroBatch, MappedReads, TotalReads, GenesDetected, AlignmentRate,
                IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate, TranscriptsDetected) %>%
  		 		 dplyr::filter(SampleID %in% colnames(geneCountsMerged), !is.na(SampleID)) %>%
  		 		 dplyr::filter(Region %in% c('DLPFC')) %>%
			 	 dplyr::mutate(Cohort = forcats::fct_recode(Cohort, `MSSM-Penn-Pitt` = 'MSSM-Penn-Pitt', `NIMH-HBCC`='NIMH-HBCC', `NIMH-HBCC`='NIMH-HBCC-4')) %>%
			 	 dplyr::filter(Dx %in% c('Control', "SCZ"))  %>%
         dplyr::filter(ageOfDeath > 17)

# Set Control to baseline
METADATA$Dx = factor(METADATA$Dx, c('Control', 'SCZ'))

ind = METADATA$SampleID[which(METADATA$ReportExclude == 1 | METADATA$GenotypeExclude)]
writeLines(paste('Following',length(ind),'samples are marked exclude'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$Ethnicity) | is.na(METADATA$Institution) | is.na(METADATA$Dx)]
writeLines(paste('Following', length(ind), 'counts are missing any metadata'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$PMI)]
writeLines(paste('Following', length(ind), 'counts are missing PMI'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$Reported_Gender)]
writeLines(paste('Following', length(ind), 'counts are missing gender'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$ageOfDeath)]
writeLines(paste('Following', length(ind), 'counts are missing age of death'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$EV.1)]
writeLines(paste('Following', length(ind), 'counts are missing ancestry information'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

# drop individuals who where sequenced twice here
tab = table(METADATA$Individual_ID)

dropSamples = sapply( names(tab[tab > 1]), function(id){

  idx = which(METADATA$Individual_ID == id)
  df = METADATA[idx,]

  # remove individuals that have less then max reads
  i = which(df$MappedReads < max(df$MappedReads))
  rownames(METADATA)[idx[i]]
  } )
METADATA = METADATA[!(rownames(METADATA) %in% dropSamples),]
```

```{r cellFractionEstimates}
# include estimated cell fractions in METADATA
df_cellFractions = read.table(synGet('syn22333694')$path, row.names=1)

celFrac_ilr = ilr(df_cellFractions)
colnames(celFrac_ilr) = paste0("cellFrac_ilr_", 1:3)

METADATA = merge(METADATA, df_cellFractions, by.x = 'SampleID', by.y='row.names')
METADATA = merge(METADATA, celFrac_ilr, by.x = 'SampleID', by.y='row.names')
```

```{r preprocess.data1, results='asis'}
# Match covariates to expression data
indToRetain = intersect(METADATA$SampleID, colnames(geneCountsMerged))

geneCountsMerged = geneCountsMerged[,indToRetain]

rownames(METADATA) = METADATA$SampleID
```

## Summarize samples
```{r summary}
METADATA %>% group_by(Dx, Institution) %>% dplyr::summarise(count = n()) %>% 
  kable(caption="Diagnosis, Institution") %>% kable_styling(full_width=FALSE)

METADATA %>% group_by( Institution, Reported_Gender) %>% dplyr::summarise(count = n()) %>% 
  kable(caption="Institution, Sex") %>% kable_styling(full_width=FALSE)

METADATA %>% group_by(Dx, Institution, Reported_Gender) %>% dplyr::summarise(count = n()) %>% 
kable(caption="Diagnosis, Institution, Sex") %>% kable_styling(full_width=FALSE) 
```

```{r joint.cca, fig.width=7, fig.height=7}
variables = c('Reported_Gender', 'CellType', 'ageOfDeath', 'Institution', "Dx")

df_cca = lapply(variables, function(v1){

  if( v1 == "CellType"){
    form1 = ~ GABA + GLU + Olig + MgAs
  }else{
    form1 = as.formula(paste0('~', v1))
  }
  M1 = model.matrix(form1, METADATA)

  res = lapply(variables, function(v2){

    if( v2 == "CellType"){
      form2 = ~ GABA + GLU + Olig + MgAs
    }else{
      form2 = as.formula(paste0('~', v2))
    }
    M2 = model.matrix(form2, METADATA)

    fit = cancor(M1, M2)
    V_sqrt = sqrt(mean(fit$cor^2)) # Cramer's sqrt V

    data.frame(v1, v2, V_sqrt)
  })
  do.call(rbind, res)
})
df_cca = do.call(rbind, df_cca)

C = matrix(df_cca$V_sqrt, ncol=length(variables))
rownames(C) = variables
colnames(C) = variables

plotCorrMatrix(C, main='CCA')
```

```{r explain.CellType}
CTs = c('GABA', 'GLU', 'Olig', 'MgAs')

form = ~ ageOfDeath:Institution + Reported_Gender + Dx + Institution + Dx

Y = clr(METADATA[,CTs])

vp = fitExtractVarPartModel( t(Y), form, METADATA, BPPARAM=SerialParam())
plotPercentBars(vp) + ggtitle("Explain cell type composition")

ymax = max(rowSums(vp[,-ncol(vp)]))*110
plotPercentBars(vp) + scale_y_continuous(expand=c(0,0), limits=c(0, ymax))
```

```{r sex.graph, fig.width=7, fig.height=7}
library(ggplot2)
library(ggraph)
library(tidygraph)

df_cca2 = df_cca[with(df_cca, v1!=v2),]

rstat_nodes <- data.frame(name = unique(df_cca2$v1))

rstat_edges <- data.frame(from = match(df_cca2$v1, rstat_nodes$name),
                         to = match(df_cca2$v2, rstat_nodes$name),
                         Rsq = df_cca2$V_sqrt^2 )

net = tbl_graph(nodes = rstat_nodes, edges = rstat_edges, directed=FALSE)

ggraph(net,  layout = 'igraph', algorithm = 'star') + 
  geom_edge_link(aes(width = Rsq, color=Rsq), alpha = 1) +
  geom_node_point(size=20, color="grey70") +
  geom_node_text(aes(label = name), repel = FALSE) +
  theme_graph() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) +
  scale_edge_width(name=bquote(R^2)) +
  scale_edge_color_gradient(name=bquote(R^2), low="white", high="red", lim=c(0, max(df_cca2$V_sqrt^2))) +
  ggtitle("Pairwise squared correlation")
```

```{r sex.confounding, fig.width=5, fig.height=5}
form = ~ Reported_Gender + GABA + GLU + Olig + MgAs + ageOfDeath 
par(mfrow=c(2,2))
res = lapply( unique(METADATA$Institution), function(INST){
  i = METADATA$Institution == INST
  C = canCorPairs(form, METADATA[i,]) 
  plotCorrMatrix(C, main=INST)
})
```

```{r sex.by.age}
ggplot(METADATA, aes(Reported_Gender, ageOfDeath, fill=Reported_Gender)) + geom_boxplot() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + facet_wrap(~Institution, nrow=1) + scale_fill_manual(values = c("red", "blue"))
```

### Sex by cell fraction
```{r summary.sex, fig.width=8, fig.height=10}
figList = lapply(colnames(df_cellFractions), function(CT){
  ggplot(METADATA, aes_string('Reported_Gender', CT, fill='Reported_Gender')) + geom_boxplot() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1) + scale_fill_manual(values = c("red", "blue"))
})
plot_grid(plotlist=figList, ncol=1) 
```

### Age by cell fraction
```{r summary.Age, fig.width=8, fig.height=10}
figList = lapply(colnames(df_cellFractions), function(CT){
  ggplot(METADATA, aes_string('ageOfDeath', CT)) + geom_point() + geom_smooth(method="gam") + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1)
}) 
plot_grid(plotlist=figList, ncol=1) 
```

## Age by cell fraction log ratio
```{r summary.Age.log.ratio, fig.width=8, fig.height=8}
figList = lapply(colnames(celFrac_ilr), function(CT){
  ggplot(METADATA, aes_string('ageOfDeath', CT)) + geom_point() + geom_smooth(method="gam") + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1)
})
plot_grid(plotlist=figList, ncol=1) 
```

```{r process.counts}
# keep genes that are express in at least one combination of cohort and gender
isexpr.lst = dlply(METADATA, .(Cohort), .fun = function(mtd){
	rowSums(cpm(geneCountsMerged[,mtd$SampleID])>.5) >= .3*ncol(geneCountsMerged[,mtd$SampleID])
	})

isexpr = do.call(rbind, isexpr.lst)
isexpr = colSums(isexpr) == 2 # keep gene expressed in both cohorts 
table(isexpr)

# Normalize each cohort separately
dge.lst = lapply( levels(METADATA$Cohort), function(Cohort){
  # normalize counts
  dge = DGEList( geneCountsMerged[isexpr,METADATA$Cohort==Cohort] )
  calcNormFactors(dge, "none") # TMM
})
names(dge.lst) = levels(METADATA$Cohort)

# form = ~ (1|Dx) + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)

form.lst = list(
  `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
  `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)

vobj.lst = lapply(dge.lst, function(dge){

  i = match(colnames(dge), rownames(METADATA))
  info = METADATA[i,]

  form_use = form.lst[[info$Cohort[1]]]

  voomWithDreamWeights(dge, form_use, info, save.plot=TRUE, BPPARAM = SnowParam(18))
  })
names(vobj.lst) = names(dge.lst)
```

### Voom trend plots
```{r plot.voom, fig.height=5, fig.width=10}
fig1 = plot_voom(vobj.lst[[1]]) + ggtitle(names(vobj.lst)[1])
fig2 = plot_voom(vobj.lst[[2]]) + ggtitle(names(vobj.lst)[2])
plot_grid( fig1, fig2, nrow=1)

df = lapply( names(vobj.lst), function(key){
  data.frame(Cohort = key,  vobj.lst[[key]]$voom.line)
  })
df = do.call(rbind, df)

ggplot(df, aes(x, y, color=Cohort)) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + geom_line(size=1.2) + xlab(bquote(log[2](count + 0.5))) + ylab(expression( sqrt("standard deviation"))) + ggtitle("Compare voom trends") + ylim(0, max(df$y)) + scale_color_manual(name="Status", values=c("#3758ba", "#33e65c")) + guides(color = guide_legend(override.aes = list(size = 1))) 
```

<!---

# Examine genes on chrX and chrY

# dge = DGEList(geneCountsMerged)
# dge = calcNormFactors(dge, "none")
# geneExpr = cpm(dge, log=TRUE)

# df = merge( data.frame(Expression = geneExpr['ENSG00000229807.11',]), METADATA, by="row.names")
# ggplot(df, aes(Reported_Gender, Expression)) + geom_violin() + theme_bw() + facet_wrap(~Cohort)

# id = 'ENSG00000229807.11'
# isexpr[id]

# df = as.data.frame(vobj.lst[[2]]$voom.xy)
# with(df, plot(x,y, cex=.2, pch=20))
# with(df[id,], points(x,y, col="red", pch=20, cex=2))

--->

```{r Create.residuals}

### Create residuals
form.lst = list(
  `MSSM-Penn-Pitt`  = ~ Dx*Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
  `NIMH-HBCC`       = ~ Dx*Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)

eval_residuals = function( form ){
  # Compute residuals, but add back in Dx*Reported_Gender
  # for form = ~ Dx*Reported_Gender
  f = function(fit){
    residuals(fit) + get_prediction( fit, form)
  }

  residuals.lst = lapply( vobj.lst, function(vobj){

    i = match(colnames(vobj), rownames(METADATA))
    info = METADATA[i,]

    form_use = form.lst[[info$Cohort[1]]]

    # Compute residuals which adding back Diagnosis and Sex effects
    resid.lst = fitVarPartModel(vobj, form_use, info, fxn = f, BPPARAM = SnowParam(18))  
    do.call(rbind, resid.lst)
    })
  names(residuals.lst) = names(vobj.lst)

  residuals.lst
}

resid.add = list( Dx_Sex  = eval_residuals( ~ Dx*Reported_Gender ),
                  Dx      = eval_residuals( ~ Dx ),
                  Sex     = eval_residuals( ~ Reported_Gender ))

residuals.lst = resid.add[['Dx_Sex']]
```

```{r check.correlation, eval=FALSE}
ngenes = 300
C1 = cor(t(residuals.lst[[1]][1:ngenes,]))
C2 = cor(t(residuals.lst[[2]][1:ngenes,]))

df = data.table(A   = as.numeric(C1[lower.tri(C1)]),
                B   = as.numeric(C2[lower.tri(C2)]))

df$density <- get_density(df$A, df$B, n = 80)
ggplot(df, aes(A, B, color=density)) + geom_point() + geom_point(size=.4) + theme_bw(15) + theme(aspect.ratio=1, legend.position="bottom", plot.title = element_text(hjust = 0.5)) + geom_abline(color="red") + scale_color_viridis() + xlim(-1,1) + ylim(-1,1) + xlab(paste0('Correlation (', names(residuals.lst)[1], ')')) + ylab(paste0('Correlation (', names(residuals.lst)[2], ')'))
```

# Schizophrenia vs controls

### Test marginals for Scz vs controls
```{r differential.expression.simple}
form.lst = list(
  `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
  `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)

fit_marginal.lst = lapply( vobj.lst, function(vobj){

  i = match(colnames(vobj), rownames(METADATA))
  info = METADATA[i,]

  form_use = form.lst[[info$Cohort[1]]]

  fit = dream(vobj, form_use, info, BPPARAM = SnowParam(18))  
  eBayes( fit )
  })
names(fit_marginal.lst) = names(vobj.lst)

coefTest = c('DxSCZ', 'Reported_GenderMale', 'ageOfDeath')
```

```{r load.genesets, eval=TRUE}
# Load Gene Ontology database 
gs.GO = get_GeneOntology()
gs.MSigDB = get_MSigDB()

# convert from GeneSetCollection to list used by camera and zenith
gs.GO.collection = recodeToList( gs.GO )
gs.MSigDB.collection = recodeToList( gs.MSigDB )
```

```{r run.zenith, eval=TRUE}
res_zenith = lapply( names(fit_marginal.lst), function(key){

  fit = fit_marginal.lst[[key]]

  res.GO = run_zenith(  fit, 
                        coefTest,
                        gs.GO.collection)
  res.GO$Resource = "GO"   
  res.GO$Cohort = key

  res.MSigDB = run_zenith(fit, 
                          coefTest,
                          gs.MSigDB.collection ) 
  res.MSigDB$Resource = "MSigDB"
  res.MSigDB$Cohort = key

  rbind(res.GO, res.MSigDB)
})
names(res_zenith) = names(fit_marginal.lst)

# only autosomes
res_zenith_sex_autosomes = lapply( names(fit_marginal.lst), function(key){

  fit = fit_marginal.lst[[key]]

  # only keep autosomes, for Reported_GenderMale
  tab = coef(fit) %>% getGeneSymbol()
  geneids = rownames(tab)[tab$Chrom %in% 1:22]
  fit = fit[geneids,]

  res.GO = run_zenith(  fit, 
                        'Reported_GenderMale',
                        gs.GO.collection)
  res.GO$Resource = "GO"   
  res.GO$Cohort = key

  res.MSigDB = run_zenith(fit, 
                          'Reported_GenderMale',
                          gs.MSigDB.collection ) 
  res.MSigDB$Resource = "MSigDB"
  res.MSigDB$Cohort = key

  rbind(res.GO, res.MSigDB)
})
names(res_zenith_sex_autosomes) = names(fit_marginal.lst)
```

## Meta-analysis of zenith results
```{r zenith.meta, eval=TRUE}
res_meta_zenith = zenith_meta_analysis( res_zenith )                  
res_meta_zenith.sex_autosome = zenith_meta_analysis( res_zenith_sex_autosomes )
```

```{r get.signatures, fig.width=10, fig.height=5}
# Get DE signatures
df_DE_results = lapply(fit_marginal.lst, function(fit){
  res = lapply(coefTest, function(coef){
    res = topTable(fit, coef, number=Inf) 
    res$Coef = coef
    res %>% getGeneSymbol()  
    })
  names(res) = coefTest
  res
  })
names(df_DE_results) = names(fit_marginal.lst)
```

```{r test.scz, fig.height=5, fig.width=10}
# Schizophrenia testing
# extract from each cohort
coef = 'DxSCZ'
# ymin = min(sapply(df_DE_results, function(x) min(x[[coef]]$P.Value)))

figList.Scz = lapply( names(df_DE_results), function(key){
  df = df_DE_results[[key]][[coef]]

  main = paste(key, coef, sep = ': ' )
  plotVolcano( df, df$Symbol[1:10] ) + ggtitle( main ) #+ ylim(0, -log10(ymin))
})

# volcano plots
plot_grid(plotlist=figList.Scz, nrow=1)
```

## Meta-analysis across cohorts
```{r meta.analysis.DE}
# Store results for each coef and cohort
tab.de.lst = lapply(coefTest, function(coef){    
  tab.lst = lapply( df_DE_results, function(lst){
    lst[[coef]]
  })
  names(tab.lst) = names(df_DE_results)
  tab.lst
  })
names(tab.de.lst) = coefTest

# Create list preparing for meta-analysis for SCZ
df_meta = lapply(coefTest, function(coef){
  # Run meta-analysis
  run_meta_analysis( tab.de.lst[[coef]], "FE") %>% getGeneSymbol
})#, mc.cores=6)
names(df_meta) = coefTest
```

```{r concordance.scz, fig.height=6, fig.width=10}
# Concordance analysis
resConcord.SCZ = plot_concordance(tab.de.lst[[coef]], showGenes = df_meta[[coef]]$Symbol[1:20])

main = "Concordance of SCZ"
plot_grid(  resConcord.SCZ$fig_logFC + ggtitle(main),
            resConcord.SCZ$fig_tstat + ggtitle(main), nrow=1)

data.frame(Cohort = names(tab.de.lst[[coef]]),
          "# DE"  = resConcord.SCZ$n.de,
          pi1     = resConcord.SCZ$pi_single,
          check.names=FALSE) %>% 
          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)

resConcord.SCZ$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
resConcord.SCZ$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
```

```{r plot.meta.analysis.SCZ, fig.height=5, fig.width=10}
# volcano plot for meta-analysis
main = paste('Meta-analysis', coef, sep = ': ' )
plotVolcano( df_meta[[coef]], df_meta[[coef]]$Symbol[1:20] ) + ggtitle( main )

data.frame(pi1 = 1 - qvalue(df_meta[[coef]]$P)$pi0,
          `# DE` = sum(df_meta[[coef]]$adj.P.Val < 0.05),
          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for SCZ meta-analysis") %>% kable_styling(full_width=FALSE)

```

## Plot top examples
```{r plot.forrest.dlpfc, fig.width=8, fig.height=20}  
plot_gene = function(residuals.lst, df_meta, geneSymbol, variable, fill=c("#1E90FF", "#FF8D1E"), mode=""){

  ensGene = rownames(df_meta)[which(df_meta$Symbol == geneSymbol)]

  exprResid = lapply( names(residuals.lst), function(key){
      data.frame(Expression = residuals.lst[[key]][ensGene,])
    })
  exprResid = do.call(rbind, exprResid)

  df = merge(exprResid, METADATA, by="row.names")

  if( is.factor(df[[variable]]) || length(unique(df[[variable]])) < 5 ){
    fig = ggplot(df, aes_string(variable, 'Expression')) + geom_violin(aes_string(fill=variable)) + geom_boxplot(width=0.1, fill="grey80") + scale_fill_manual(values = fill)
  }else{    
    fig = ggplot(df, aes_string(variable, 'Expression')) + geom_point() + geom_smooth(method="lm") #+ scale_color_manual(values = fill)
  }

  fig + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + ylab('Expression residuals') + ggtitle(geneSymbol) + facet_wrap(~Cohort)
}

coef = 'DxSCZ'
figList = lapply(df_meta[[coef]]$Symbol[1:10], function(geneSymbol){
  fig1 = plot_forrest( c(tab.de.lst[[coef]], list("Meta-analysis" = df_meta[[coef]])), geneSymbol)

  fig2 = plot_gene(residuals.lst, df_meta[[coef]], geneSymbol, "Dx") + xlab("Diagnosis")

  plot_grid(fig1, fig2, nrow=1)
})

plot_grid(plotlist=figList, ncol=1)
```

### Concordance with other datasets
```{r CMC.concordance}
# CMC Analysis 1
CMC1 <- readr::read_tsv(synGet("syn5607603")$path) %>% 
  dplyr::rename(gene_id = genes,
                `CMC 1 t-statistic` = t) %>% 
  dplyr::select(gene_id, `CMC 1 t-statistic`)
# HBCC-DLPFC microarray
CMC_HBCC <- readr::read_tsv(synGet("syn4941553")$path) %>% 
  dplyr::rename(gene_id = genes,
                `HBCC microarray t-statistic` = t) %>% 
  dplyr::select(gene_id, `HBCC microarray t-statistic`)
# Gandal, Science 2018, Table S1
# Use Tab:DGE
# wget https://science.sciencemag.org/highwire/filestream/720110/field_highwire_adjunct_files/1/aat8127_Table_S1.xlsx
gandal <- readxl::read_xlsx( synGet("syn22255217")$path, sheet = "DGE") %>% 
  dplyr::rename(gene_id = ensembl_gene_id,
                `Gandal, SCZ t-statistic` = `SCZ.t.value`,
                `Gandal, BD t-statistic` = `BD.t.value`,
                `Gandal, ASD t-statistic` = `ASD.t.value`) %>%
  dplyr::select(gene_id, `Gandal, SCZ t-statistic`,
                `Gandal, BD t-statistic`, `Gandal, ASD t-statistic`)
# AMP-AD RNASeq reprocessing GRCh37 - https://www.synapse.org/#!Synapse:syn9702085
MSSM <- readr::read_tsv(synGet("syn10526259")$path) %>%
  dplyr::filter(Comparison %in% "AD-CONTROL",
                Model %in% "Diagnosis") %>%
  dplyr::select(ensembl_gene_id, Tissue, t) %>%
  dplyr::group_by(ensembl_gene_id, Tissue) %>% 
  dplyr::filter(row_number() == 1) %>% 
  tidyr::pivot_wider(names_from = Tissue,
                     values_from = t) %>%
  dplyr::rename(`AMP-AD MSSM FP AD t-statistic` = FP,
                `AMP-AD MSSM IGF AD t-statistic` = IFG,
                `AMP-AD MSSM PHG AD t-statistic` = PHG,
                `AMP-AD MSSM STG AD t-statistic` = STG,
                gene_id = ensembl_gene_id) %>% 
  dplyr::select(gene_id, `AMP-AD MSSM STG AD t-statistic`) 
MSSM <- MSSM[!is.na(MSSM$`AMP-AD MSSM STG AD t-statistic`),]

MAYO <- readr::read_tsv(synGet("syn8468023")$path) %>% 
  dplyr::filter(Comparison %in% "AD-CONTROL",
                Model %in% "Diagnosis") %>%
  dplyr::select(ensembl_gene_id, `Tissue.ref`, t) %>%
  dplyr::group_by(ensembl_gene_id, `Tissue.ref`) %>% 
  dplyr::filter(row_number() == 1) %>% 
  tidyr::pivot_wider(names_from = `Tissue.ref`,
                     values_from = t) %>% 
  dplyr::rename(`AMP-AD MAYO CBE AD t-statistic` = CBE,
                `AMP-AD MAYO TCX AD t-statistic` = TCX,
                gene_id = ensembl_gene_id) %>% 
  dplyr::select(-`AMP-AD MAYO TCX AD t-statistic`)

ROSMAP <- readr::read_tsv(synGet("syn8456721")$path) %>%
    dplyr::filter(Comparison %in% "AD-CONTROL",
                Model %in% "Diagnosis") %>%
  dplyr::rename(gene_id = ensembl_gene_id, 
                `AMP-AD ROSMAP DLPFC AD t-statistic` = t) %>% 
  dplyr::select(gene_id, `AMP-AD ROSMAP DLPFC AD t-statistic`)

out_group <- inner_join(MSSM, MAYO) %>% 
  inner_join(., ROSMAP) %>% 
  inner_join(., gandal)

# merge with meta-analysis

coef = 'DxSCZ'
df = df_meta[[coef]]
df_two = data.frame(gene_id = trim_ensembl_ids(rownames(df)),
                `CMC - meta-analysis` = with(df, logFC/se), check.names=FALSE)

df = tab.de.lst[[coef]][[1]]
df.tmp = data.frame(gene_id = trim_ensembl_ids(rownames(df)),
                `t` = df$t, check.names=FALSE)
colnames(df.tmp)[2] = names(tab.de.lst[[coef]])[1]
df_two = df_two %>% inner_join(., df.tmp)

df = tab.de.lst[[coef]][[2]]
df.tmp = data.frame(gene_id = trim_ensembl_ids(rownames(df)),
                `t` = df$t, check.names=FALSE)
colnames(df.tmp)[2] = names(tab.de.lst[[coef]])[2]
df_two = df_two %>% inner_join(., df.tmp)

# Remove transcript Ids
df_two <- df_two %>% 
  tidyr::separate(gene_id, c("gene_id", "transcript")) %>% 
  dplyr::select(-one_of("transcript"))
df_two <- inner_join(df_two, CMC1, by = c("gene_id")) 
df_two <- inner_join(df_two, CMC_HBCC, by = c("gene_id"))
df_two <- inner_join(df_two, out_group, by = c("gene_id")) %>% 
  group_by(gene_id) %>% 
  dplyr::filter(row_number()==1) %>% 
  tibble::column_to_rownames(var = "gene_id")

colnames(df_two) <- gsub(" t-statistic", "", colnames(df_two))

col2 <- colorRampPalette(rev(c("#67001F", "#B2182B", "#D6604D", "#F4A582",
                           "#FDDBC7", "#FFFFFF", "#D1E5F0", "#92C5DE",
                           "#4393C3", "#2166AC", "#053061")))

cor <- cor(df_two)
# clustering based on distance
hcl = hclust(as.dist(1-cor), method='ward.D2')

results <- pvclust(cor, method.hclust='ward.D2')
# reorder based on clustering
reorder = cor[results$hclust$order, results$hclust$order]

# plot clustering
plot(results, hang=-1)
# plot(hcl, hang=-1)
corrplot(reorder, col = col2(200))
```

```{r test.multtest, eval=FALSE}

library(irlba)
library(pinnacle)
library(poolR)
library(matrixStats)

i = with(res_meta_zenith, (Resource == 'GO') & (Coef == "DxSCZ"))
df = res_meta_zenith[i,]
df$qvalue = qvalue(df$PValue)$qvalue
# df$p.adjust = p.adjust(df$PValue, method="bon" )
df$p.adjust = pmin(1, 2805 * df$PValue)

M = recodeToSparseMatrix(gs.GO)
M = M[rownames(M) %in% trim_ensembl_ids(rownames(df_meta_SCZ)),]
keep = (colSums(M) >= 10) & (colnames(M) %in% rownames(df) )
M = M[,keep]

m = colMeans(M)
s = apply(M, 2, var) * nrow(M) / (nrow(M) + 1)

dcmp = irlba( M, nv = 500, center=m, scale=sqrt(s), right_only=FALSE)
ev = dcmp$d^2 / nrow(M)

plot(cumsum(ev) / ncol(M))

df_unsort = df[colnames(M),]
edcmp = list(vectors = dcmp$v, values = dcmp$d^2 / nrow(M))

res = p.corr.adjust( df_unsort$PValue, dcmp = edcmp, method.sim="eigen", B = nrow(df_unsort)*4)

plot( cumsum(dcmp$d^2 / sum(dcmp$d^2)))

C = cor(as.matrix(M[,1:1000]))
evals = eigen(C)$values

dcmp = svd(scale(M) / sqrt(nrow(M)))
dcmp$d^2

plot(cumsum(dcmp$d^2 / sum(dcmp$d^2)))

meff(C, method='gao')

dcmp = svd(scale(M[,1:100]))

meff(eigen = dcmp$d^2 , method = "galwey")

```

```{r show.gene_set_test.marginal, eval=TRUE}
show_zenith_results = function( res, coef, Resource, FDR = 0.05){

  i = which((res$Resource == Resource) & (res$Coef == coef))
  df = res[i,]

	# filter by FDR
	df = df[df$FDR <= FDR,]
  df$Coef = c()
  df$se = c()
  df$Resource = c()

	df %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption=paste(Resource, coef), row.names=FALSE) %>% kable_styling(full_width=FALSE)
}

show_zenith_results( res_meta_zenith, "DxSCZ", "GO", FDR=0.05)
show_zenith_results( res_meta_zenith, "DxSCZ", "MSigDB", FDR=0.05)
```

# Sex Analysis
```{r test.Sex, fig.height=5, fig.width=10}
# Schizophrenia testing
# extract from each chort
coef = 'Reported_GenderMale'
figList.Sex = lapply( names(df_DE_results), function(key){
  df = df_DE_results[[key]][[coef]]
  main = paste(key, coef, sep = ': ' )
  plotVolcano( df, df$Symbol[1:20] ) + ggtitle( main )
})

# volcano plots
plot_grid(plotlist=figList.Sex, nrow=1)
```

### Autosomes
```{r test.Sex.autosome, fig.height=5, fig.width=10}
# Schizophrenia testing
# extract from each chort
coef = 'Reported_GenderMale'
figList.Sex = lapply( names(df_DE_results), function(key){
  df = df_DE_results[[key]][[coef]]
  main = paste(key, coef, sep = ': ' )
  df = df[df$Chrom %in% 1:22,]
  plotVolcano( df, df$Symbol[1:20] ) + ggtitle( main )
})

# volcano plots
plot_grid(plotlist=figList.Sex, nrow=1)
```

## Meta-analysis across cohorts
```{r concordance.Sex, fig.height=6, fig.width=10}
# Concordance analysis
resConcord.Sex = plot_concordance(tab.de.lst[[coef]], showGenes = df_meta[[coef]]$Symbol[1:20])

main = "Concordance of Sex"
plot_grid(  resConcord.Sex$fig_logFC + ggtitle(main),
            resConcord.Sex$fig_tstat + ggtitle(main), nrow=1)

data.frame(Cohort = names(tab.de.lst[[coef]]),
          "# DE"  = resConcord.Sex$n.de,
          pi1     = resConcord.Sex$pi_single,
          check.names=FALSE) %>% 
          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)

resConcord.Sex$tab_pi  %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
resConcord.Sex$tab_corr  %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
```

```{r concordance.Sex.autosome, fig.height=6, fig.width=10}

tab.de.lst.autosome = lapply(tab.de.lst[[coef]], function(x) x[x$Chrom %in% 1:22,])

# Concordance analysis
resConcord.Sex = plot_concordance(tab.de.lst.autosome, showGenes = tab.de.lst.autosome$Symbol[1:20])

main = "Concordance of Sex - autosomes"
plot_grid(  resConcord.Sex$fig_logFC + ggtitle(main),
            resConcord.Sex$fig_tstat + ggtitle(main), nrow=1)

data.frame(Cohort = names(resConcord.Sex$pi_single),
          "# DE"  = resConcord.Sex$n.de,
          pi1     = resConcord.Sex$pi_single,
          check.names=FALSE) %>% 
          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)

resConcord.Sex$tab_pi %>% mutate_if(is.numeric, format, digits=3)  %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
resConcord.Sex$tab_corr  %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
```

```{r plot.meta.analysis.Sex, fig.height=5, fig.width=10}
# volcano plot for meta-analysis
main = 'Meta-analysis: Sex'
fig1 = plotVolcano( df_meta[[coef]], df_meta[[coef]]$Symbol[1:20] ) + ggtitle( main )

df = df_meta[[coef]][df_meta[[coef]]$Chrom %in% 1:22,]
fig2 = plotVolcano( df, df$Symbol[1:20] ) + ggtitle( paste(main, "autosome") )
plot_grid(fig1, fig2, ncol=2)

data.frame(pi1 = 1 - qvalue(df_meta[[coef]]$P)$pi0,
          `# DE` = sum(df_meta[[coef]]$adj.P.Val < 0.05),
          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for Sex meta-analysis") %>% kable_styling(full_width=FALSE)
```

## Sex circos
```{r sex.circos}
coef = 'Reported_GenderMale'
sapply( names(df_DE_results), function(key){
  plot_circos_logFC( df_DE_results[[key]][[coef]] ) + ggtitle(key)
  })

plot_circos_logFC( df_meta[[coef]] ) + ggtitle("Meta analysis")
```

## Plot top examples
```{r plot.forrest.dlpfc.Sex, fig.width=8, fig.height=20}  
figList = lapply(df_meta[[coef]]$Symbol[1:10], function(geneSymbol){

  fig1 = plot_forrest( c(tab.de.lst[[coef]], list("Meta-analysis" = df_meta[[coef]])), geneSymbol)

  fig2 = plot_gene( residuals.lst, df_meta[[coef]], geneSymbol, "Reported_Gender", fill=c("red", "blue")) + xlab("Sex")

  plot_grid(fig1, fig2, nrow=1)
})

plot_grid(plotlist=figList, ncol=1)
```

## Meta-analysis of zenith results
```{r show.gene_set_test.marginal.Sex, eval=TRUE}
show_zenith_results( res_meta_zenith, "Reported_GenderMale", "GO", FDR=0.05)         
show_zenith_results( res_meta_zenith, "Reported_GenderMale", "MSigDB", FDR=0.05)        
```

### Only autosomes
```{r show.gene_set_test.marginal.Sex.autosome, eval=TRUE}
show_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "GO", FDR=0.1) 
show_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "MSigDB", FDR=0.1)   
```

# Age Analysis
```{r test.Age, fig.height=5, fig.width=10}
# Schizophrenia testing
# extract from each chort
coef = 'ageOfDeath'
figList.Age = lapply( names(df_DE_results), function(key){
  df = df_DE_results[[key]][[coef]]
  main = paste(key, coef, sep = ': ' )
  plotVolcano( df, df$Symbol[1:20] ) + ggtitle( main )
})

# volcano plots
plot_grid(plotlist=figList.Age, nrow=1)
```

## Meta-analysis across cohorts
```{r concordance.Age, fig.height=6, fig.width=10}
# Concordance analysis
resConcord.Age = plot_concordance(tab.de.lst[[coef]], showGenes = df_meta[[coef]]$Symbol[1:20])

main = "Concordance of Age"
plot_grid(  resConcord.Age$fig_logFC + ggtitle(main),
            resConcord.Age$fig_tstat + ggtitle(main), nrow=1)

data.frame(Cohort = names(tab.de.lst[[coef]]),
          "# DE"  = resConcord.Age$n.de,
          pi1     = resConcord.Age$pi_single,
          check.names=FALSE) %>% 
          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)

resConcord.Age$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
resConcord.Age$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
```

```{r plot.meta.analysis.Age, fig.height=5, fig.width=10}
# volcano plot for meta-analysis
main = 'Meta-analysis: Age'
plotVolcano( df_meta[[coef]], df_meta[[coef]]$Symbol[1:20] ) + ggtitle( main )

data.frame(pi1 = 1 - qvalue(df_meta[[coef]]$P)$pi0,
          `# DE` = sum(df_meta[[coef]]$adj.P.Val < 0.05),
          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for Age meta-analysis") %>% kable_styling(full_width=FALSE)
```

## Plot top examples
```{r plot.forrest.dlpfc.Age, fig.width=8, fig.height=20}  
figList = lapply(df_meta[[coef]]$Symbol[1:10], function(geneSymbol){

  fig1 = plot_forrest( c(tab.de.lst[[coef]], list("Meta-analysis" = df_meta[[coef]])), 
    geneSymbol)

  expr.list = lapply(vobj.lst, function(x)x$E)

  fig2 = plot_gene( expr.list, df_meta[[coef]], geneSymbol, "ageOfDeath", fill=c("red", "blue")) + xlab("Age")

  plot_grid(fig1, fig2, nrow=1)
})

plot_grid(plotlist=figList, ncol=1)
```

## Meta-analysis of zenith results
```{r show.gene_set_test.marginal.Age, eval=TRUE}
show_zenith_results( res_meta_zenith, "ageOfDeath", "GO", FDR=0.01)   
show_zenith_results( res_meta_zenith, "ageOfDeath", "MSigDB", FDR=0.01)    
```

# Sex by Disease interaction
```{r dream.sex.disease}
form.lst = list(
  `MSSM-Penn-Pitt`  = ~ Dx*Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
  `NIMH-HBCC`       = ~ Dx*Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)

# drop chrY genes
df_genes  = data.frame(ENSEMBL = rownames(vobj.lst[[1]])) %>% getGeneSymbol('ENSEMBL')
keep.ensID = df_genes$ENSEMBL[df_genes$Chrom %in% c(1:22, 'X', 'MT')]

fit_inter_sex.disease.lst = lapply( vobj.lst, function(vobj){

  i = match(colnames(vobj), rownames(METADATA))
  info = METADATA[i,]

  form_use = form.lst[[info$Cohort[1]]]

  fit = dream(vobj[keep.ensID,], form_use, info, BPPARAM = SnowParam(18))  
  eBayes( fit )
  })
names(fit_inter_sex.disease.lst) = names(vobj.lst)

coefTest = c('DxSCZ:Reported_GenderMale')
```

```{r run.zenith.inter_sex.disease, eval=TRUE}
res_zenith.inter_sex.disease = lapply( names(fit_inter_sex.disease.lst), function(key){

  fit = fit_inter_sex.disease.lst[[key]]

  res.GO = run_zenith(  fit,   
                        coefTest,
                        gs.GO.collection)
  res.GO$Resource = "GO"   
  res.GO$Cohort = key

  res.MSigDB = run_zenith(fit, 
                          coefTest,
                          gs.MSigDB.collection) 
  res.MSigDB$Resource = "MSigDB"
  res.MSigDB$Cohort = key

  rbind(res.GO, res.MSigDB)
})
names(res_zenith.inter_sex.disease) = names(fit_inter_sex.disease.lst)

# only autosomes
res_zenith.inter_sex.disease_autosome = lapply( names(fit_inter_sex.disease.lst), function(key){

  fit = fit_inter_sex.disease.lst[[key]]

  # only keep autosomes, for Reported_GenderMale
  tab = coef(fit) %>% getGeneSymbol()
  geneids = rownames(tab)[tab$Chrom %in% 1:22]
  fit = fit[geneids,]

  res.GO = run_zenith(  fit, 
                        coefTest,
                        gs.GO.collection)
  res.GO$Resource = "GO"   
  res.GO$Cohort = key

  res.MSigDB = run_zenith(fit, 
                          coefTest,
                          gs.MSigDB.collection) 
  res.MSigDB$Resource = "MSigDB"
  res.MSigDB$Cohort = key

  rbind(res.GO, res.MSigDB)
})
names(res_zenith.inter_sex.disease_autosome) = names(fit_inter_sex.disease.lst)
```

## Meta-analysis of zenith results
```{r zenith.meta.inter_sex.disease, eval=TRUE}
res_meta_zenith.inter_sex.disease = zenith_meta_analysis( res_zenith.inter_sex.disease )  
res_meta_zenith.inter_sex.disease_autosome = zenith_meta_analysis( res_zenith.inter_sex.disease_autosome )  
```

```{r get.signatures.inter_sex.disease, fig.width=10, fig.height=5}
# Get DE signatures
df_DE_results.inter_sex.disease = lapply(fit_inter_sex.disease.lst, function(fit){
  res = lapply(coefTest, function(coef){
    res = topTable(fit, coef, number=Inf) 
    res$Coef = coef
    res %>% getGeneSymbol()  
    })
  names(res) = coefTest 
  res
  })
names(df_DE_results.inter_sex.disease) = names(fit_inter_sex.disease.lst)
```

```{r test.scz.inter_sex.disease, fig.height=5, fig.width=10}
coef = 'DxSCZ:Reported_GenderMale'

figList.Scz = lapply( names(df_DE_results), function(key){
  df = df_DE_results.inter_sex.disease[[key]][[coef]]

  main = paste(key, coef, sep = ': ' )
  plotVolcano( df, df$Symbol[1:10] ) + ggtitle( main )
})

# volcano plots
plot_grid(plotlist=figList.Scz, nrow=1)
```

## Meta-analysis across cohorts
```{r meta.analysis.DE.inter_sex.disease}
# Store results for each coef and cohort
tab.de.inter_sex.disease.lst = lapply(coefTest, function(coef){    
  tab.de.inter_sex.disease.lst = lapply( df_DE_results.inter_sex.disease, function(lst){
    lst[[coef]]
  })
  names(tab.de.inter_sex.disease.lst) = names(df_DE_results.inter_sex.disease)
  tab.de.inter_sex.disease.lst
  })
names(tab.de.inter_sex.disease.lst) = coefTest

# Create list preparing for meta-analysis for SCZ
df_meta.inter_sex.disease = lapply(coefTest, function(coef){
  # Run meta-analysis
  run_meta_analysis( tab.de.inter_sex.disease.lst[[coef]], "FE") %>% getGeneSymbol
}) 
names(df_meta.inter_sex.disease) = coefTest  
```

```{r concordance.scz.inter_sex.disease, fig.height=6, fig.width=10}
# Concordance analysis.  
resConcord.inter_sex.disease = plot_concordance(tab.de.inter_sex.disease.lst[[coef]], showGenes = df_meta.inter_sex.disease[[coef]]$Symbol[1:20])

main = "Concordance of DxSCZ:Reported_GenderMale"
plot_grid(  resConcord.inter_sex.disease$fig_logFC + ggtitle(main),
            resConcord.inter_sex.disease$fig_tstat + ggtitle(main), nrow=1)

data.frame(Cohort = names(tab.de.inter_sex.disease.lst[[coef]]),
          "# DE"  = resConcord.inter_sex.disease$n.de,
          pi1     = resConcord.inter_sex.disease$pi_single,
          check.names=FALSE) %>% 
          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)

resConcord.inter_sex.disease$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
resConcord.inter_sex.disease$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
```

```{r plot.meta.analysis.SCZ.inter_sex.disease, fig.height=5, fig.width=10}
# volcano plot for meta-analysis.  
main = paste('Meta-analysis', coef, sep = ': ' )
plotVolcano( df_meta.inter_sex.disease[[coef]], df_meta.inter_sex.disease[[coef]]$Symbol[1:20] ) + ggtitle( main )

data.frame(pi1 = 1 - qvalue(df_meta.inter_sex.disease[[coef]]$P)$pi0,
          `# DE` = sum(df_meta.inter_sex.disease[[coef]]$adj.P.Val < 0.05),
          `# DE @ 10%` = sum(df_meta.inter_sex.disease[[coef]]$adj.P.Val < 0.1),
          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for SCZ:Sex meta-analysis") %>% kable_styling(full_width=FALSE)
```

## Plot top examples
```{r plot.forrest.dlpfc.inter_sex.disease, fig.width=8, fig.height=40}

figList = lapply(df_meta.inter_sex.disease[[coef]]$Symbol[1:20], function(geneSymbol){
  fig1 = plot_forrest( c(tab.de.inter_sex.disease.lst[[coef]], list("Meta-analysis" = df_meta.inter_sex.disease[[coef]])), geneSymbol, plot2='p')

  fig2 = plot_gene(residuals.lst, df_meta.inter_sex.disease[[coef]], geneSymbol, "Dx") + xlab("Diagnosis") + facet_wrap(~Reported_Gender + Cohort, nrow=1)

  plot_grid(fig1, fig2, nrow=1)
})

plot_grid(plotlist=figList, ncol=1)
```

## Meta-analysis of zenith results
```{r show.gene_set_test.marginal.inter_sex.disease, eval=TRUE}
coef = 'DxSCZ:Reported_GenderMale'   
show_zenith_results( res_meta_zenith.inter_sex.disease, coef, "GO", FDR=0.1)  
show_zenith_results( res_meta_zenith.inter_sex.disease, coef, "MSigDB", FDR=0.1)

show_zenith_results( res_meta_zenith.inter_sex.disease_autosome, coef, "GO", FDR=0.1) 
show_zenith_results( res_meta_zenith.inter_sex.disease_autosome, coef, "MSigDB", FDR=0.1)    
```

# Dx test with in Males and then Females

```{r store, eval=TRUE}
if( params$upload ){    
  # Code
  # Make sure these match the folder name and parent Id in the first chunk.
  CODE <- Folder(name = "Joint analysis of DLPFC count data", parentId = "syn22416298")
  CODE <- synStore(CODE)

  # Store data 
  storeData = function( name, description){
    file = paste0(name, ".RDS")
    saveRDS( eval(parse(text=name)), file=file)

    nEXP_OBJ = File(file, 
                    name = description, 
                    parentId = CODE$properties$id)
    synStore(nEXP_OBJ, used = ALL_USED_IDs, executed = thisFile)
  }

  # Standard analysis
  storeData( "geneCountsMerged", "Merged DLPFC counts for both cohorts")
  storeData( "METADATA", "Merged metadata used here for both cohorts")
  storeData( "isexpr", "Genes used in this analysis")
  storeData( "vobj.lst", "Voom results for each cohort")
  storeData( "resid.add", "Residuals correcting for covariates and add back - (Dx_Sex Dx Sex)")
  storeData( "fit_marginal.lst", "dream fit for marginal analysis")
  storeData( "res_zenith", "zenith results from fit_marginal.lst")
  storeData( "res_meta_zenith", "Meta-analysis of zenith results from fit_marginal.lst")
  storeData( "res_meta_zenith.sex_autosome", "Meta-analysis of zenith results from fit_marginal.lst - just autosomal genes")
  storeData( "df_DE_results", "Differential expressed genes within each cohort")
  storeData( "df_meta", "Meta-analysis of differential expression across cohorts")

  # Sex by Disease interaction
  storeData( "fit_inter_sex.disease.lst", "dream fit for interaction analysis")
  storeData( "res_zenith.inter_sex.disease", "zenith results for interaction analysis")
  storeData( "res_meta_zenith.inter_sex.disease", "Meta-analysis of zenith results for interaction analysis")
  storeData( "res_meta_zenith.inter_sex.disease_autosome", "Meta-analysis of zenith results for interaction analysis -  autosomal genes only")
  storeData( "df_DE_results.inter_sex.disease", "Differential expression result for each cohort for interaction analysis")
  storeData( "df_meta.inter_sex.disease", "Meta-analysis for differential expression for interaction analysis")
}

#   rmarkdown::render("combined_analysis/combine_all_DLPFC.Rmd")

```

### Source
[markdown](`r thisFile`)