### 2020-09-23 combined_analysis/plot_diff.networks.Rmd
Initial commit

---
title: "CommonMind Consortium: DLPFC"
subtitle: "Plots for Differential network analysis - DLPFC"
author:
- name: "[Gabriel Hoffman](http://gabrielhoffman.github.io)$^{1}$ and Kelsey Montgomery$^{2}$"
  affiliation: | 
    [1] Icahn School of Medicine at Mount Sinai, New York
    [2] Sage Bionetworks, Seattle, WA
date: "Run on `r Sys.time()`"
documentclass: article
output:
  html_document:
    toc: true
    toc_float: true
upload: FALSE
---

<!---

cd /hpc/users/hoffmg01/work/covarr-de
ml git pandoc udunits proj gdal geos
git pull
R

system("git pull")

rmarkdown::render("combined_analysis/plot_diff.networks.Rmd")

--->

```{r setup, include=FALSE}
suppressPackageStartupMessages({
library(synapser)

synLogin()

library(WGCNA)
library(variancePartition)
library(decorate)
library(ggrepel)
library(cowplot) 
library(metap) 
library(zenith) 
library(knitr) 
library(plyr) 
library(githubr)
library(qvalue)  
library(BiocParallel)  
})   
source("../common_functions/common_functions.R")   
# source("common_functions/common_functions.R")

knitr::opts_chunk$set(
  echo = FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,   
  tidy = FALSE,
  cache = TRUE,
  cache.lazy = FALSE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE, eval=FALSE}

parentId = 'syn22416298';
activityName = 'Plots for Differential network analysis - DLPFC';
activityDescription = 'Plots for Differential network analysis - DLPFC';

thisFileName <- 'diff.networks.Rmd'

# Github link
thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('combined_analysis/',thisFileName))
```

```{r load.data}
# Load data
resid.add = readRDS(synGet('syn22416489')$path)
vobj.lst = readRDS(synGet('syn22416446')$path)
METADATA = readRDS(synGet('syn22416411')$path)
ALL_USED_IDs = c('syn22416489', 'syn22416446', 'syn22416411')

METADATA$Reported_Gender = factor(METADATA$Reported_Gender, c("Male", "Female"))
```

```{r load.dream}
fit_marginal.lst = readRDS(synGet('syn22416500')$path)
fit_inter_sex.disease.lst = readRDS(synGet('syn22416524')$path)
df_meta = readRDS(synGet('syn22416508')$path)
df_meta.inter_sex.disease = readRDS(synGet('syn22416532')$path)

ALL_USED_IDs = c(ALL_USED_IDs, c('syn22416500', 'syn22416524', 'syn22416508', 'syn22416532'))
```