### 2020-10-14 combined_analysis/CTS/combine.Rmd
initial commit

---
title: "Combine cell type specific analysis"
subtitle: ""
author:
- name: "[Gabriel Hoffman](http://gabrielhoffman.github.io)"
  affiliation: | 
    Icahn School of Medicine at Mount Sinai, New York
date: "Run on `r Sys.time()`"
documentclass: article
output:
  html_document:
    toc: true
    toc_float: true
---

<!---

cd /hpc/users/hoffmg01/work/covarr-de
# rm -rf combined_analysis/combine_all_DLPFC_cache
ml git pandoc
git pull
R

system("git pull")

rmarkdown::render("combined_analysis/CTS/combine.Rmd")

--->

```{r synapse.parameters, include=FALSE, cache=TRUE, eval=FALSE}

library(githubr)
parentId = 'syn23241817';
activityName = "Combine cell type specific analysis"
activityDescription = "Combine cell type specific analysis"

thisFileName <- 'combine.Rmd'

# Github link
thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('combined_analysis/CTS/',thisFileName))
```

```{r setup, include=FALSE}
suppressPackageStartupMessages({
library(synapser)

synLogin()

library(ggplot2)
library(dplyr)
library(ComplexHeatmap)
library(readxl)
library(gtools)
library(corrplot)
library(pvclust)
library(metafor)
library(data.table)
library(cowplot)
library(ggraph)
library(tidygraph)
library(kableExtra)
library(ashr)
library(githubr)
library(ggrepel)
library(corrplot)
library(zenith)
library(compositions)
library(knitr)
library(patchwork)
library(plyr) 
library(BiocParallel)

library(limma)
library(edgeR)
library(variancePartition)

}) 
source("../../common_functions/common_functions.R")   
# source("common_functions/common_functions.R")

knitr::opts_chunk$set(
  echo = FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  dev = c('png', 'pdf'),
  cache = TRUE)
```