### 2021-01-08 combined_analysis/combine_all_DLPFC.Rmd
add new plots

---
title: "CommonMind Consortium: DLPFC"
subtitle: "Analysis of read count data by cohort"
author:
- name: "[Gabriel Hoffman](http://gabrielhoffman.github.io)$^{1}$ and Kelsey Montgomery$^{2}$"
  affiliation: | 
    [1] Icahn School of Medicine at Mount Sinai, New York
    [2] Sage Bionetworks, Seattle, WA
date: "Run on `r Sys.time()`"
documentclass: article
output:
  html_document:
    toc: true
    toc_float: true
params:
  upload: FALSE
---

<!---

cd /hpc/users/hoffmg01/work/covarr-de
# rm -rf combined_analysis/combine_all_DLPFC_cache
ml git pandoc
git pull
R

system("git pull")

rmarkdown::render("combined_analysis/combine_all_DLPFC.Rmd")

--->

```{r knit2synapse, eval= FALSE}
library(synapser)
library(knit2synapse)
synLogin()
knit2synapse::createAndKnitToFolderEntity(file = "combined_analysis/combine_all_DLPFC.Rmd",
                                          parentId ="syn22416298",
                                          folderName = 'Joint analysis of DLPFC count data',
                                          overwrite=TRUE, 
                                          knitmd=TRUE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}

library(githubr)
parentId = 'syn22416298';
activityName = 'Joint analysis of DLPFC count data';
activityDescription = 'Joint analysis of DLPFC count data';

thisFileName <- 'combine_all_DLPFC.Rmd'

# Github link
thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('combined_analysis/',thisFileName))
```

<!---

TODO

zenith 1 sided

zenith plots
replication or correlation as a function of expression magnitude

difference networks
-------------------

DONE

Effect of ilr should depend on Instaition
suffix 1.
zenith for Sex with autosomes
check zenith squared test statistics

--->

```{r setup, include=FALSE}
suppressPackageStartupMessages({
library(synapser)

synLogin()

library(ggplot2)
library(dplyr)
library(ComplexHeatmap)
library(readxl)
library(gtools)
library(corrplot)
library(fdrtool)
library(pvclust)
library(RColorBrewer)
library(metafor)
library(data.table)
library(cowplot)
library(ggraph)
library(tidygraph)
library(kableExtra)
library(githubr)
library(ggrepel)
library(corpcor)
library(corrplot)
library(zenith)
library(compositions)
library(knitr)
library(patchwork)
library(plyr) 
library(BiocParallel)

library(limma)
library(edgeR)
library(variancePartition)

}) 
source("../common_functions/common_functions.R")   
# source("common_functions/common_functions.R")

knitr::opts_chunk$set(
  echo = FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  dev = c('png', 'pdf'),
  cache = TRUE)

colorArray = colors()[grep("grey|gray", colors(), invert=TRUE)]

```

```{r read.counts}
# Read raw counts from MPP and HBCC
counts.MPP = fread(synGet('syn21867938')$path, data.table=FALSE)
rownames(counts.MPP) = counts.MPP$Geneid
counts.MPP = counts.MPP[,-c(1:6)]

# HBCC - DLFPC
counts.HBCC = fread(synGet('syn21886235')$path, data.table=FALSE)
rownames(counts.HBCC) = counts.HBCC$Geneid
counts.HBCC = counts.HBCC[,-c(1:6)]

# # HBCC - sgACC
counts.HBCC4 = fread(synGet('syn21867866')$path, data.table=FALSE)
rownames(counts.HBCC4) = counts.HBCC4$Geneid
counts.HBCC4 = counts.HBCC4[,-c(1:6)]

geneCountsMerged = cbind(counts.MPP, counts.HBCC, counts.HBCC4)
```

### Data download
#### Obtain effective count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
downloadFile <- function(id){
  fread(synGet(id)$path, data.table = F)
}
downloadFile_version <- function(id , version){
  fread(synGet(id, version = version)$path, data.table = F)
}

# Get RNASeq QCmetadata
METADATA_QC_DLPFC_ID = 'syn16816488' 
ALL_USED_IDs = METADATA_QC_DLPFC_ID
metadata = downloadFile_version(METADATA_QC_DLPFC_ID, version = 19) %>%
  dplyr::select(`Individual_ID`, Institution, Cohort, `Reported_Gender`, Sex, Ethnicity, ageOfDeath, `PMI_(in_hours)`, Dx, 'pH',
                Sample_RNA_ID, one_of('rnaSeq_isolation:RIN',
                                      'rnaSeq_dissection:Brain_Region', 
                                      'rnaSeq_isolation:RNA_Isolation_Batch',
                                      'rnaSeq_report:Ribozero_Batch', 'rnaSeq_report:Library_Batch', 
                                      'rnaSeq_report:Flowcell_Batch', 'rnaSeq_report:Exclude?',
                                      'rnaSeq_report:Mapped_Reads',
                                      'rnaSeq_report:Intragenic_Rate','rnaSeq_report:Intronic_Rate',
                                      'rnaSeq_report:Intergenic_Rate','rnaSeq_report:Genes_Detected',
                                      'rnaSeq_report:Expression_Profiling_Efficiency',
                                      'rnaSeq_report:rRNA_Rate',
                                      'rnaSeq_report:Total_Reads', 'rnaSeq_report:Percent_Aligned','rnaSeq_report:Transcripts_Detected'))

# Get ancestry vector calculated using gemtools
ANCESTRY_ID = 'syn9922992'
ALL_USED_IDs = c(ALL_USED_IDs, ANCESTRY_ID)
ANCESTRY.HBCC = downloadFile(ANCESTRY_ID) %>% 
  plyr::rename(c(ID = 'SNP_report:Genotyping_Sample_ID'))

ANCESTRY_ID = 'syn2511399'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = ANCESTRY_ID
ANCESTRY.MPP = downloadFile(ANCESTRY_ID) %>% 
  plyr::rename(c('DNA_report..Genotyping.Sample_ID' = 'SNP_report:Genotyping_Sample_ID'))

ANCESTRY = rbind(ANCESTRY.HBCC[,colnames(ANCESTRY.MPP)[-2]], ANCESTRY.MPP[,colnames(ANCESTRY.MPP)[-2]])

# Get genotype ids from synapse. 
GENOTYPE_ID = 'syn16816490'
ALL_USED_IDs = c(ALL_USED_IDs, GENOTYPE_ID)
GENOTYPE = downloadFile(GENOTYPE_ID) %>% 
  dplyr::select(Individual_ID, `SNP_report:Genotyping_Sample_ID`, `SNP_report:Exclude?`) %>% 
  dplyr::inner_join(ANCESTRY) %>% 
  dplyr::filter(is.na(`SNP_report:Exclude?`))
```

### Data preprocessing
```{r preprocess.data}
# Merge metadata 
METADATA = metadata %>%
  dplyr::left_join(GENOTYPE) %>%      
  dplyr::rename(Region = `rnaSeq_dissection:Brain_Region`,
                PMI = `PMI_(in_hours)`,
                RIN = `rnaSeq_isolation:RIN`, 
                ReportExclude = `rnaSeq_report:Exclude?`,
                GenotypeExclude = `SNP_report:Exclude?`,
                SampleID = Sample_RNA_ID,
                LibraryBatch = `rnaSeq_report:Library_Batch`,
                FlowcellBatch = `rnaSeq_report:Flowcell_Batch`,
                IsolationBatch = `rnaSeq_isolation:RNA_Isolation_Batch`, #isolation batch to replace ribozero batch
                RibozeroBatch = `rnaSeq_report:Ribozero_Batch`,
                MappedReads = `rnaSeq_report:Mapped_Reads`,
                IntragenicRate = `rnaSeq_report:Intragenic_Rate`, IntronicRate = `rnaSeq_report:Intronic_Rate`, 
                IntergenicRate = `rnaSeq_report:Intergenic_Rate`, GenesDetected = `rnaSeq_report:Genes_Detected`,
                ExpProfEfficiency = `rnaSeq_report:Expression_Profiling_Efficiency`, rRNARate = `rnaSeq_report:rRNA_Rate`,
                TotalReads = `rnaSeq_report:Total_Reads`, AlignmentRate = `rnaSeq_report:Percent_Aligned`, TranscriptsDetected = `rnaSeq_report:Transcripts_Detected`) %>%
  dplyr::select(SampleID, Individual_ID, Institution, Cohort, Reported_Gender,Sex, Ethnicity, ageOfDeath, PMI, Dx, Region,
                RIN, ReportExclude, GenotypeExclude, EV.1, EV.2, EV.3, EV.4, EV.5, pH,
                LibraryBatch, FlowcellBatch, IsolationBatch, RibozeroBatch, MappedReads, TotalReads, GenesDetected, AlignmentRate,
                IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate, TranscriptsDetected) %>%
  		 		 dplyr::filter(SampleID %in% colnames(geneCountsMerged), !is.na(SampleID)) %>%
  		 		 dplyr::filter(Region %in% c('DLPFC')) %>%
			 	 dplyr::mutate(Cohort = forcats::fct_recode(Cohort, `MSSM-Penn-Pitt` = 'MSSM-Penn-Pitt', `NIMH-HBCC`='NIMH-HBCC', `NIMH-HBCC`='NIMH-HBCC-4')) %>%
			 	 dplyr::filter(Dx %in% c('Control', "SCZ"))  %>%
         dplyr::filter(ageOfDeath > 17)

# Set Control to baseline
METADATA$Dx = factor(METADATA$Dx, c('Control', 'SCZ'))

ind = METADATA$SampleID[which(METADATA$ReportExclude == 1 | METADATA$GenotypeExclude)]
writeLines(paste('Following',length(ind),'samples are marked exclude'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$Ethnicity) | is.na(METADATA$Institution) | is.na(METADATA$Dx)]
writeLines(paste('Following', length(ind), 'counts are missing any metadata'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$PMI)]
writeLines(paste('Following', length(ind), 'counts are missing PMI'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$Reported_Gender)]
writeLines(paste('Following', length(ind), 'counts are missing gender'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$ageOfDeath)]
writeLines(paste('Following', length(ind), 'counts are missing age of death'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

ind = METADATA$SampleID [is.na(METADATA$EV.1)]
writeLines(paste('Following', length(ind), 'counts are missing ancestry information'))
writeLines(paste(ind, collapse = ', '))
METADATA <- METADATA  %>% dplyr::filter(!(SampleID %in% ind))

# drop individuals who where sequenced twice here
tab = table(METADATA$Individual_ID)

dropSamples = sapply( names(tab[tab > 1]), function(id){

  idx = which(METADATA$Individual_ID == id)
  df = METADATA[idx,]

  # remove individuals that have less then max reads
  i = which(df$MappedReads < max(df$MappedReads))
  rownames(METADATA)[idx[i]]
  } )
METADATA = METADATA[!(rownames(METADATA) %in% dropSamples),]
```

```{r cellFractionEstimates}
# include estimated cell fractions in METADATA
df_cellFractions = read.table(synGet('syn22333694')$path, row.names=1)

celFrac_ilr = ilr(df_cellFractions)
colnames(celFrac_ilr) = paste0("cellFrac_ilr_", 1:3)

METADATA = merge(METADATA, df_cellFractions, by.x = 'SampleID', by.y='row.names')
METADATA = merge(METADATA, celFrac_ilr, by.x = 'SampleID', by.y='row.names')
```

```{r preprocess.data1, results='asis'}
# Match covariates to expression data
indToRetain = intersect(METADATA$SampleID, colnames(geneCountsMerged))

geneCountsMerged = geneCountsMerged[,indToRetain]

rownames(METADATA) = METADATA$SampleID
```

## Summarize samples
```{r summary}
METADATA %>% group_by(Dx, Institution) %>% dplyr::summarise(count = n()) %>% 
  kable(caption="Diagnosis, Institution") %>% kable_styling(full_width=FALSE)

METADATA %>% group_by( Institution, Reported_Gender) %>% dplyr::summarise(count = n()) %>% 
  kable(caption="Institution, Sex") %>% kable_styling(full_width=FALSE)

METADATA %>% group_by(Dx, Institution, Reported_Gender) %>% dplyr::summarise(count = n()) %>%  
kable(caption="Diagnosis, Institution, Sex") %>% kable_styling(full_width=FALSE) 
```

```{r summarize.plots}

METADATA %>% group_by(Dx, Institution, Reported_Gender) %>% dplyr::summarise(count = n()) %>% ggplot(aes(Institution, count, fill=Dx)) + geom_bar(stat="identity", position=position_dodge()) + facet_wrap(~ Reported_Gender) + theme_bw(16) + theme(aspect.ratio=1, axis.text.x = element_text(angle = 45, hjust=1)) + scale_fill_manual("Diagnosis", values=c("grey", "darkred"))
```

```{r summarize.plots2, fig.height=12}
variables = c( 'ageOfDeath',"RIN", "PMI", 'rRNARate', 'IntronicRate', 'IntragenicRate', 'IntergenicRate', "pH")

METADATA %>% group_by(Dx, Institution, Reported_Gender) %>% reshape2::melt(, id.vars=c('Dx', 'Institution', 'Reported_Gender'), measure.vars=variables ) %>% ggplot(aes(Institution, value, fill=Dx)) + geom_boxplot() + facet_grid( variable ~ Reported_Gender, scales='free_y') + theme_bw(16) + theme(aspect.ratio=1, axis.text.x = element_text(angle = 45, hjust=1), legend.position="none") + scale_fill_manual("Diagnosis", values=c("grey", "darkred"))
```

Wilcoxon test comparing Controls and SZ within each Institution and Sex
```{r compare.case.control, fig.height=12}
df = METADATA %>% group_by(Dx, Institution, Reported_Gender) %>% reshape2::melt(, id.vars=c('Dx', 'Institution', 'Reported_Gender'), measure.vars=variables ) %>% data.table

# test difference bween CTL and SZ within each Institution, Sex
df[,wilcox.test(value ~ Dx)$p.value, by=c("Institution", "Reported_Gender", "variable")] %>% ggplot(aes(Institution, -log10(V1))) + geom_bar(stat="identity", position=position_dodge()) + facet_grid(variable ~ Reported_Gender) + theme_bw(16) + theme(aspect.ratio=1, axis.text.x = element_text(angle = 45, hjust=1)) + scale_fill_manual("Diagnosis", values=c("grey", "darkred")) + ylab(bquote(-log[10]~P))
```

ANCOVA evaluating heterogenity in values across cohorts for each sex
```{r ancova, fig.height=12}
df2 = df[,anova(lm(value ~ Institution))$`Pr(>F)`[1], by=c("Reported_Gender", "variable")] 
df2[,p.adj := p.adjust(V1, "bonferroni")]

ggplot(df2, aes(Reported_Gender, -log10(V1), fill=p.adj < 0.05)) + geom_bar(stat="identity", position=position_dodge()) + facet_wrap(~variable, ncol=1) + theme_bw(16) + theme(aspect.ratio=1, axis.text.x = element_text(angle = 45, hjust=1)) + ylab(bquote(-log[10]~P)) + scale_fill_manual("Passes 5%\nBonferroni", values=c("grey", "navy"))

df2[,p := format(V1, digits=2,scientific=TRUE)][,c('Reported_Gender', 'variable', 'p')] %>% kable(caption="ANCOVA p-values", scientific=TRUE) %>% kable_styling(full_width=FALSE)
```

```{r joint.cca, fig.width=7, fig.height=7}

# pH is not included because it is not available om many of the samples
sum(!is.na(METADATA$pH))
sum(is.na(METADATA$pH))

variables = c('Reported_Gender', 'CellType', 'ageOfDeath', 'Institution', "Dx")
variables2 = c("RIN", "PMI", 'rRNARate', 'IntronicRate', 'IntragenicRate', 'IntergenicRate')
variables = c(variables, variables2)

df_cca = lapply(variables, function(v1){

  if( v1 == "CellType"){
    form1 = ~ GABA + GLU + Olig + MgAs
  }else{
    form1 = as.formula(paste0('~', v1))
  }
  M1 = model.matrix(form1, METADATA)

  res = lapply(variables, function(v2){

    if( v2 == "CellType"){
      form2 = ~ GABA + GLU + Olig + MgAs
    }else{
      form2 = as.formula(paste0('~', v2))
    }
    M2 = model.matrix(form2, METADATA)

    # If some rows are dropped for including NA's, 
    # only analyze rows present in both variables
    rowids = intersect(rownames(M1), rownames(M2))

    fit = cancor(M1[rowids,], M2[rowids,])
    V_sqrt = sqrt(mean(fit$cor^2)) # Cramer's sqrt V

    data.frame(v1, v2, V_sqrt)
  })
  do.call(rbind, res)
})
df_cca = do.call(rbind, df_cca)

C = matrix(df_cca$V_sqrt, ncol=length(variables))
rownames(C) = variables
colnames(C) = variables

plotCorrMatrix(C, main='CCA')
```

<!---
Evaluate and perform hypothesis test for partial correlations
```{r pcor}
# Compute partial correlations
pC = cor2pcor(C)
colnames(pC) = colnames(C)
rownames(pC) = rownames(C)

res = lapply(colnames(C), function(v1){
  res = lapply(colnames(C), function(v2){

    r = pC[v1, v2]
    n = nrow(METADATA)
    k = ncol(C)
    t = r * sqrt((n-2-k) / (1-r^2))

    p = 2*pt( abs(t), n - 2 -k, lower.tail=FALSE)

    a = sort(c(v1, v2))

    data.frame(V1 = a[1], V2 = a[2], partial.corr = r, pvalue = p)
  })
  do.call(rbind, res)
 })
res = data.table(do.call(rbind, res))
res = res[V1 != V2,]
res[,title:=paste(V1, V2)]
res = res[,.SD[1,], by='title']

res$p.adj = p.adjust( res$pvalue, "BH")

ggplot(res, aes(V1, V2, fill=partial.corr, label=ifelse(p.adj<0.05, '*', ''))) + geom_tile() + theme_bw() + scale_fill_gradient2(name = "Partial corr", low="blue", mid="white", high="red", limit=c(-1, 1)) + theme(aspect.ratio=1, panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), legend.position="bottom") + geom_text() + coord_flip() + xlab("Variable 1") + ylab("Variable 2")
```
--->

```{r explain.CellType}
CTs = c('GABA', 'GLU', 'Olig', 'MgAs')

form = ~ ageOfDeath:Institution + Reported_Gender + Dx + Institution + RIN + rRNARate + PMI + IntronicRate + IntragenicRate + IntergenicRate

Y = clr(METADATA[,CTs])

df = lapply( colnames(Y), function(id){

  form2 = as.formula(paste0('Y[,id] ~ ', as.character(form)[2]))

  fit = lm(form2, METADATA)

  vp = calcVarPart(fit)
  n = length(vp) - 1 # omit residuals

  data.frame( Test     = rownames(anova(fit))[1:n],
              CellType = id, 
              p        = anova(fit)[,'Pr(>F)'][1:n], 
              frac     =  calcVarPart(fit)[1:n])
})
df = data.table(do.call(rbind, df))

df$Test[df$Test=='Reported_Gender'] = "Sex"
df$Test[df$Test=='ageOfDeath:Institution'] = "Age of death"
df$p.adj = p.adjust(df$p, "bonferroni")

df2 = df[,sum(frac), by="Test"]
setorder(df2, 'V1')

df$Test = factor(df$Test, df2$Test)

ymax = max(df[,sum(frac),by="CellType"]$V1)*105
ggplot(df, aes(CellType, frac*100, fill=Test)) + geom_bar(stat="identity") + coord_flip() + theme_bw() + theme(aspect.ratio=1, panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +  ylab("Variance explained (%)") + theme(axis.line = element_line(colour = "transparent"), 
        axis.line.x = element_line(colour = "black"), panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), panel.border = element_blank(), 
        panel.background = element_blank(), axis.ticks.y = element_blank(), 
        legend.key = element_blank(), plot.margin = unit(c(0, 
            0.3, 0, 0.8), "cm")) + scale_y_continuous(expand=c(0,0), limits=c(0, ymax)) + xlab('') + scale_fill_manual(values = rev(brewer.pal( length(table(df$Test)), "Set3")))

ratio = length(CTs) / nlevels(df$Test)

ggplot(df, aes(CellType, Test, fill=-log10(pmax(1e-16,p)), label=ifelse(p.adj<0.05, '*', ''))) + geom_tile() + theme_bw() + scale_fill_gradient(name = bquote(-log[10]~P), low="white", high="red", limit=c(0, NA)) + theme(aspect.ratio=ratio, panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), legend.position="bottom") + geom_text() + coord_flip()
```

```{r sex.graph, fig.width=7, fig.height=7, dev="png"}
plotStarCorr = function(df_cca, cutoff=1e-3){

  df_cca2 = df_cca[with(df_cca, v1!=v2),]
  df_cca2$key = apply(df_cca2, 1, function(x){
    x = sort(c(x['v1'], x['v2']))
    paste(x[1], x[2])
    }) 
  df_cca2 = df_cca2[!duplicated(df_cca2$key),]

  df_cca2 = df_cca2[df_cca2$V_sqrt^2 > cutoff,]

  rstat_nodes <- data.frame(name = unique(c(df_cca2$v1, df_cca2$v2)))

  rstat_edges <- data.frame(from = match(df_cca2$v1, rstat_nodes$name),
                           to = match(df_cca2$v2, rstat_nodes$name),
                           Rsq = df_cca2$V_sqrt^2 )

  net = tbl_graph(nodes = rstat_nodes, edges = rstat_edges, directed=FALSE)

  ggraph(net,  layout = 'igraph', algorithm = 'star') + 
    geom_edge_link(aes(width = Rsq, color=Rsq), alpha = 1) +
    geom_node_point(size=20, color="grey70") +
    geom_node_text(aes(label = name), repel = FALSE) +
    theme_graph() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) +
    scale_edge_width(name=bquote(R^2)) +
    scale_edge_color_gradient(name=bquote(R^2), low="cornsilk", high="red", lim=c(0, max(df_cca2$V_sqrt^2))) +  ggtitle("Pairwise squared correlation")
}

plotStarCorr(df_cca, 0)
```

```{r sex.graph.sub, fig.width=7, fig.height=7, dev="png"}
i1 = with(df_cca, v1 %in% setdiff(variables, variables2))
i2 = with(df_cca, v2 %in% setdiff(variables, variables2))

plotStarCorr(df_cca[i1 & i2,], 0)
```

```{r sex.confounding, fig.width=5, fig.height=5}
form = ~ Reported_Gender + GABA + GLU + Olig + MgAs + ageOfDeath + PMI + rRNARate + RIN + IntronicRate + IntragenicRate + IntergenicRate
par(mfrow=c(2,2))
res = lapply( unique(METADATA$Institution), function(INST){
  i = METADATA$Institution == INST
  C = canCorPairs(form, METADATA[i,]) 
  plotCorrMatrix(C, main=INST)
})
```

Despite the fact that the estimated GABA fraction is lower in females, this observation is not significant after accounting for other variables, notably age of death. 
```{r sex.confounding.testing, fig.width=7, fig.height=4}

res = lapply( unique(METADATA$Institution), function(INST){
  res = lapply( c('GABA', 'GLU', 'Olig', 'MgAs'), function(CT){
    i = METADATA$Institution == INST
    y = factor(METADATA[i,]$Reported_Gender, c("Male", "Female"))

    form = as.formula(paste0('y ~ ', CT, '+ ageOfDeath + PMI + rRNARate + RIN + IntronicRate + IntragenicRate + IntergenicRate'))

    fit = glm(form,  METADATA[i,], family="binomial")

    result = data.frame(coef(summary(fit))[-1,])
    result$coef = rownames(result)
    result$Institution = INST
    result$CellType = CT
    rownames(result) = NULL

    result
    })
  do.call(rbind, res)
})
res = do.call(rbind, res)
colnames(res)[colnames(res) == 'Pr...z..'] = 'pvalue'
colnames(res)[colnames(res) == 'Std..Error'] = 'se'
res$p.adj = p.adjust(res$pvalue, "BH")

lim = max(abs(res$z.value))

res2 = res[res$coef %in%  c('GABA', 'GLU', 'Olig', 'MgAs'),]
res2$p.adj = p.adjust(res2$pvalue, "BH")

ggplot(res2, aes(Institution, Estimate, color=coef)) + geom_point() + theme_bw() + ylab("Regression coefficient") + facet_wrap(~CellType, nrow=1) + theme(aspect.ratio=1) + geom_errorbar(aes(ymin = Estimate - se, ymax = Estimate + se), width=0.1) + geom_hline(yintercept=0, linetype="dashed")
```

```{r significance.CT.diff}
colorCodes = c(
  GABA = '#66A61E',
  GLU = '#E6AB02',
  Olig = '#E7298A',
  MgAs = '#7570B3',
  Bulk = "grey50")

res2$color = colorCodes[match(res2$coef, names(colorCodes))]

ggplot(res2, aes(Institution, -log10(pvalue))) + geom_point(color=res2$color) + theme_bw() + ylab("Regression coefficient") + facet_wrap(~CellType, nrow=1) + theme(aspect.ratio=1, axis.text.x = element_text(angle = 45, hjust=1)) + geom_hline(yintercept=-log10(0.05/nrow(res2)), linetype="dashed") + ylab(bquote(-log[10]~P)) + ylim(0, 3)
```

Sow results for hypothesis testing for: `ageOfDeath + PMI + rRNARate + RIN`

```{r sex.confounding.testing2, fig.width=7, fig.height=4}

res = lapply( unique(METADATA$Institution), function(INST){
  i = METADATA$Institution == INST
  y = factor(METADATA[i,]$Reported_Gender, c("Male", "Female"))

  form = y ~ ageOfDeath + PMI + rRNARate + RIN + IntronicRate + IntragenicRate + IntergenicRate

  fit = glm(form,  METADATA[i,], family="binomial")

  result = data.frame(coef(summary(fit))[-1,])
  result$coef = rownames(result)
  result$Institution = INST
  rownames(result) = NULL

  result
})
res = do.call(rbind, res)
colnames(res)[colnames(res) == 'Pr...z..'] = 'pvalue'
res$p.adj = p.adjust(res$pvalue, "BH")

ggplot(res, aes(Institution, z.value, fill=coef)) + geom_bar(stat="identity", position = position_dodge())  + theme_bw() + ylab("Z-score from logistic regression") + theme(aspect.ratio=1)
```

Evaluating each term in the model: `Reported_Gender ~ Institution*(ageOfDeath + PMI + rRNARate + RIN + IntronicRate + IntragenicRate + IntergenicRate)` using logistic regression
```{r anova.test}
# run ANOVA LRT omitting each variable
y = factor(METADATA$Reported_Gender, c("Male", "Female"))

form = y ~ Institution*(ageOfDeath + PMI + rRNARate + RIN + IntronicRate + IntragenicRate + IntergenicRate)
fit_full = glm(form,  METADATA, family="binomial")

res = lapply(c('ageOfDeath', 'PMI', 'rRNARate','RIN', 'IntronicRate', 'IntragenicRate', 'IntergenicRate'), function(V){

  form2 = gsub(V, "1", form)
  form2 = as.formula(paste(form2[2], '~', form2[3]))

  fit_null = glm(form2,  METADATA, family="binomial")

  res = anova(fit_full, fit_null, test="Chisq")

  data.frame(Variable = V, pvalue = res$`Pr(>Chi)`[2])
})
res = do.call(rbind, res)
res$p.adj = p.adjust(res$pvalue, "bonferroni")

res %>% kable(caption="Hypothesis test ", digits=8) %>% kable_styling(full_width=FALSE)
```

```{r sex.by.age}
ggplot(METADATA, aes(Reported_Gender, ageOfDeath, fill=Reported_Gender)) + geom_boxplot() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + facet_wrap(~Institution, nrow=1) + scale_fill_manual(values = c("red", "blue"))
```

```{r fraction.female}
df = xtabs(~Reported_Gender + Institution, METADATA)
df2 = apply(df, 2, function(x) x / sum(x))

plotPercentBars(data.frame(t(df2))) + scale_fill_manual(values=c("red", "blue")) + ylab("Fraction of samples") + xlab("Institution")
```

### Sex by cell fraction
```{r summary.sex, fig.width=8, fig.height=10}
figList = lapply(colnames(df_cellFractions), function(CT){
  ggplot(METADATA, aes_string('Reported_Gender', CT, fill='Reported_Gender')) + geom_boxplot() + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1) + scale_fill_manual(values = c("red", "blue"))
})
plot_grid(plotlist=figList, ncol=1) 
```

### Age by cell fraction
```{r summary.Age, fig.width=8, fig.height=10}
figList = lapply(colnames(df_cellFractions), function(CT){
  ggplot(METADATA, aes_string('ageOfDeath', CT)) + geom_point() + geom_smooth(method="gam") + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1)
}) 
plot_grid(plotlist=figList, ncol=1) 
```

## Age by cell fraction log ratio
```{r summary.Age.log.ratio, fig.width=8, fig.height=8}
figList = lapply(colnames(celFrac_ilr), function(CT){
  ggplot(METADATA, aes_string('ageOfDeath', CT)) + geom_point() + geom_smooth(method="gam") + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + facet_wrap(~Institution, nrow=1)
})
plot_grid(plotlist=figList, ncol=1) 
```

```{r process.counts}
# keep genes that are express in at least one combination of cohort and gender
isexpr.lst = dlply(METADATA, .(Cohort), .fun = function(mtd){
	rowSums(cpm(geneCountsMerged[,mtd$SampleID])>.5) >= .3*ncol(geneCountsMerged[,mtd$SampleID])
	})

isexpr = do.call(rbind, isexpr.lst)
isexpr = colSums(isexpr) == 2 # keep gene expressed in both cohorts 
table(isexpr)

# Normalize each cohort separately
dge.lst = lapply( levels(METADATA$Cohort), function(Cohort){
  # normalize counts
  dge = DGEList( geneCountsMerged[isexpr,METADATA$Cohort==Cohort] )
  calcNormFactors(dge, "none") # TMM
})
names(dge.lst) = levels(METADATA$Cohort)

# form = ~ (1|Dx) + Reported_Gender + RIN + ageOfDeath + scale(IntronicRate)*Cohort + scale(IntragenicRate)*Cohort + scale(IntergenicRate)*Cohort + scale(rRNARate)*Cohort + (1|Institution:Cohort) + (1|Region:Cohort) + (1|Individual_ID)

form.lst = list(
  `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
  `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)

vobj.lst = lapply(dge.lst, function(dge){

  i = match(colnames(dge), rownames(METADATA))
  info = METADATA[i,]

  form_use = form.lst[[info$Cohort[1]]]

  voomWithDreamWeights(dge, form_use, info, save.plot=TRUE, BPPARAM = SnowParam(18))
  })
names(vobj.lst) = names(dge.lst)
```

### Voom trend plots
```{r plot.voom, fig.height=5, fig.width=10}
fig1 = plot_voom(vobj.lst[[1]]) + ggtitle(names(vobj.lst)[1])
fig2 = plot_voom(vobj.lst[[2]]) + ggtitle(names(vobj.lst)[2])
plot_grid( fig1, fig2, nrow=1)

df = lapply( names(vobj.lst), function(key){
  data.frame(Cohort = key,  vobj.lst[[key]]$voom.line)
  })
df = do.call(rbind, df)

ggplot(df, aes(x, y, color=Cohort)) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + geom_line(size=1.2) + xlab(bquote(log[2](count + 0.5))) + ylab(expression( sqrt("standard deviation"))) + ggtitle("Compare voom trends") + ylim(0, max(df$y)) + scale_color_manual(name="Status", values=c("#3758ba", "#33e65c")) + guides(color = guide_legend(override.aes = list(size = 1))) 
```

<!---

# Examine genes on chrX and chrY

# dge = DGEList(geneCountsMerged)
# dge = calcNormFactors(dge, "none")
# geneExpr = cpm(dge, log=TRUE)

# df = merge( data.frame(Expression = geneExpr['ENSG00000229807.11',]), METADATA, by="row.names")
# ggplot(df, aes(Reported_Gender, Expression)) + geom_violin() + theme_bw() + facet_wrap(~Cohort)

# id = 'ENSG00000229807.11'
# isexpr[id]

# df = as.data.frame(vobj.lst[[2]]$voom.xy)
# with(df, plot(x,y, cex=.2, pch=20))
# with(df[id,], points(x,y, col="red", pch=20, cex=2))

--->

```{r Create.residuals}

### Create residuals
form.lst = list(
  `MSSM-Penn-Pitt`  = ~ Dx*Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
  `NIMH-HBCC`       = ~ Dx*Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)

eval_residuals = function( form ){
  # Compute residuals, but add back in Dx*Reported_Gender
  # for form = ~ Dx*Reported_Gender
  f = function(fit){
    residuals(fit) + get_prediction( fit, form)
  }

  residuals.lst = lapply( vobj.lst, function(vobj){

    i = match(colnames(vobj), rownames(METADATA))
    info = METADATA[i,]

    form_use = form.lst[[info$Cohort[1]]]

    # Compute residuals which adding back Diagnosis and Sex effects
    resid.lst = fitVarPartModel(vobj, form_use, info, fxn = f, BPPARAM = SnowParam(18))  
    do.call(rbind, resid.lst)
    })
  names(residuals.lst) = names(vobj.lst)

  residuals.lst
}

resid.add = list( Dx_Sex  = eval_residuals( ~ Dx*Reported_Gender ),
                  Dx      = eval_residuals( ~ Dx ),
                  Sex     = eval_residuals( ~ Reported_Gender ))

residuals.lst = resid.add[['Dx_Sex']]
```

```{r check.correlation, eval=FALSE}
ngenes = 300
C1 = cor(t(residuals.lst[[1]][1:ngenes,]))
C2 = cor(t(residuals.lst[[2]][1:ngenes,]))

df = data.table(A   = as.numeric(C1[lower.tri(C1)]),
                B   = as.numeric(C2[lower.tri(C2)]))

df$density <- get_density(df$A, df$B, n = 80)
ggplot(df, aes(A, B, color=density)) + geom_point() + geom_point(size=.4) + theme_bw(15) + theme(aspect.ratio=1, legend.position="bottom", plot.title = element_text(hjust = 0.5)) + geom_abline(color="red") + scale_color_viridis() + xlim(-1,1) + ylim(-1,1) + xlab(paste0('Correlation (', names(residuals.lst)[1], ')')) + ylab(paste0('Correlation (', names(residuals.lst)[2], ')'))
```

# Schizophrenia vs controls

### Test marginals for Scz vs controls
```{r differential.expression.simple}
form.lst = list(
  `MSSM-Penn-Pitt`  = ~ Dx + Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
  `NIMH-HBCC`       = ~ Dx + Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)

fit_marginal.lst = lapply( vobj.lst, function(vobj){

  i = match(colnames(vobj), rownames(METADATA))
  info = METADATA[i,]

  form_use = form.lst[[info$Cohort[1]]]

  fit = dream(vobj, form_use, info, BPPARAM = SnowParam(18))  
  eBayes( fit )
  })
names(fit_marginal.lst) = names(vobj.lst)

coefTest = c('DxSCZ', 'Reported_GenderMale', 'ageOfDeath')
```

```{r load.genesets, eval=TRUE}
# Load Gene Ontology database 
gs.GO = get_GeneOntology()
gs.MSigDB = get_MSigDB()

# convert from GeneSetCollection to list used by camera and zenith
gs.GO.collection = recodeToList( gs.GO )
gs.MSigDB.collection = recodeToList( gs.MSigDB )
```

```{r run.zenith, eval=TRUE}
res_zenith = lapply( names(fit_marginal.lst), function(key){

  fit = fit_marginal.lst[[key]]

  res.GO = run_zenith(  fit, 
                        coefTest,
                        gs.GO.collection)
  res.GO$Resource = "GO"   
  res.GO$Cohort = key

  res.MSigDB = run_zenith(fit, 
                          coefTest,
                          gs.MSigDB.collection ) 
  res.MSigDB$Resource = "MSigDB"
  res.MSigDB$Cohort = key

  rbind(res.GO, res.MSigDB)
})
names(res_zenith) = names(fit_marginal.lst)

# only autosomes
res_zenith_sex_autosomes = lapply( names(fit_marginal.lst), function(key){

  fit = fit_marginal.lst[[key]]

  # only keep autosomes, for Reported_GenderMale
  tab = coef(fit) %>% getGeneSymbol()
  geneids = rownames(tab)[tab$Chrom %in% 1:22]
  fit = fit[geneids,]

  res.GO = run_zenith(  fit, 
                        'Reported_GenderMale',
                        gs.GO.collection)
  res.GO$Resource = "GO"   
  res.GO$Cohort = key

  res.MSigDB = run_zenith(fit, 
                          'Reported_GenderMale',
                          gs.MSigDB.collection ) 
  res.MSigDB$Resource = "MSigDB"
  res.MSigDB$Cohort = key

  rbind(res.GO, res.MSigDB)
})
names(res_zenith_sex_autosomes) = names(fit_marginal.lst)
```

## Meta-analysis of zenith results
```{r zenith.meta, eval=TRUE}
res_meta_zenith = zenith_meta_analysis( res_zenith )                  
res_meta_zenith.sex_autosome = zenith_meta_analysis( res_zenith_sex_autosomes )
```

```{r get.signatures, fig.width=10, fig.height=5}
# Get DE signatures
df_DE_results = lapply(fit_marginal.lst, function(fit){
  res = lapply(coefTest, function(coef){
    res = topTable(fit, coef, number=Inf) 
    res$Coef = coef
    res %>% getGeneSymbol()  
    })
  names(res) = coefTest
  res
  })
names(df_DE_results) = names(fit_marginal.lst)
```

```{r test.scz, fig.height=5, fig.width=10}
# Schizophrenia testing
# extract from each cohort
coef = 'DxSCZ'
# ymin = min(sapply(df_DE_results, function(x) min(x[[coef]]$P.Value)))

figList.Scz = lapply( names(df_DE_results), function(key){
  df = df_DE_results[[key]][[coef]]

  main = paste(key, coef, sep = ': ' )
  plotVolcano( df, df$Symbol[1:10] ) + ggtitle( main ) #+ ylim(0, -log10(ymin))
})

# volcano plots
plot_grid(plotlist=figList.Scz, nrow=1)
```

## Meta-analysis across cohorts
```{r meta.analysis.DE}
# Store results for each coef and cohort
tab.de.lst = lapply(coefTest, function(coef){    
  tab.lst = lapply( df_DE_results, function(lst){
    lst[[coef]]
  })
  names(tab.lst) = names(df_DE_results)
  tab.lst
  })
names(tab.de.lst) = coefTest

# Create list preparing for meta-analysis for SCZ
df_meta = lapply(coefTest, function(coef){
  # Run meta-analysis
  run_meta_analysis( tab.de.lst[[coef]], "FE") %>% getGeneSymbol
})#, mc.cores=6)
names(df_meta) = coefTest
```

```{r concordance.scz, fig.height=6, fig.width=10}
# Concordance analysis
resConcord.SCZ = plot_concordance(tab.de.lst[[coef]], showGenes = df_meta[[coef]]$Symbol[1:20])

main = "Concordance of SCZ"
plot_grid(  resConcord.SCZ$fig_logFC + ggtitle(main),
            resConcord.SCZ$fig_tstat + ggtitle(main), nrow=1)

data.frame(Cohort = names(tab.de.lst[[coef]]),
          "# DE"  = resConcord.SCZ$n.de,
          pi1     = resConcord.SCZ$pi_single,
          check.names=FALSE) %>% 
          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)

resConcord.SCZ$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
resConcord.SCZ$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
```

```{r plot.meta.analysis.SCZ, fig.height=5, fig.width=10}
# volcano plot for meta-analysis
main = paste('Meta-analysis', coef, sep = ': ' )
plotVolcano( df_meta[[coef]], df_meta[[coef]]$Symbol[1:20] ) + ggtitle( main )

data.frame(pi1 = 1 - qvalue(df_meta[[coef]]$P)$pi0,
          `# DE` = sum(df_meta[[coef]]$adj.P.Val < 0.05),
          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for SCZ meta-analysis") %>% kable_styling(full_width=FALSE)
```

## Plot top examples
```{r plot.forrest.dlpfc, fig.width=8, fig.height=20}  
plot_gene = function(residuals.lst, df_meta, geneSymbol, variable, fill=c("#1E90FF", "#FF8D1E"), mode=""){

  ensGene = rownames(df_meta)[which(df_meta$Symbol == geneSymbol)]

  exprResid = lapply( names(residuals.lst), function(key){
      data.frame(Expression = residuals.lst[[key]][ensGene,])
    })
  exprResid = do.call(rbind, exprResid)

  df = merge(exprResid, METADATA, by="row.names")

  if( is.factor(df[[variable]]) || length(unique(df[[variable]])) < 5 ){
    fig = ggplot(df, aes_string(variable, 'Expression')) + geom_violin(aes_string(fill=variable)) + geom_boxplot(width=0.1, fill="grey80") + scale_fill_manual(values = fill)
  }else{    
    fig = ggplot(df, aes_string(variable, 'Expression')) + geom_point() + geom_smooth(method="lm") #+ scale_color_manual(values = fill)
  }

  fig + theme_bw() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + ylab('Expression residuals') + ggtitle(geneSymbol) + facet_wrap(~Cohort)
}

coef = 'DxSCZ'
figList = lapply(df_meta[[coef]]$Symbol[1:10], function(geneSymbol){
  fig1 = plot_forrest( c(tab.de.lst[[coef]], list("Meta-analysis" = df_meta[[coef]])), geneSymbol)

  fig2 = plot_gene(residuals.lst, df_meta[[coef]], geneSymbol, "Dx") + xlab("Diagnosis")

  plot_grid(fig1, fig2, nrow=1, rel_widths=c(.8, 1))
})

plot_grid(plotlist=figList, ncol=1)
```

### Concordance with other datasets
```{r CMC.concordance}
# CMC Analysis 1
CMC1 <- readr::read_tsv(synGet("syn5607603")$path) %>% 
  dplyr::rename(gene_id = genes,
                `CMC 1 t-statistic` = t) %>% 
  dplyr::select(gene_id, `CMC 1 t-statistic`)
# HBCC-DLPFC microarray
CMC_HBCC <- readr::read_tsv(synGet("syn4941553")$path) %>% 
  dplyr::rename(gene_id = genes,
                `HBCC microarray t-statistic` = t) %>% 
  dplyr::select(gene_id, `HBCC microarray t-statistic`)
# Gandal, Science 2018, Table S1
# Use Tab:DGE
# wget https://science.sciencemag.org/highwire/filestream/720110/field_highwire_adjunct_files/1/aat8127_Table_S1.xlsx
gandal <- readxl::read_xlsx( synGet("syn22255217")$path, sheet = "DGE") %>% 
  dplyr::rename(gene_id = ensembl_gene_id,
                `Gandal, SCZ t-statistic` = `SCZ.t.value`,
                `Gandal, BD t-statistic` = `BD.t.value`,
                `Gandal, ASD t-statistic` = `ASD.t.value`) %>%
  dplyr::select(gene_id, `Gandal, SCZ t-statistic`,
                `Gandal, BD t-statistic`, `Gandal, ASD t-statistic`)
# AMP-AD RNASeq reprocessing GRCh37 - https://www.synapse.org/#!Synapse:syn9702085
MSSM <- readr::read_tsv(synGet("syn10526259")$path) %>%
  dplyr::filter(Comparison %in% "AD-CONTROL",
                Model %in% "Diagnosis") %>%
  dplyr::select(ensembl_gene_id, Tissue, t) %>%
  dplyr::group_by(ensembl_gene_id, Tissue) %>% 
  dplyr::filter(row_number() == 1) %>% 
  tidyr::pivot_wider(names_from = Tissue,
                     values_from = t) %>%
  dplyr::rename(`AMP-AD MSSM FP AD t-statistic` = FP,
                `AMP-AD MSSM IGF AD t-statistic` = IFG,
                `AMP-AD MSSM PHG AD t-statistic` = PHG,
                `AMP-AD MSSM STG AD t-statistic` = STG,
                gene_id = ensembl_gene_id) %>% 
  dplyr::select(gene_id, `AMP-AD MSSM STG AD t-statistic`) 
MSSM <- MSSM[!is.na(MSSM$`AMP-AD MSSM STG AD t-statistic`),]

MAYO <- readr::read_tsv(synGet("syn8468023")$path) %>% 
  dplyr::filter(Comparison %in% "AD-CONTROL",
                Model %in% "Diagnosis") %>%
  dplyr::select(ensembl_gene_id, `Tissue.ref`, t) %>%
  dplyr::group_by(ensembl_gene_id, `Tissue.ref`) %>% 
  dplyr::filter(row_number() == 1) %>% 
  tidyr::pivot_wider(names_from = `Tissue.ref`,
                     values_from = t) %>% 
  dplyr::rename(`AMP-AD MAYO CBE AD t-statistic` = CBE,
                `AMP-AD MAYO TCX AD t-statistic` = TCX,
                gene_id = ensembl_gene_id) %>% 
  dplyr::select(-`AMP-AD MAYO TCX AD t-statistic`)

ROSMAP <- readr::read_tsv(synGet("syn8456721")$path) %>%
    dplyr::filter(Comparison %in% "AD-CONTROL",
                Model %in% "Diagnosis") %>%
  dplyr::rename(gene_id = ensembl_gene_id, 
                `AMP-AD ROSMAP DLPFC AD t-statistic` = t) %>% 
  dplyr::select(gene_id, `AMP-AD ROSMAP DLPFC AD t-statistic`)

# out_group <- inner_join(MSSM, MAYO) %>% 
#   inner_join(., ROSMAP) %>% 
#   inner_join(., gandal)

out_group <- inner_join(gandal, ROSMAP)

# merge with meta-analysis

coef = 'DxSCZ'
df = df_meta[[coef]]
df_two = data.frame(gene_id = trim_ensembl_ids(rownames(df)),
                `CMC - meta-analysis` = with(df, logFC/se), check.names=FALSE)

df = tab.de.lst[[coef]][[1]]
df.tmp = data.frame(gene_id = trim_ensembl_ids(rownames(df)),
                `t` = df$t, check.names=FALSE)
colnames(df.tmp)[2] = names(tab.de.lst[[coef]])[1]
df_two = df_two %>% inner_join(., df.tmp)

df = tab.de.lst[[coef]][[2]]
df.tmp = data.frame(gene_id = trim_ensembl_ids(rownames(df)),
                `t` = df$t, check.names=FALSE)
colnames(df.tmp)[2] = names(tab.de.lst[[coef]])[2]
df_two = df_two %>% inner_join(., df.tmp)

# Remove transcript Ids
df_two <- df_two %>% 
  tidyr::separate(gene_id, c("gene_id", "transcript")) %>% 
  dplyr::select(-one_of("transcript"))
df_two <- inner_join(df_two, CMC1, by = c("gene_id")) 
df_two <- inner_join(df_two, CMC_HBCC, by = c("gene_id"))
df_two <- inner_join(df_two, out_group, by = c("gene_id")) %>% 
  group_by(gene_id) %>% 
  dplyr::filter(row_number()==1) %>% 
  tibble::column_to_rownames(var = "gene_id")

colnames(df_two) <- gsub(" t-statistic", "", colnames(df_two))

col2 <- colorRampPalette(rev(c("#67001F", "#B2182B", "#D6604D", "#F4A582",
                           "#FDDBC7", "#FFFFFF", "#D1E5F0", "#92C5DE",
                           "#4393C3", "#2166AC", "#053061")))

cor <- cor(df_two, method="spearman")
# clustering based on distance
hcl = hclust(as.dist(1-cor), method='ward.D2')

results <- pvclust(cor, method.hclust='ward.D2')
# reorder based on clustering
reorder = cor[results$hclust$order, results$hclust$order]

# plot clustering
plot(results, hang=-1)
# plot(hcl, hang=-1)
corrplot(reorder, col = col2(200))
```

## t-statistics for CASKIN1
```{r CASKIN1}
 df_two['ENSG00000167971',]  %>% kable(caption="t-statistics for CASKIN1", row.names=FALSE) %>% kable_styling(full_width=FALSE)
```

```{r show.gene_set_test.marginal, eval=TRUE}
show_zenith_results = function( res, coef, Resource, FDR = 0.05){

  i = which((res$Resource == Resource) & (res$Coef == coef))
  df = res[i,]

	# filter by FDR
	df = df[df$FDR <= FDR,]
  df$Coef = c()
  df$se = c()
  df$Resource = c()

	df %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption=paste(Resource, coef), row.names=FALSE) %>% kable_styling(full_width=FALSE)
}

show_zenith_results( res_meta_zenith, "DxSCZ", "GO", FDR=0.05)
show_zenith_results( res_meta_zenith, "DxSCZ", "MSigDB", FDR=0.05)
```

```{r plot.gene_set_test.marginal, eval=TRUE}

  plot_zenith_results = function( res, coef, Resource, FDR = 0.05, compare=NULL){

  i = which((res$Resource == Resource) & (res$Coef == coef))
  df = res[i,]

  # filter by FDR
  df$FDR = pmax(1e-300, df$FDR)
  df = df[df$FDR <= FDR,]
  df$Coef = c()
  df$se = c()
  df$Resource = c()

  df = df[order(df$PValue, decreasing=TRUE),]
  df$Set = factor(df$Set, df$Set)
  ymax = max(-log10(df$FDR)) * 1.05

  if( !is.null(compare) ){
    df$Direction = compare[(sign(df$effect) + 1)/2 + 1]
    fig = ggplot(df, aes(Set, -log10(FDR), fill = Direction )) + geom_bar(stat="identity") + scale_fill_manual(name = "Upregulated in", values = c('red', 'blue'))
  }else{
    fig = ggplot(df, aes(Set, -log10(FDR) )) + geom_bar(stat="identity", fill="navy")
  }

  fig + theme_bw() + theme(aspect.ratio=1,plot.title = element_text(hjust = 0.5)) + coord_flip() + ylab(bquote(-log[10]~FDR)) + xlab('') + ggtitle(coef) + scale_y_continuous(expand=c(0,0), limits=c(0, ymax)) + geom_hline(yintercept=-log10(0.05), color="black", linetype="dashed") 
}

 plot_zenith_results( res_meta_zenith, "DxSCZ", "GO", FDR=0.05) 
```

```{r, fig.width=9}
  plot_zenith_results( res_meta_zenith, "DxSCZ", "MSigDB", FDR=0.05)  
```

# Sex Analysis
```{r test.Sex, fig.height=5, fig.width=10}
# Schizophrenia testing
# extract from each chort
coef = 'Reported_GenderMale'
figList.Sex = lapply( names(df_DE_results), function(key){
  df = df_DE_results[[key]][[coef]]
  main = paste(key, coef, sep = ': ' )
  plotVolcanoSex( df, df$Symbol[1:20] ) + ggtitle( main )
})

# volcano plots
plot_grid(plotlist=figList.Sex, nrow=1)
```

### Autosomes
```{r test.Sex.autosome, fig.height=5, fig.width=10}
# Schizophrenia testing
# extract from each chort
coef = 'Reported_GenderMale'
figList.Sex = lapply( names(df_DE_results), function(key){
  df = df_DE_results[[key]][[coef]]
  main = paste(key, coef, sep = ': ' )
  df = df[df$Chrom %in% 1:22,]
  plotVolcano( df, df$Symbol[1:20] ) + ggtitle( main )
})

# volcano plots
plot_grid(plotlist=figList.Sex, nrow=1)
```

## Meta-analysis across cohorts
```{r concordance.Sex, fig.height=6, fig.width=10}
# Concordance analysis
resConcord.Sex = plot_concordance(tab.de.lst[[coef]], showGenes = df_meta[[coef]]$Symbol[1:20])

main = "Concordance of Sex"
plot_grid(  resConcord.Sex$fig_logFC + ggtitle(main),
            resConcord.Sex$fig_tstat + ggtitle(main), nrow=1)

data.frame(Cohort = names(tab.de.lst[[coef]]),
          "# DE"  = resConcord.Sex$n.de,
          pi1     = resConcord.Sex$pi_single,
          check.names=FALSE) %>% 
          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)

resConcord.Sex$tab_pi  %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
resConcord.Sex$tab_corr  %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
```

```{r concordance.Sex.autosome, fig.height=6, fig.width=10}

tab.de.lst.autosome = lapply(tab.de.lst[[coef]], function(x) x[x$Chrom %in% 1:22,])

# Concordance analysis
df = df_meta[[coef]][df_meta[[coef]]$Chrom %in% 1:22,]
resConcord.Sex = plot_concordance(tab.de.lst.autosome, showGenes = df$Symbol[1:20])

main = "Concordance of Sex - autosomes"
plot_grid(  resConcord.Sex$fig_logFC + ggtitle(main),
            resConcord.Sex$fig_tstat + ggtitle(main), nrow=1)

data.frame(Cohort = names(resConcord.Sex$pi_single),
          "# DE"  = resConcord.Sex$n.de,
          pi1     = resConcord.Sex$pi_single,
          check.names=FALSE) %>% 
          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)

resConcord.Sex$tab_pi %>% mutate_if(is.numeric, format, digits=3)  %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
resConcord.Sex$tab_corr  %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
```

```{r plot.meta.analysis.Sex, fig.height=5, fig.width=10}
# volcano plot for meta-analysis
main = 'Meta-analysis: Sex'
fig1 = plotVolcanoSex( df_meta[[coef]], df_meta[[coef]]$Symbol[1:15] ) + ggtitle( main )

df = df_meta[[coef]][df_meta[[coef]]$Chrom %in% 1:22,]
fig2 = plotVolcano( df, df$Symbol[1:20] ) + ggtitle( paste(main, "autosome") )
plot_grid(fig1, fig2, ncol=2)

data.frame(pi1 = 1 - qvalue(df_meta[[coef]]$P)$pi0,
          `# DE` = sum(df_meta[[coef]]$adj.P.Val < 0.05),
          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for Sex meta-analysis") %>% kable_styling(full_width=FALSE)

data.frame(pi1 = 1 - qvalue(df$P)$pi0,
          `# DE` = sum(df$adj.P.Val < 0.05),
          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for Sex meta-analysis - autosomes") %>% kable_styling(full_width=FALSE)
```

## Sex circos
```{r sex.circos}
coef = 'Reported_GenderMale'
sapply( names(df_DE_results), function(key){
  plot_circos_logFC( df_DE_results[[key]][[coef]] ) + ggtitle(key)
  })

plot_circos_logFC( df_meta[[coef]] ) + ggtitle("Meta analysis")
```

## Plot top examples - not chrY
```{r plot.forrest.dlpfc.Sex.autosomes, fig.width=8, fig.height=20}  
Symbols = df_meta[[coef]][df_meta[[coef]]$Chrom !='Y',]$Symbol[1:10]
figList = lapply(Symbols, function(geneSymbol){

  fig1 = plot_forrest( c(tab.de.lst[[coef]], list("Meta-analysis" = df_meta[[coef]])), geneSymbol)

  fig2 = plot_gene( residuals.lst, df_meta[[coef]], geneSymbol, "Reported_Gender", fill=c("red", "blue")) + xlab("Sex")

  plot_grid(fig1, fig2, nrow=1)
})

plot_grid(plotlist=figList, ncol=1)
```

## Plot top examples - autosomes
```{r plot.forrest.dlpfc.Sex, fig.width=8, fig.height=20}  
Symbols = df_meta[[coef]][df_meta[[coef]]$Chrom %in% 1:22,]$Symbol[1:10]

figList = lapply(Symbols, function(geneSymbol){

  fig1 = plot_forrest( c(tab.de.lst[[coef]], list("Meta-analysis" = df_meta[[coef]])), geneSymbol)

  fig2 = plot_gene( residuals.lst, df_meta[[coef]], geneSymbol, "Reported_Gender", fill=c("red", "blue")) + xlab("Sex")

  plot_grid(fig1, fig2, nrow=1)
})

plot_grid(plotlist=figList, ncol=1)
```

## Meta-analysis of zenith results
```{r show.gene_set_test.marginal.Sex, eval=TRUE}
show_zenith_results( res_meta_zenith, "Reported_GenderMale", "GO", FDR=0.05)         
show_zenith_results( res_meta_zenith, "Reported_GenderMale", "MSigDB", FDR=0.05)        
```

Evaluate if GABA pathways are different between sexes
```{r zenith.gaba}
df = data.table(res_meta_zenith)
df = df[Coef == 'Reported_GenderMale',] 
df2 = df[grep("GABA", Set),]

# correct only for number of tests that have GABA in the name
df2[,FDR := p.adjust(PValue, "fdr")] 
df2 %>% kable(caption="Test only gene sets with GABA in the name") %>% kable_styling(full_width=FALSE)
```

```{r plot.gene_set_test.marginal.Sex, eval=TRUE, fig.width=12}
  plot_zenith_results( res_meta_zenith, "Reported_GenderMale", "GO", FDR=0.05, compare=c('Female', "Male"))               
  plot_zenith_results( res_meta_zenith, "Reported_GenderMale", "MSigDB", FDR=0.05, compare=c('Female', "Male"))         
```

Show genes in neurexin gene set
```{r neurexin}
geneNames = gs.GO.collection[[grep('GO0042043', names(gs.GO.collection))]]

df = df_meta[['Reported_GenderMale']]

i = match( geneNames, trim_ensembl_ids(rownames(df)))

df = df[i[!is.na(i)],]
df$ischrY = df$Chrom == "Y"

figList = lapply( unique(df$ischrY), function(TF){

  df2 = df[df$ischrY == TF,]
  ggplot(df2, aes(Symbol, logFC, color=adj.P.Val < 0.05)) + geom_point() + geom_errorbar(aes(ymin = logFC - se, ymax = logFC + se, width=.1)) + theme_bw(16) + theme(aspect.ratio=1) + geom_hline(yintercept=0, linetype="dashed") + coord_flip() + scale_color_manual(name="FDR < 5%", values=c("black", "red"))
  })
plot_grid(plotlist = figList, ncol=1)
```

### Only autosomes
```{r show.gene_set_test.marginal.Sex.autosome, eval=TRUE}
show_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "GO", FDR=0.1) 
show_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "MSigDB", FDR=0.1)   
```

```{r plot.gene_set_test.marginal.Sex.autosome, eval=TRUE}
  plot_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "GO", FDR=0.1, compare=c('Female', "Male"))  
  plot_zenith_results( res_meta_zenith.sex_autosome, "Reported_GenderMale", "MSigDB", FDR=0.1, compare=c('Female', "Male"))    
```

# Age Analysis
```{r test.Age, fig.height=5, fig.width=10}
# Schizophrenia testing
# extract from each chort
coef = 'ageOfDeath'
figList.Age = lapply( names(df_DE_results), function(key){
  df = df_DE_results[[key]][[coef]]
  main = paste(key, coef, sep = ': ' )
  plotVolcano( df, df$Symbol[1:20] ) + ggtitle( main )
})

# volcano plots
plot_grid(plotlist=figList.Age, nrow=1)
```

## Meta-analysis across cohorts
```{r concordance.Age, fig.height=6, fig.width=10}
# Concordance analysis
resConcord.Age = plot_concordance(tab.de.lst[[coef]], showGenes = df_meta[[coef]]$Symbol[1:20])

main = "Concordance of Age"
plot_grid(  resConcord.Age$fig_logFC + ggtitle(main),
            resConcord.Age$fig_tstat + ggtitle(main), nrow=1)

data.frame(Cohort = names(tab.de.lst[[coef]]),
          "# DE"  = resConcord.Age$n.de,
          pi1     = resConcord.Age$pi_single,
          check.names=FALSE) %>% 
          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)

resConcord.Age$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
resConcord.Age$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
```

```{r plot.meta.analysis.Age, fig.height=5, fig.width=10}
# volcano plot for meta-analysis
main = 'Meta-analysis: Age'
plotVolcano( df_meta[[coef]], df_meta[[coef]]$Symbol[1:20] ) + ggtitle( main )

data.frame(pi1 = 1 - qvalue(df_meta[[coef]]$P)$pi0,
          `# DE` = sum(df_meta[[coef]]$adj.P.Val < 0.05),
          check.names=FALSE ) %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for Age meta-analysis") %>% kable_styling(full_width=FALSE)
```

## Plot top examples
```{r plot.forrest.dlpfc.Age, fig.width=8, fig.height=20}  
figList = lapply(df_meta[[coef]]$Symbol[1:10], function(geneSymbol){

  fig1 = plot_forrest( c(tab.de.lst[[coef]], list("Meta-analysis" = df_meta[[coef]])), 
    geneSymbol)

  expr.list = lapply(vobj.lst, function(x)x$E)

  fig2 = plot_gene( expr.list, df_meta[[coef]], geneSymbol, "ageOfDeath", fill=c("red", "blue")) + xlab("Age")

  plot_grid(fig1, fig2, nrow=1)
})

plot_grid(plotlist=figList, ncol=1)
```

## Meta-analysis of zenith results
```{r show.gene_set_test.marginal.Age, eval=TRUE}
show_zenith_results( res_meta_zenith, "ageOfDeath", "GO", FDR=0.01)   
show_zenith_results( res_meta_zenith, "ageOfDeath", "MSigDB", FDR=0.01)    
```

```{r plot.gene_set_test.marginal.Age, eval=TRUE}
  plot_zenith_results( res_meta_zenith, "ageOfDeath", "GO", FDR=0.01)    
  plot_zenith_results( res_meta_zenith, "ageOfDeath", "MSigDB", FDR=0.01)    
```

# Comparison with GTEx
```{r gtex.sex, fig.height=6, fig.width=10}

# read beta and se from GTEx
df_gtex.beta = read.table(synGet("syn22891761")$path, header=TRUE, row.names=1)
df_gtex.se = read.table(synGet("syn22891767")$path, header=TRUE, row.names=1)
df_gtex.lfsr = read.table(synGet("syn22891755")$path, header=TRUE, row.names=1)
df_signif = fread(synGet("syn22891771")$path) 
df_signif = df_signif[tissue =='Brain_Frontal_Cortex_BA9', ]
# identical(rownames(df_gtex.beta), rownames(df_gtex.se))
# identical(rownames(df_gtex.lfsr), rownames(df_gtex.se))

# data.frame of GTEx results
df_gtex_BA9 = data.frame( ENSEMBL.full = rownames(df_gtex.beta),
                          ENSEMBL = trim_ensembl_ids( rownames(df_gtex.beta) ),
                          logFC = -1*df_gtex.beta[,'Brain_Frontal_Cortex_BA9'],
                          se = df_gtex.se[,'Brain_Frontal_Cortex_BA9'],
                          adj.P.Val = df_gtex.lfsr[,'Brain_Frontal_Cortex_BA9'])

df_gtex_BA9$isSignif = FALSE
df_gtex_BA9$isSignif[ df_gtex_BA9$ENSEMBL.full %in% df_signif$gene] = TRUE

df_gtex_BA9 = df_gtex_BA9 %>% getGeneSymbol( "ENSEMBL")

df_gtex_BA9$t = with(df_gtex_BA9, logFC / se)
df_gtex_BA9$P.Value = with(df_gtex_BA9, 2*pnorm(abs(t), lower.tail=FALSE))
# df_gtex_BA9$adj.P.Val = p.adjust(df_gtex_BA9$P.Value, "fdr")
df_gtex_BA9 = df_gtex_BA9[!is.na(df_gtex_BA9$logFC),]
rownames(df_gtex_BA9) = df_gtex_BA9$ENSEMBL

# data.frame of CMC results from meta-analysis
df = df_meta[['Reported_GenderMale']]
df$t = with(df, logFC / se)
rownames(df) = trim_ensembl_ids( rownames(df) )

# all genes
resConcord.gtex = plot_concordance( list('CMC v3' = df, 'GTEx BA9' = df_gtex_BA9), showGenes=df$Symbol[1:10] )

main = "Concordance with GTEx (all genes)"
plot_grid(  resConcord.gtex$fig_logFC + ggtitle(main),
            resConcord.gtex$fig_tstat + ggtitle(main), nrow=1)

# resConcord.gtex$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
resConcord.gtex$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)

# because GTEx uses MASH to combine 44 tissue, standard plot_concordance() does give 
# correct replication values
df.merge = merge(df, df_gtex_BA9, by="row.names")

table(df.merge$isSignif)

df_repl = data.frame(Discovery = c("GTEx", "CMC"), rate = c(NA, NA))

# Discover in CMC
# with(df.merge, table(isSignif[adj.P.Val.x < 0.05]))
# with(df, 1 - qvalue(P.Value.y[adj.P.Val.x < 0.05])$pi0)
lfsr = with(df.merge, adj.P.Val.y[adj.P.Val.x < 0.05]) 
df_repl$rate[1] = sum(1-lfsr) / length(lfsr)

# Discover in GTEx
# with(df, 1 - qvalue(P.Value.x[adj.P.Val.y < 0.05])$pi0)
# df_repl$rate[2] = with(df.merge, 1 - qvalue(P.Value.x[isSignif])$pi0)

res = with(df.merge, ash( logFC.x, se.x))
df.merge$lsfr = res$result$lfsr

lfsr = with(df.merge, lsfr[adj.P.Val.x < 0.05]) 
df_repl$rate[2] = sum(1-lfsr) / length(lfsr)

df_repl %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)

# how any DEGs?
# sum(apply(df_gtex.lfsr, 1, function(x) any(x <= 0.05, na.rm=TRUE)))

# df_combine = apply(df_gtex.lfsr, 2, function(x) sum(x <= 0.05, na.rm=TRUE))
# df_combine = data.frame(Tissue = names(df_combine), Count = df_combine)
# df_combine$Tissue = factor(df_combine$Tissue, rev(df_combine$Tissue))

# ggplot(df_combine, aes(Tissue, Count)) + geom_bar(stat="identity") + theme_bw() + coord_flip()

# # All genes: 
# sum(df_gtex.lfsr[,'Brain_Frontal_Cortex_BA9'] < 0.05, na.rm=TRUE)
# sum(df_gtex.lfsr[,'Brain_Frontal_Cortex_BA9'] < 0.05, na.rm=TRUE)

# # expressed in CMC
# sum(df_gtex_BA9$adj.P.Val < 0.05, na.rm=TRUE)

# sum( df$adj.P.Val.y < 0.05)

# # get significant genes directly
# df_combine = df_signif[,data.frame(Count = length(gene)), by="tissue"]
# df_combine$tissue = factor(df_combine$tissue, rev(df_combine$tissue))

# ggplot(df_combine, aes(tissue, Count)) + geom_bar(stat="identity") + theme_bw() + coord_flip()

```

```{r gtex.sex.2, fig.height=6, fig.width=10}
# autosomes
df = df[df$Chrom %in% 1:22,]

resConcord.gtex = plot_concordance( list('CMC v3' = df, 'GTEx BA9' = df_gtex_BA9), showGenes=df$Symbol[1:10] )

main = "Concordance with GTEx (autosomes)"
plot_grid(  resConcord.gtex$fig_logFC + ggtitle(main),
            resConcord.gtex$fig_tstat + ggtitle(main), nrow=1)

# resConcord.gtex$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
resConcord.gtex$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)

df.merge = merge(df, df_gtex_BA9, by="row.names")

table(df.merge$isSignif)

df_repl = data.frame(Discovery = c("GTEx", "CMC"), rate = c(NA, NA))

# Discover in CMC
# with(df.merge, table(isSignif[adj.P.Val.x < 0.05]))
# with(df, 1 - qvalue(P.Value.y[adj.P.Val.x < 0.05])$pi0)
lfsr = with(df.merge, adj.P.Val.y[adj.P.Val.x < 0.05]) 
df_repl$rate[1] = sum(1-lfsr) / length(lfsr)

# Discover in GTEx
# with(df, 1 - qvalue(P.Value.x[adj.P.Val.y < 0.05])$pi0)
# df_repl$rate[2] = with(df.merge, 1 - qvalue(P.Value.x[isSignif])$pi0)

res = with(df.merge, ash( logFC.x, se.x))
df.merge$lsfr = res$result$lfsr

lfsr = with(df.merge, lsfr[adj.P.Val.x < 0.05]) 
df_repl$rate[2] = sum(1-lfsr) / length(lfsr)

df_repl %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)
```

# Sex by Disease interaction
```{r dream.sex.disease}
form.lst = list(
  `MSSM-Penn-Pitt`  = ~ Dx*Reported_Gender + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
  `NIMH-HBCC`       = ~ Dx*Reported_Gender + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)

# drop chrY genes
df_genes  = data.frame(ENSEMBL = rownames(vobj.lst[[1]])) %>% getGeneSymbol('ENSEMBL')
keep.ensID = df_genes$ENSEMBL[df_genes$Chrom %in% c(1:22, 'X', 'MT')]

fit_inter_sex.disease.lst = lapply( vobj.lst, function(vobj){

  i = match(colnames(vobj), rownames(METADATA))
  info = METADATA[i,]

  form_use = form.lst[[info$Cohort[1]]]

  fit = dream(vobj[keep.ensID,], form_use, info, BPPARAM = SnowParam(18))  
  eBayes( fit )
  })
names(fit_inter_sex.disease.lst) = names(vobj.lst)

coefTest = c('DxSCZ:Reported_GenderMale')
```

```{r run.zenith.inter_sex.disease, eval=TRUE}
res_zenith.inter_sex.disease = lapply( names(fit_inter_sex.disease.lst), function(key){

  fit = fit_inter_sex.disease.lst[[key]]

  res.GO = run_zenith(  fit,   
                        coefTest,
                        gs.GO.collection)
  res.GO$Resource = "GO"   
  res.GO$Cohort = key

  res.MSigDB = run_zenith(fit, 
                          coefTest,
                          gs.MSigDB.collection) 
  res.MSigDB$Resource = "MSigDB"
  res.MSigDB$Cohort = key

  rbind(res.GO, res.MSigDB)
})
names(res_zenith.inter_sex.disease) = names(fit_inter_sex.disease.lst)

# only autosomes
res_zenith.inter_sex.disease_autosome = lapply( names(fit_inter_sex.disease.lst), function(key){

  fit = fit_inter_sex.disease.lst[[key]]

  # only keep autosomes, for Reported_GenderMale
  tab = coef(fit) %>% getGeneSymbol()
  geneids = rownames(tab)[tab$Chrom %in% 1:22]
  fit = fit[geneids,]

  res.GO = run_zenith(  fit, 
                        coefTest,
                        gs.GO.collection)
  res.GO$Resource = "GO"   
  res.GO$Cohort = key

  res.MSigDB = run_zenith(fit, 
                          coefTest,
                          gs.MSigDB.collection) 
  res.MSigDB$Resource = "MSigDB"
  res.MSigDB$Cohort = key

  rbind(res.GO, res.MSigDB)
})
names(res_zenith.inter_sex.disease_autosome) = names(fit_inter_sex.disease.lst)
```

## Meta-analysis of zenith results
```{r zenith.meta.inter_sex.disease, eval=TRUE}
res_meta_zenith.inter_sex.disease = zenith_meta_analysis( res_zenith.inter_sex.disease )  
res_meta_zenith.inter_sex.disease_autosome = zenith_meta_analysis( res_zenith.inter_sex.disease_autosome )  
```

```{r get.signatures.inter_sex.disease, fig.width=10, fig.height=5}
# Get DE signatures
df_DE_results.inter_sex.disease = lapply(fit_inter_sex.disease.lst, function(fit){
  res = lapply(coefTest, function(coef){
    res = topTable(fit, coef, number=Inf) 
    res$Coef = coef
    res %>% getGeneSymbol()  
    })
  names(res) = coefTest 
  res
  })
names(df_DE_results.inter_sex.disease) = names(fit_inter_sex.disease.lst)
```

```{r test.scz.inter_sex.disease, fig.height=5, fig.width=10}
coef = 'DxSCZ:Reported_GenderMale'

figList.Scz = lapply( names(df_DE_results), function(key){
  df = df_DE_results.inter_sex.disease[[key]][[coef]]

  main = paste(key, coef, sep = ': ' )
  plotVolcano( df, df$Symbol[1:10] ) + ggtitle( main )
})

# volcano plots
plot_grid(plotlist=figList.Scz, nrow=1)
```

## Meta-analysis across cohorts
```{r meta.analysis.DE.inter_sex.disease}
# Store results for each coef and cohort
tab.de.inter_sex.disease.lst = lapply(coefTest, function(coef){    
  tab.de.inter_sex.disease.lst = lapply( df_DE_results.inter_sex.disease, function(lst){
    lst[[coef]]
  })
  names(tab.de.inter_sex.disease.lst) = names(df_DE_results.inter_sex.disease)
  tab.de.inter_sex.disease.lst
  })
names(tab.de.inter_sex.disease.lst) = coefTest

# Create list preparing for meta-analysis for SCZ
df_meta.inter_sex.disease = lapply(coefTest, function(coef){
  # Run meta-analysis
  run_meta_analysis( tab.de.inter_sex.disease.lst[[coef]], "FE") %>% getGeneSymbol
}) 
names(df_meta.inter_sex.disease) = coefTest  
```

```{r concordance.scz.inter_sex.disease, fig.height=6, fig.width=10}
# Concordance analysis.  
resConcord.inter_sex.disease = plot_concordance(tab.de.inter_sex.disease.lst[[coef]], showGenes = df_meta.inter_sex.disease[[coef]]$Symbol[1:20])

main = "Concordance of DxSCZ:Reported_GenderMale"
plot_grid(  resConcord.inter_sex.disease$fig_logFC + ggtitle(main),
            resConcord.inter_sex.disease$fig_tstat + ggtitle(main), nrow=1)

data.frame(Cohort = names(tab.de.inter_sex.disease.lst[[coef]]),
          "# DE"  = resConcord.inter_sex.disease$n.de,
          pi1     = resConcord.inter_sex.disease$pi_single,
          check.names=FALSE) %>% 
          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)

resConcord.inter_sex.disease$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
resConcord.inter_sex.disease$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
```

```{r plot.meta.analysis.SCZ.inter_sex.disease, fig.height=5, fig.width=10}
# volcano plot for meta-analysis.  
main = paste('Meta-analysis', coef, sep = ': ' )
plotVolcano( df_meta.inter_sex.disease[[coef]], df_meta.inter_sex.disease[[coef]]$Symbol[1:20] ) + ggtitle( main )  
```

Here we used multiple methods (BH, qvalue and FDRTool), to compute the local and global FDR and evalute the number of genes passing the 25% cutoff.
```{r pv.hist.fdrtool}
pv = df_meta.inter_sex.disease[[coef]]$P

res_fdr = fdrtool( pv, "pvalue", plot=FALSE, verbose=FALSE)

cutoff = .25
df = data.frame(`BH` = sum(p.adjust(pv, "BH") < cutoff),
          `qvalue (global)` = sum(qvalue(pv)$qvalue < cutoff),
          `qvalue (local)` = sum(qvalue(pv)$lfdr < cutoff),
          `FDRtool (global)` = sum(res_fdr$qval < cutoff),
          `FDRtool (local)` = sum(res_fdr$lfdr < cutoff),
          check.names=FALSE )  %>% t 
colnames(df) = "Count DE genes"

df %>% kable(caption="Number of DE genes by each multiple testing method at cutoff of 25%") %>% kable_styling(full_width=FALSE)
```

The pi1 value is `r 1 - qvalue(pv)$pi0`.

Histogram of p-values from interaction testing is close to unifoirm

```{r p.hist, fig.height=5, fig.width=5}
hist( pv, xlab="p-value", main="Histogram of p-values")
```

## Plot top examples
```{r plot.forrest.dlpfc.inter_sex.disease, fig.width=8, fig.height=40}

figList = lapply(df_meta.inter_sex.disease[[coef]]$Symbol[1:20], function(geneSymbol){
  fig1 = plot_forrest( c(tab.de.inter_sex.disease.lst[[coef]], list("Meta-analysis" = df_meta.inter_sex.disease[[coef]])), geneSymbol, plot2='p')

  fig2 = plot_gene(residuals.lst, df_meta.inter_sex.disease[[coef]], geneSymbol, "Dx") + xlab("Diagnosis") + facet_wrap(~Reported_Gender + Cohort, nrow=1)

  plot_grid(fig1, fig2, nrow=1)
})

plot_grid(plotlist=figList, ncol=1)
```

## Meta-analysis of zenith results
```{r show.gene_set_test.marginal.inter_sex.disease, eval=TRUE}
coef = 'DxSCZ:Reported_GenderMale'   
show_zenith_results( res_meta_zenith.inter_sex.disease, coef, "GO", FDR=0.1)  
show_zenith_results( res_meta_zenith.inter_sex.disease, coef, "MSigDB", FDR=0.1)

show_zenith_results( res_meta_zenith.inter_sex.disease_autosome, coef, "GO", FDR=0.1) 
show_zenith_results( res_meta_zenith.inter_sex.disease_autosome, coef, "MSigDB", FDR=0.1)    
```

```{r plot.gene_set_test.marginal.inter_sex.disease, eval=TRUE}
coef = 'DxSCZ:Reported_GenderMale'   
  plot_zenith_results( res_meta_zenith.inter_sex.disease, coef, "GO", FDR=0.1)  
  plot_zenith_results( res_meta_zenith.inter_sex.disease, coef, "MSigDB", FDR=0.1)

  plot_zenith_results( res_meta_zenith.inter_sex.disease_autosome, coef, "GO", FDR=0.1) 
  plot_zenith_results( res_meta_zenith.inter_sex.disease_autosome, coef, "MSigDB", FDR=0.1)    
```

# Dx test separately in Males and then Females
```{r dream.sex.disease.separately}
form.lst = list(
  `MSSM-Penn-Pitt`  = ~ Dx + RIN +scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + Institution*(ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3),
  `NIMH-HBCC`       = ~ Dx + RIN + scale(IntronicRate) + scale(IntragenicRate) + scale(IntergenicRate) + scale(rRNARate) + ageOfDeath + cellFrac_ilr_1 + cellFrac_ilr_2 + cellFrac_ilr_3)

# DE for males and females separately
fit.separ = lapply( vobj.lst, function(vobj){

  fit.lst = lapply(unique(METADATA$Reported_Gender), function(Sex){

    i = match(colnames(vobj), rownames(METADATA))
    info = METADATA[i,]
    idx = which(info$Reported_Gender == Sex)

    form_use = form.lst[[info$Cohort[1]]]

    fit = dream(vobj[,idx], form_use, info[idx,], BPPARAM = SnowParam(18))  
    eBayes( fit )
  })
  names(fit.lst) = unique(METADATA$Reported_Gender)
  fit.lst
  })
names(fit.separ) = names(vobj.lst)

# meta-analysis
df_meta.separ = lapply(unique(METADATA$Reported_Gender), function(Sex){
  tab.lst = lapply( names(vobj.lst), function(chrt){
    topTable(fit.separ[[chrt]][[Sex]], coef="DxSCZ", sort.by='none', number=Inf)
  })

  res = run_meta_analysis( tab.lst, "FE") %>% getGeneSymbol
  res$t = with(res, logFC/se)
  res
})
names(df_meta.separ) = unique(METADATA$Reported_Gender)
```

```{r separate.test.sz}

df = merge(df_meta.separ[[1]], df_meta.separ[[2]], by="row.names")

df$tstat = with(df, (logFC.x - logFC.y) / sqrt(se.x^2 + se.y^2))

df$p = 2*pnorm(abs(df$tstat), 0, 1, lower.tail=FALSE)
df$FDR = p.adjust(df$p, "fdr")

1 - qvalue(df$p)$pi0

head(df[df$FDR < 0.1,])

coef(lm(t.x ~ t.y, df))
coef(lm(t.y ~ t.x, df))

with(df, hist(t.y - t.x))

with(df, wilcox.test(t.y, t.x, paired=TRUE))

with(df, hist(logFC.y - logFC.x))

```

```{r separ.concordance.scz, fig.height=6, fig.width=10}
# Concordance analysis
resConcord.separ = plot_concordance( df_meta.separ, showGenes = df_meta[['DxSCZ']]$Symbol[1:20])

main = "Concordance of SCZ across sexes"
plot_grid(  resConcord.separ$fig_logFC + ggtitle(main),
            resConcord.separ$fig_tstat + ggtitle(main), nrow=1)

data.frame(Cohort = names(df_meta.separ),
          "# DE"  = resConcord.separ$n.de,
          pi1     = resConcord.separ$pi_single,
          check.names=FALSE) %>% 
          mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 within dataset") %>% kable_styling(full_width=FALSE)

resConcord.separ$tab_pi %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="pi1 for replication") %>% kable_styling(full_width=FALSE)
resConcord.separ$tab_corr %>% mutate_if(is.numeric, format, digits=3) %>% kable(caption="Correlation between t-statistics") %>% kable_styling(full_width=FALSE)
```

## Plot top examples
```{r plot.forrest.dlpfc.separ, fig.width=8, fig.height=40}

figList = lapply(df_meta.separ[[1]]$Symbol[1:20], function(geneSymbol){
  fig1 = plot_forrest( df_meta.separ, geneSymbol, plot2='p')

  plot_grid(fig1, nrow=1)
})

plot_grid(plotlist=figList, ncol=1)
```

```{r store, eval=TRUE}
if( params$upload ){    
  # Code
  # Make sure these match the folder name and parent Id in the first chunk.
  CODE <- Folder(name = "Joint analysis of DLPFC count data", parentId = "syn22416298")
  CODE <- synStore(CODE)

  # Store data 
  storeData = function( name, description){
    file = paste0(name, ".RDS")
    saveRDS( eval(parse(text=name)), file=file)

    nEXP_OBJ = File(file, 
                    name = description, 
                    parentId = CODE$properties$id)
    synStore(nEXP_OBJ, used = ALL_USED_IDs, executed = thisFile)
  }

  # Standard analysis
  storeData( "geneCountsMerged", "Merged DLPFC counts for both cohorts")
  storeData( "METADATA", "Merged metadata used here for both cohorts")
  storeData( "isexpr", "Genes used in this analysis")
  storeData( "vobj.lst", "Voom results for each cohort")
  storeData( "resid.add", "Residuals correcting for covariates and add back - (Dx_Sex Dx Sex)")
  storeData( "fit_marginal.lst", "dream fit for marginal analysis")
  storeData( "res_zenith", "zenith results from fit_marginal.lst")
  storeData( "res_meta_zenith", "Meta-analysis of zenith results from fit_marginal.lst")
  storeData( "res_meta_zenith.sex_autosome", "Meta-analysis of zenith results from fit_marginal.lst - just autosomal genes")
  storeData( "df_DE_results", "Differential expressed genes within each cohort")
  storeData( "df_meta", "Meta-analysis of differential expression across cohorts")

  # Sex by Disease interaction
  storeData( "fit_inter_sex.disease.lst", "dream fit for interaction analysis")
  storeData( "res_zenith.inter_sex.disease", "zenith results for interaction analysis")
  storeData( "res_meta_zenith.inter_sex.disease", "Meta-analysis of zenith results for interaction analysis")
  storeData( "res_meta_zenith.inter_sex.disease_autosome", "Meta-analysis of zenith results for interaction analysis -  autosomal genes only")
  storeData( "df_DE_results.inter_sex.disease", "Differential expression result for each cohort for interaction analysis")
  storeData( "df_meta.inter_sex.disease", "Meta-analysis for differential expression for interaction analysis")
}

#   rmarkdown::render("combined_analysis/combine_all_DLPFC.Rmd")

```

### Source
[markdown](`r thisFile`)