### 2021-05-19 macaque/macaque_DE.Rmd
initial commit

---
title: "Rhesus macaque drug study"
author: "[Gabriel Hoffman](http://gabrielhoffman.github.io)"
date: "Run on `r Sys.time()`"
output: html_document
---

<!---

cd /Users/gabrielhoffman/workspace/repos/CMC/covarr-de/macaque
# rm -rf combined_analysis/combine_all_EPHA1_cache

R

rmarkdown::render("macaque_DE.Rmd")

--->

```{r setup, include=FALSE, echo=FALSE}
library(data.table)
library(synapser)
library(limma)
library(edgeR)
library(corrplot)
library(magrittr)
library(knitr)
library(kableExtra)
library(RColorBrewer)
library(biomaRt)

synLogin()

knitr::opts_chunk$set(
  echo = TRUE,
  warning=FALSE,
  message=TRUE,
  error = FALSE,
  tidy = FALSE,
  dev = c('png', 'pdf'),
  cache = TRUE)
```

## Get data
```{r get.data}
# Rhesus data: https://www.synapse.org/#!Synapse:syn11617745

# reads RNA-seq data
geneCounts = data.frame(fread( synGet('syn11615855')$path))
rownames(geneCounts) = geneCounts$GeneFeature
geneCounts = geneCounts[,-1]

# read sample metadata
info = data.frame(fread( synGet('syn11638462')$path))
rownames(info) = info$DLPFC_RNA_Sequencing_Sample_ID

# sort info based on geneCounts
idx = match(colnames(geneCounts), rownames(info))
info = info[idx,]

# Create drug and dosage variable
info$treatmentType = factor(info$treatmentType, c('placebo', 'clozapine', 'haloperidol'))
info$DrugExposure = with(info, paste0(treatmentType, ' (', dose..mg.kg.d., ' mg/kg)'))
info$DrugExposure = gsub(" \\(NA mg/kg\\)", '', info$DrugExposure)
info$DrugExposure = factor(info$DrugExposure, unique(info$DrugExposure))

# get mapping between Rh mac and human genes
#################

# genome annotation
rhmac = useMart("ensembl", dataset = "mmulatta_gene_ensembl")

geneInfo <- getBM(attributes=c("ensembl_gene_id", "hsapiens_homolog_ensembl_gene","hsapiens_homolog_associated_gene_name"), mart=rhmac, values = rownames(geneCounts))
```

## Differential expession analysis
```{r DE.analysis, fig.width=5, fig.height=5}
# Differential analysis
dsgn = model.matrix( ~ DrugExposure, info)

# voom
keep = filterByExpr(geneCounts, dsgn)
vobj = voom(geneCounts[keep,], dsgn, plot=TRUE)

# lmFit
fit = lmFit(vobj, dsgn)
fit = eBayes(fit)

# get and combine signatures
df_signatures = lapply( colnames(coef(fit))[-1], function(drug){
  tab = topTable(fit, coef=drug, sort.by="none", number=Inf)[,'logFC', drop=FALSE]
  colnames(tab) = gsub("^Drug", '', drug)
  tab
})
df_signatures = do.call(cbind, df_signatures)
df_signatures = merge(df_signatures, geneInfo, by.x="row.names", by.y="ensembl_gene_id")
```

## Compare with CommonMind
```{r cmc}
df_cmc = readRDS( synGet('syn22416508')$path)$DxSCZ

trim_ensembl_ids = function(x){
  gsub("(.*)\\.(.*)", "\\1", x) 
}
df_cmc$ENSEMBL = trim_ensembl_ids( rownames(df_cmc) )

df_signatures = merge(df_signatures, 
                      data.frame(CMCv4 = with(df_cmc, logFC), ENSEMBL=df_cmc$ENSEMBL),
                       by.x="hsapiens_homolog_ensembl_gene", 
                       by.y="ENSEMBL")
```

```{r compare, fig.width=5, fig.height=5}
cols = c(gsub("^Drug", '',colnames(coef(fit))[-1]), 'CMCv4')

C = cor(df_signatures[,cols], method="spearman")

corrplot(C, col = colorRampPalette(c("blue", "red"))(100), main = "Spearman correlation")
```

```{r table}
data.frame(`Fraction of variance` = C['CMCv4',-4]^2) %>%
  kbl() %>%
  kable_styling(full_width=FALSE)
```

```{r test, eval=FALSE}
lapply(names(C['CMCv4',-4]), function(drug){
  cor.test(df_signatures[,drug], df_signatures[,'CMCv4'], method="spearman")
})
```