### 2021-07-28 deconvolution/fans4/molec_profiling.Rmd
freeze 2

---
title: "RNA-seq from 4 cell populations"
subtitle: ''
author: "Developed by [Gabriel Hoffman](http://gabrielhoffman.github.io/)"
date: "Run on `r Sys.time()`"
documentclass: article
output: 
  html_document:
  toc: true
  smart: false
---

<!---

cd /hpc/users/hoffmg01/work/cmc_analysis/src
# rm -rf molec_profiling_cache/
ml git pandoc
git pull
R

system("git pull")
rmarkdown::render("molec_profiling.Rmd", output_dir='./', intermediates_dir='./')

# http
# https://hoffmg01.u.hpc.mssm.edu/molec_profiling.html

cd /Users/gabrielhoffman/workspace/repos/CMC/covarr-de/deconvolution/fans4
R

rmarkdown::render("molec_profiling.Rmd", output_dir='./', intermediates_dir='./')

--->

--->

```{r knit2synapse, eval=FALSE}
library(synapser)
library(knit2synapse)
synLogin()
knit2synapse::createAndKnitToFolderEntity(file = "deconvolution/fans4/molec_profiling.Rmd",
                                          parentId ="syn22231629",
                                          folderName = 'FANS4 processing and testing',
                                          overwrite=TRUE, 
                                          knitmd=TRUE)
```

```{r load.packages, echo=FALSE, message=FALSE, results='hide', cache=FALSE}
suppressPackageStartupMessages({
library(variancePartition)
library(limma)
library(edgeR)
library(mvIC)
library(reshape2)
library(qvalue)
library(data.table)
library(BiocParallel)
library(knitr)
library(org.Hs.eg.db)
library(kableExtra)
library(ggplot2)
library(cowplot)
library(scran)
library(SummarizedExperiment)
library(gridExtra)
library(openxlsx)
library(synapser)
library(githubr)
synLogin()
library(EnsDb.Hsapiens.v86)
detach("package:EnsDb.Hsapiens.v86", unload=TRUE)
library(EnsDb.Hsapiens.v86)
ensdb = EnsDb.Hsapiens.v86
})

source("../../common_functions/common_functions.R")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE,
  cache.lazy = FALSE,  dev = c("png", "pdf"), 
  fig.width=7, fig.height=7)

options(markdown.HTML.stylesheet = 'css/custom.css')
```

```{r synapse.parameters, include=FALSE, cache=TRUE, eval=TRUE}
thisFileName <- 'molec_profiling.Rmd'
# thisFile = thisFileName
# Github link
thisRepo <- getRepo(repository = "CommonMindConsortium/covarr-de", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('deconvolution/fans4/',thisFileName))
```

### Reading RNA-seq data
```{r read.data, message=TRUE}
Ids = c('syn22321012', 'syn22321013', 'syn22321014', 'syn22321016')

# Read in RNA-seq featureCounts
###############################

# Phase 1
file = synGet('syn22321012')$path
df_featureCounts_gene = fread( file )

# Phase 2+3
# syn26000845 is really phase2 again
ids = c('syn26000845','syn25976117', 'syn25976127')[-1]
df_read = lapply(ids, function(id){
  df = fread(synGet(id)$path)
  colnames(df) = gsub("^Sample_", '', colnames(df))
  colnames(df) = gsub("-", '.', colnames(df))
  df
  })

# save batch info
df_batch = data.frame(Name = colnames(df_featureCounts_gene)[-c(1:6)], Batch=1)
df_batch = rbind(df_batch, data.frame(Name = colnames(df_read[[1]][,-c(1:6)]), Batch=2))
df_batch = rbind(df_batch, data.frame(Name = colnames(df_read[[2]][,-c(1:6)]), Batch=3))
df_batch$Batch = as.character(df_batch$Batch)

# sapply(df_read, ncol)
df_featureCounts_gene = cbind(df_featureCounts_gene, 
                          df_read[[1]][,-c(1:6)], 
                          df_read[[2]][,-c(1:6)])

geneInfo = df_featureCounts_gene[,1:6]
geneInfo$EnsId = trim_ensembl_ids( geneInfo$Geneid )
geneInfo = geneInfo %>% getGeneSymbol( 'Geneid' )

geneCounts = df_featureCounts_gene[,7:ncol(df_featureCounts_gene)]
geneCounts = as.matrix(geneCounts)
rownames(geneCounts) = geneInfo$Geneid

# identify samples with improper names
idx = grep("Undetermined", colnames(geneCounts))
message("Dropping samples with improper names:\n",  paste(colnames(geneCounts)[idx], collapse="\n"))
geneCounts = geneCounts[,-idx]

idx = which(colSums(geneCounts) < 1e5)
message("Dropping samples with zero read counts:\n",  paste(colnames(geneCounts)[idx], collapse="\n"))
geneCounts = geneCounts[,-idx]

# remove genes with duplicated EnsId
exclude = duplicated(geneInfo$EnsId)
geneInfo = geneInfo[!exclude,]
geneCounts = geneCounts[!exclude,]
rownames(geneCounts) = trim_ensembl_ids( rownames(geneCounts) )

# convert to HUGO names
# df_hugo = select(org.Hs.eg.db, geneInfo$EnsId, "SYMBOL", "ENSEMBL")
# geneInfo = merge(geneInfo, df_hugo, by.x='EnsId', by.y="ENSEMBL")

```

```{r metadata}
# file = "/sc/arion/projects/CommonMind/roussp01a/MOLECULAR_PROFILING/meta-files/MolecularProfiling_MSSM_120samples_05.24.17-4.xlsx"
file = synGet('syn22321013')$path
# focused on this experiment
metadata1 = read.xlsx( file )

# from Synapse
# file = "/sc/arion/projects/CommonMind/roussp01a/MOLECULAR_PROFILING/meta-files/CMC_METADATA.csv"
file = synGet('syn22321014')$path
metadataCMC = fread( file )
```

### RNA-seq QC metrics
```{r QC, fig.width=18, fig.height=74}
# Phase 1 
# file = '/sc/arion/projects/CommonMind/roussp01a/MOLECULAR_PROFILING/rnaseq/step1-merged/QCmetrics/qc_metrics.RDS'
file = synGet('syn22321016')$path
qc_metrics.in = t(readRDS( file ))
colnames(qc_metrics.in) = qc_metrics.in[1,]
qc_metrics.in = as.data.frame(qc_metrics.in[-c(1:2),], stringsAsFactors=FALSE)
qc_metrics = apply(qc_metrics.in, 2, as.numeric)
rownames(qc_metrics) = gsub('-', '.', rownames(qc_metrics.in))

# Phase 2+3
ids = c('syn26000850', 'syn25976122', 'syn25976136')[-1]
df_read = lapply(ids, function(id){
  qc.in = read.table(synGet(id)$path)
  rownames(qc.in) = gsub('-', '.', rownames(qc.in))
  qc.in
  })

# # sapply(df_read, ncol)
# df_featureCounts_gene = cbind(df_featureCounts_gene, 
#                           df_read[[1]][,-c(1:6)], 
#                           df_read[[2]][,-c(1:6)])

# extract information from sample names
info = data.frame( Name = colnames(geneCounts), stringsAsFactors=FALSE)
rownames(info) = info$Name
info$DissectionID = sapply(strsplit( info$Name, '\\.'), function(x) x[1])
info$CellType = sapply(strsplit( info$Name, '\\.'), function(x) x[2])
info$Assay = sapply(strsplit( info$Name, '\\.'), function(x) x[3])

# set color codes for each cell type
colorCodes = c(
  GABA = '#66A61E',
  GLU = '#E6AB02',
  Olig = '#E7298A',
  MgAs = '#7570B3')

# set factor levels to be the same as the order of the colors
# this makes plotting easier
info$CellType = factor(info$CellType,names(colorCodes))

# add information about each sample
info = merge(info, metadata1, by.x="DissectionID", by.y='Dissection.ID', all.x=TRUE)
info = merge(info, qc_metrics, by.x="Name", by.y='row.names', all.x=TRUE)
rownames(info) = info$Name

# info = merge(info, metadataCMC, by.x='Individual.ID', by.y="Individual ID")
#  info[1:4, 1:20]

# Merge with batch info
info = merge(info, df_batch, by="Name")
rownames(info) = info$Name

# set order to be the same as geneCounts
# Must be last change to info
info = info[match(colnames(geneCounts), info$Name),]

# exclude columns with no variance
index = apply(qc_metrics, 2, function(x) var(x, na.rm=TRUE)) > 0
include = which(index==TRUE)
include = include[! names(include) %in% "PF_BASES"]

df = reshape2::melt(info[,c('DissectionID', 'CellType', colnames(qc_metrics)[include])], id.vars=c('DissectionID', 'CellType'))

ggplot(df, aes( x=CellType, y=value, fill=CellType)) + geom_boxplot() + facet_wrap(~variable, scales="free_y", ncol=5) + theme_bw(16) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + scale_fill_manual("Cell Type", values=colorCodes) 
```

### Study Summary
```{r}
kable(table(info$Dx, info$CellType)) %>% kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

```{r}
kable(table(info$Individual.ID)) %>% kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

```{r preprocess}
isexpr = rowSums(edgeR::cpm(geneCounts)>1) >= 0.1*ncol(geneCounts)
table(isexpr)

dge = DGEList( geneCounts[isexpr,] )
dge = calcNormFactors(dge)
```

```{r TMM}
df = merge(dge$samples, info, by="row.names")

ggplot(df, aes(CellType, norm.factors, fill=CellType)) + geom_boxplot() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=colorCodes) + ylab("TMM factor")
```

```{r expression.density}

geneExpr = edgeR::cpm(dge, log=TRUE)

df = data.table(reshape2::melt(geneExpr))
setkey(df, "Var2")
df = merge(df, info[,c("Name", 'CellType')], by.x = "Var2", by.y="Name")

ggplot(df, aes(value, color=CellType)) + geom_density() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=colorCodes) + xlab(bquote(log[2]~CPM))

ggplot(df, aes(value, color=Var2)) + geom_density() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + facet_wrap(~CellType, ncol=4) + xlab(bquote(log[2]~CPM))
```

### PCA
```{r PCA, fig.width=10, fig.height=5}
dcmp = prcomp( t(geneExpr), scale=TRUE)

df_pca = data.frame(dcmp$x)

df_pca = merge(df_pca, info, by.x="row.names", by.y="Name")

varFrac = dcmp$sdev^2 / sum(dcmp$sdev^2)

xlab = paste0('PC1 (', format(varFrac[1]*100, digits=3), '%)')
ylab = paste0('PC2 (', format(varFrac[2]*100, digits=3), '%)')
fig1 = ggplot(df_pca, aes(PC1, PC2, color=CellType, shape=Batch)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=colorCodes) + xlab(xlab) + ylab(ylab)

xlab = paste0('PC2 (', format(varFrac[2]*100, digits=3), '%)')
ylab = paste0('PC3 (', format(varFrac[3]*100, digits=3), '%)')
fig2 = ggplot(df_pca, aes(PC2, PC3, color=CellType, shape=Batch)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=colorCodes) + xlab(xlab) + ylab(ylab)

plot_grid(fig1, fig2 + theme(legend.position="none"), axis="tbr", align="v")
```

#### Batch corrected
```{r PCA.batch, fig.width=10, fig.height=5}
fit = dream(geneExpr, ~ Batch, info)

dcmp = prcomp( t(residuals(fit)), scale=TRUE)

df_pca = data.frame(dcmp$x)

df_pca = merge(df_pca, info[,c("Name", "CellType", "Batch")], by.x="row.names", by.y="Name")

varFrac = dcmp$sdev^2 / sum(dcmp$sdev^2)

xlab = paste0('PC1 (', format(varFrac[1]*100, digits=3), '%)')
ylab = paste0('PC2 (', format(varFrac[2]*100, digits=3), '%)')
fig1 = ggplot(df_pca, aes(PC1, PC2, color=CellType, shape=Batch)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=colorCodes) + xlab(xlab) + ylab(ylab)

xlab = paste0('PC2 (', format(varFrac[2]*100, digits=3), '%)')
ylab = paste0('PC3 (', format(varFrac[3]*100, digits=3), '%)')
fig2 = ggplot(df_pca, aes(PC2, PC3, color=CellType, shape=Batch)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=colorCodes) + xlab(xlab) + ylab(ylab)

plot_grid(fig1, fig2 + theme(legend.position="none"), axis="tbr", align="v")
```

```{r predict.celltype}
library(nnet)
library(PRROC)

hush=function(code){
  sink("NUL") # use /dev/null in UNIX
  tmp = code
  sink()
  return(tmp)
}

df_class = lapply(1:10, function(k){

  fit = hush(multinom(df_pca$CellType ~ X[,1:k,drop=FALSE]))

  data.table(k =k, n_mismatch = sum(predict(fit) != df_pca$CellType))
  })
df_class = do.call(rbind, df_class)

fig1 = ggplot(df_class, aes(factor(k), n_mismatch)) + geom_bar(stat="identity") + theme_classic() + scale_y_continuous(expand=c(0,0)) + ggtitle("Ability to predict each cell type") + theme(aspect.ratio=1)

df_class = lapply(1:10, function(k){

  fit = hush(multinom(df_pca$CellType ~ X[,1:k,drop=FALSE]))

  Y_pred = predict(fit, type="probs")

  df_class = lapply( levels(df_pca$CellType), function(CT){

    y0 = Y_pred[df_pca$CellType == CT,CT]
    y1 = Y_pred[df_pca$CellType != CT,CT]

    pr = pr.curve(scores.class0 = y0, scores.class1 = y1, curve=TRUE, rand.compute=TRUE)

    data.table(k = k, CellType = CT, AUPR = pr$auc.integral, AUPR.rand = pr$rand$auc.integral)
    })
  do.call(rbind, df_class)
  })
df_class = do.call(rbind, df_class)

fig2 = ggplot(df_class, aes(k, AUPR, color=CellType)) + geom_point() + geom_line() + theme_classic() + scale_color_manual(values = colorCodes) + scale_x_continuous(breaks = 1:10) + ggtitle("Ability to predict each cell type") + theme(aspect.ratio=1)

plot_grid(fig1, fig2, axis="tbr", align="v")
```

### Sex check
```{r sex.check}

XIST = 'ENSG00000229807'
UTY = 'ENSG00000183878'

df = data.frame(  Name  = colnames(geneExpr), 
                  XIST  = geneExpr[XIST,],
                  UTY   = geneExpr[UTY,])
df = merge( df, info[,c('Name', 'Gender', 'Individual.ID')], by="Name" )
df$Gender = factor(df$Gender, c("Male", 'Female'))

ggplot(df, aes(XIST, UTY, color=Gender)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c("blue", "red")) + ggtitle("Sex check")
```

Samples with wrong sex
```{r sex.off}
kable(df[df$Gender == 'Female' & df$UTY > 5,], digits=2, row.names=FALSE) %>% kable_styling(bootstrap_options = "striped", full_width = FALSE)  
```

### Esimate precision weights
```{r voom}
form = ~ (1|CellType) + (1|Individual.ID) + (1|Gender)
vobj = voomWithDreamWeights(dge, form, info, plot=TRUE, save.plot=TRUE, BPPARAM=SnowParam(6, progressbar=TRUE))
```

### mvIC
Need to get proper covariates here

```{r mvIC, eval=FALSE}
colnames(info) = gsub(' ', '_', colnames(info))
colnames(info) = gsub(':', '', colnames(info))
colnames(info) = gsub('\\(', '', colnames(info))
colnames(info) = gsub(')', '', colnames(info))
colnames(info) = gsub('/', '', colnames(info))
colnames(info) = gsub('-', '_', colnames(info))
colnames(info) = gsub(',', '', colnames(info))

include = apply(info, 2, function(x) all(!is.na(x)) && var(x, na.rm=TRUE) >0 )

variables = colnames(info)[colnames(info) %in% names(include[include])]

mvSearch = mvForwardStepwise( vobj, ~ CellType, info, variables)
```

### variancePartition
```{r vp}
form = ~ (1|CellType) + (1|Dx) + (1|CellType:Dx) + (1|Individual.ID) + (1|Gender) + (1|Batch)
vp = fitExtractVarPartModel( geneExpr, form, info, BPPARAM=SnowParam(6, progressbar=TRUE))

plotVarPart( sortCols(vp) )
```

## Differential expression analysis
### Cell type
```{r DE_process}
form = ~ 0 + CellType + (1|Individual.ID) + (1|Gender) + (1|Batch)

# Prepare all pairwise contrasts
index = combn(nlevels(info$CellType),2)

L = lapply( 1:ncol(index), function(j){

  CT1 = levels(info$CellType)[index[1,j]]
  CT2 = levels(info$CellType)[index[2,j]]

  L = getContrast( vobj, form, info, paste0('CellType', c(CT1, CT2)))

  L = as.matrix(L)
  colnames(L) = paste0(CT1, '.', CT2)

  L
})
L = do.call(cbind, L)

L = cbind(L, data.frame(GABAGLU.Other = c(1, 1, -1, -1),
                    MgAs.Other = c(-1/3, -1/3, -1/3, 1),
                    Olig.Other = c(-1/3, -1/3, 1, -1/3) ))

fitBasic = dream( vobj, form, info, L=L, BPPARAM=SnowParam(6, progressbar=TRUE))
```

```{r plotContrasts}
plotContrasts(round(L,2)) + theme(axis.text.x=element_text(angle=45, hjust=1))   
```

```{r DE.analysis}

index = combn(nlevels(info$CellType),2)

df_deCount = lapply( 1:ncol(index), function(j){

    CT1 = levels(info$CellType)[index[1,j]]
    CT2 = levels(info$CellType)[index[2,j]]

    key = paste0(CT1, '.', CT2)

    tabl = topTable(fitBasic, coef=key, number=Inf)

    data.frame(CT1, CT2, key, count = sum(tabl$adj.P.Val < 0.05), pi1 = 1 - qvalue(tabl$P.Value)$pi0)
})
df_deCount = do.call(rbind, df_deCount)

ggplot(df_deCount, aes(CT1, CT2, fill=count)) + geom_tile() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_fill_gradient(low="white", high="red", limits=c(0, 30000)) + geom_text(aes(label=format(count, big.mark=','))) + xlab('') + ylab('') + ggtitle("Count differentiall expression genes")

ggplot(df_deCount, aes(CT1, CT2, fill=pi1)) + geom_tile() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_fill_gradient(name = quote(pi[1]), low="white", high="red", limits=c(0, 1)) + geom_text(aes(label=format(pi1, digits=3))) + xlab('') + ylab('') + ggtitle(bquote(pi[1]))
```

```{r cache=FALSE}
knitr::knit_exit()
```

```{r DE2}

index = combn(nlevels(info$CellType),2)

df = lapply( 1:ncol(index), function(j){

    CT1 = levels(info$CellType)[index[1,j]]
    CT2 = levels(info$CellType)[index[2,j]]

    key = paste0(CT1, '.', CT2)

    tabl = topTable(fitBasic, coef=key, number=Inf)

    data.frame(CT1, CT2, key, tabl)
})
df = do.call(rbind, df)

ggplot(df, aes(logFC, ..count..)) + geom_density(alpha = .2) + facet_wrap(~key) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) 
```

## Plots of top differentially expressed genes
```{r DE.marker_genes, fig.width=15, fig.height=30}

index = combn(nlevels(info$CellType),2)

ensIDs = sapply( 1:ncol(index), function(j){

    CT1 = levels(info$CellType)[index[1,j]]
    CT2 = levels(info$CellType)[index[2,j]]

    key = paste0(CT1, '.', CT2)

    tabl = topTable(fitBasic, coef=key, number=Inf)

    tabl = tabl[order(tabl$t),]

    c(head(rownames(tabl), 10), tail(rownames(tabl), 10))
})

ensIDs = unique(c(ensIDs, 'ENSG00000110693', 'ENSG00000167281'))

df = reshape2::melt(geneExpr[ensIDs,])
df = merge(df, info[,1:3], by.x="Var2", by.y="row.names")

df_hugo = select(org.Hs.eg.db, unique(as.character(df$Var1)), "SYMBOL", "ENSEMBL")
df = merge(df, df_hugo, by.x="Var1", by.y="ENSEMBL")
df = df[!is.na(df$SYMBOL),]

ggplot(df, aes(CellType, value, fill=CellType)) + geom_boxplot() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position = "none") + scale_fill_manual("Cell Type", values=colorCodes) + facet_wrap(~SYMBOL, ncol=5) + ylab(bquote(log[2]~CPM))
```

## Analysis using just pairs of samples
```{r pairwise.analysis}

index = combn(nlevels(info$CellType),2)

# for all pairs of cell types
tabList = lapply( 1:ncol(index), function(j){

  CT1 = levels(info$CellType)[index[1,j]]
  CT2 = levels(info$CellType)[index[2,j]]
  key = paste0(CT1, '.', CT2)

  idx = which(info$CellType %in% c(CT1, CT2))

  isexpr = rowSums(edgeR::cpm(geneCounts[,idx])>1) >= 0.3*ncol(geneCounts[,idx])
  table(isexpr)
  infoSub = droplevels(info[idx,])

  dge = DGEList( geneCounts[isexpr,idx] )
  dge = calcNormFactors(dge)

  form = ~ CellType + Gender
  vobj = voomWithDreamWeights(dge, form, infoSub, plot=FALSE, save.plot=TRUE, BPPARAM=SnowParam(6, progressbar=TRUE))

  form = ~ CellType + Gender
  fit = dream( vobj, form, infoSub, BPPARAM=SnowParam(6, progressbar=TRUE))
  fit = eBayes(fit)

  coef = paste0('CellType',levels(infoSub$CellType)[2])

  topTable(fit, coef=coef, number=Inf)
})

names(tabList) = apply(matrix(levels(info$CellType)[index], nrow=2), 2, function(x) paste0(x, collapse='.'))

df_de_summary = lapply( names(tabList), function(key){
  tab = data.table(tabList[[key]])

  data.frame( Comparison  = key, 
              nGenes      = nrow(tab), 
              nDE         = sum(tab$adj.P.Val < 0.05),
              nDE_gt_1    =  nrow(tab[adj.P.Val < 0.05 & abs(logFC) > 1,]), 
              nDE_gt_2    =  nrow(tab[adj.P.Val < 0.05 & abs(logFC) > 2,]), 
              nDE_gt_3    =  nrow(tab[adj.P.Val < 0.05 & abs(logFC) > 3,]), 
              nDE_gt_4    =  nrow(tab[adj.P.Val < 0.05 & abs(logFC) > 4,]), 
              pi1         = 1-qvalue(tab$P.Value)$pi0)
  })
df_de_summary = do.call(rbind, df_de_summary)

kable(df_de_summary) %>% kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

```{r pairwise.density}

index = combn(nlevels(info$CellType),2)

# for all pairs of cell types
tabList = lapply( 1:ncol(index), function(j){

  CT1 = levels(info$CellType)[index[1,j]]
  CT2 = levels(info$CellType)[index[2,j]]
  key = paste0(CT1, '.', CT2)

  idx = which(info$CellType %in% c(CT1, CT2))

  isexpr = rowSums(edgeR::cpm(geneCounts[,idx])>1) >= 0.3*ncol(geneCounts[,idx])
  table(isexpr)
  infoSub = droplevels(info[idx,])

  dge = DGEList( geneCounts[isexpr,idx] )
  dge = calcNormFactors(dge)

  form = ~ CellType + Gender
  vobj = voomWithDreamWeights(dge, form, infoSub, plot=FALSE, save.plot=TRUE, BPPARAM=SnowParam(6, progressbar=TRUE))

  form = ~ CellType + Gender
  fit = dream( vobj, form, infoSub, BPPARAM=SnowParam(6, progressbar=TRUE))
  fit = eBayes(fit)

  coef = paste0('CellType',levels(infoSub$CellType)[2])

  data.frame(key, topTable(fit, coef=coef, number=Inf))
})
df = do.call(rbind, tabList)

ggplot(df, aes(logFC)) + geom_density(alpha = .2) + facet_wrap(~key) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) 
```

```{r compare.voom}

index = combn(nlevels(info$CellType),2)

# for all pairs of cell types
df_voomTrend = lapply( 1:ncol(index), function(j){

  CT1 = levels(info$CellType)[index[1,j]]
  CT2 = levels(info$CellType)[index[2,j]]
  key = paste0(CT1, '.', CT2)

  idx = which(info$CellType %in% c(CT1, CT2))

  isexpr = rowSums(edgeR::cpm(geneCounts[,idx])>1) >= 0.3*ncol(geneCounts[,idx])
  table(isexpr)
  infoSub = droplevels(info[idx,])

  dge = DGEList( geneCounts[isexpr,idx] )
  dge = calcNormFactors(dge)

  form = ~ CellType + Gender
  vobj = voomWithDreamWeights(dge, form, infoSub, plot=FALSE, save.plot=TRUE, BPPARAM=SnowParam(6, progressbar=TRUE))

  data.frame( key, x = vobj$voom.line[['x']], y = vobj$voom.line[['y']])
})
df_voomTrend = do.call(rbind, df_voomTrend)

ggplot(df_voomTrend, aes(x, y, color=key)) + geom_line() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + ylab("sqrt( standard deviation)") + xlab("log2( count size + 0.5 )") + xlim(0, 22) + ylim(0.4, 1.4)
```

## Identify marker genes
```{r tabList.markerGenes}

tabList = lapply(colnames(L), function(coef){
  tab = topTable(fitBasic, coef=coef, number=Inf)
  # tab[which((tab$adj.P.Val < 0.05) & (abs(tab$logFC) > 5)),]
  tab
})
names(tabList) = colnames(L)

saveRDS(tabList, file='./Roussos_MSSM_MolProf_FANS_4_diffExpression.RDS')

# system("ml python; synapse add --parentid syn22243354 Roussos_MSSM_MolProf_FANS_4_diffExpression.RDS")
```

## Plots of top differentially expressed genes
```{r markerGenes, fig.width=15, fig.height=30}

lapply( names(tabList), function(key){tab = tabList[[key]]; nrow(tab[tab$logFC > 6,]) })

boxplot_gene = function(ensIDs, tab, ngenes=5){
  df = reshape2::melt(geneExpr[ensIDs,,drop=FALSE])
  df = merge(df, info[,1:3], by.x="Var2", by.y="row.names")

  df_hugo = select(org.Hs.eg.db, unique(as.character(df$Var1)), "SYMBOL", "ENSEMBL")
  df = merge(df, df_hugo, by.x="Var1", by.y="ENSEMBL")
  df = df[!is.na(df$SYMBOL),]

  df = df[df$SYMBOL %in% df_hugo$SYMBOL[!is.na(df_hugo$SYMBOL)][1:ngenes], ]

  ggplot(df, aes(CellType, value, fill=CellType)) + geom_boxplot() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position = "none") + scale_fill_manual("Cell Type", values=colorCodes) + facet_wrap(~SYMBOL, nrow=1) + ylab(bquote(log[2]~CPM))
}

figList = lapply( names(tabList), function(key){

  tab = tabList[[key]]

  # only select up regulated genes
  tab = tab[tab$logFC > 6,]

  ensIDs = rownames(tab)

  boxplot_gene( ensIDs, tab, 5) + ggtitle(key)
  } )
plot_grid(plotlist = figList, ncol=1)

```

```{r FANS.markers}
ensIDs =  c('ENSG00000110693','ENSG00000167281', 'ENSG00000100146')
boxplot_gene(ensIDs, tab) + ggtitle("FANS markers")
```

# Compare to GTEx
```{r GTEX.voom}

# file_SampleAttributesDS = "/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Analysis_v8_Annotations_SampleAttributesDS.txt"
file_SampleAttributesDS = synGet('syn22321103')$path

# file_SubjectPhenotypesDS= "/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Analysis_v8_Annotations_SubjectPhenotypesDS.txt"
file_SubjectPhenotypesDS = synGet('syn22321104')$path

# file_counts = "/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Analysis_2017-06-05_v8_RNASeQCv1.1.9_gene_reads.gct.gz"
file_counts = synGet('syn22321107')$path

gtex_sample = read.delim(file_SampleAttributesDS, header = TRUE, stringsAsFactors = FALSE)
# gtex_sample = gtex_sample[gtex_sample$SMTS == 'Brain',]

rnaseq_sample_id = fread(file_counts, header = TRUE, skip = 2, nrows = 1)
gtex_sample = gtex_sample[gtex_sample$SAMPID %in% colnames(rnaseq_sample_id),]

# read count
gtex_count <- fread(file_counts, header = TRUE, skip = 2, select = c('Name', unique(gtex_sample$SAMPID)))
gtex_count = as.data.frame(gtex_count, stringsAsFactors=FALSE)

gtex_count$Name = gsub("^(.*)(\\.\\S+)$", "\\1", gtex_count$Name )
gtex_count = gtex_count[!duplicated(gtex_count$Name), ]
rownames(gtex_count) = gtex_count$Name
gtex_count$Name = NULL
gtex_count = as.matrix(gtex_count)

# isexpr = rowSums(cpm(gtex_count>1)) >= 0.3*ncol(gtex_count)
# dge = DGEList( gtex_count[isexpr,] )
# dge = calcNormFactors(dge)
# gtex_log2cpm = cpm(dge, log=TRUE)

# df_voomTrend_GTEx = lapply( unique(gtex_sample$SMTSD), function(Tissue){

#   idx = which( gtex_sample$SMTSD == Tissue)

#   isexpr = rowSums(cpm(gtex_count[,idx])>1) >= 0.3*ncol(gtex_count[,idx])
#   table(isexpr)

#   dge = DGEList( gtex_count[isexpr,idx] )
#   dge = calcNormFactors(dge)

#   vobj = voomWithDreamWeights(dge, ~1, gtex_sample[idx,], plot=FALSE, save.plot=TRUE, BPPARAM=SnowParam(6, progressbar=TRUE))

#   data.frame( key = Tissue, x = vobj$voom.line[['x']], y = vobj$voom.line[['y']])
# })
# df_voomTrend_GTEx = do.call(rbind, df_voomTrend_GTEx)
```

```{r GTEX.voom.plot}
# ggplot(df_voomTrend_GTEx, aes(x, y, color=key)) + geom_line() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="bottom") + ylab("sqrt( standard deviation)") + xlab("log2( count size + 0.5 )") + ggtitle("GTEx brain regions") + xlim(2, 22) + ylim(0.45, 1.36)
```

```{r gtex.colors}
# https://genome.ucsc.edu/cgi-bin/hgTables?db=hg19&hgta_track=hgFixed&hgta_group=allTables&hgta_table=hgFixed.gtexTissue

# file= '/sc/arion/projects/CommonMind/data/resources/GTEx/GTEx_Tissue_colors.csv'
file = synGet('syn22321106')$path
gtexColors = read.csv( file, header=TRUE)
```

```{r CV.all.in.one, eval=TRUE}
library(scran)
# FANS4
isexpr = rowSums(edgeR::cpm(geneCounts)>1) >= 0.3*ncol(geneCounts)

se <- SummarizedExperiment( assays=list(counts = as.matrix(geneCounts[isexpr,])), colData = info)
res = modelGeneCV2( se, block=info$CellType )
fit = metadata(res)

df_fans4 = lapply( names(res$per.block), function(key){
  with( res$per.block[[key]], data.frame(key, mean, total, trend, ratio, FDR))
  })
df_fans4 = data.table(do.call(rbind, df_fans4))

df_fans4 = df_fans4[order(df_fans4$mean),]
ggplot(df_fans4, aes(log2(mean), log2(trend), color=key)) + geom_line() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + xlab(bquote(log[2]~Mean~expression)) + ylab(bquote(log[2]~CV~trend)) + scale_color_manual(values=colorCodes)+ xlim(0, 20) + ylim(-5, 5)

# GTEX
idx = which(gtex_sample$SMTSD %in% names(which(table(gtex_sample$SMTSD) > 30)))

isexpr = rowSums(edgeR::cpm(gtex_count[,idx])>1) >= 0.3*ncol(gtex_count[,idx])

se <- SummarizedExperiment( assays=list(counts = gtex_count[isexpr,idx]), colData = gtex_sample[idx,])
res = modelGeneCV2( se, block=gtex_sample$SMTSD[idx] )

df_gtex = lapply( names(res$per.block), function(key){
  with( res$per.block[[key]], data.frame(key, mean, total, trend, ratio, FDR))
  })
df_gtex = data.table(do.call(rbind, df_gtex))
df_gtex$key = factor(df_gtex$key)

i = match(levels(df_gtex$key), gtexColors$description)
colsGTEx = as.hexmode(gtexColors$color)[i[!is.na(i)]]

df_gtex = df_gtex[order(df_gtex$mean),]
df_gtex = df_gtex[df_gtex$mean > 0,]

ggplot(df_gtex, aes(log2(mean), log2(trend), color=key)) + geom_line() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + xlab(bquote(log[2]~Mean~expression)) + ylab(bquote(log[2]~CV~trend))+ scale_color_manual(values=colsGTEx) + xlim(0, 20) + ylim(-5, 5)

df_gtex$dataset = 'GTEx'
df_fans4$dataset = "FANS4"
df_combined = rbind(df_gtex, df_fans4)

ggplot(df_combined, aes(log2(mean), log2(trend), color=key, size=dataset)) + geom_line() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + xlab(bquote(log[2]~Mean~expression)) + ylab(bquote(log[2]~CV~trend)) + scale_size_manual(values=c(2, .1)) + scale_color_manual(values=c(rep("black", length(colsGTEx)), colorCodes)) + xlim(-8, 20) + ylim(-5, 3) 
```

```{r coef.variation, eval=TRUE}

CTarray = as.list(levels(info$CellType))
# CTarray[[5]] = c("GABA", "GLU")
# CTarray[[6]] = levels(info$CellType)

# different because only genes expressed in given cell type are used
res_cv_FANS = lapply(CTarray, function(CT){

  id = info$Name[which(info$CellType %in% CT)]

  isexpr = rowSums(edgeR::cpm(geneCounts[,id])>1) >= 0.1*ncol(geneCounts[,id])

  se <- SummarizedExperiment( assays=list(counts = as.matrix(geneCounts[isexpr,id])), colData = info[id,])
  res = modelGeneCV2( se ) # returns CV^2
  fit = metadata(res)

  geneTab = select(ensdb, keys=names(isexpr[isexpr]), columns=c("GENEID", "GENEBIOTYPE"), keytype="GENEID")

  list(df = data.frame( key = paste0(CT, collapse="."), res ),
       f = fit$trend,
       nExpressed = sum(isexpr),
       nExpressed.CDS = sum(geneTab$GENEBIOTYPE == 'protein_coding'))
})

x = 2^seq(-3, 15, length.out=500)
df_trend_FANS = lapply(res_cv_FANS, function(res){ 
  data.frame( Dataset="FANS4", key = res$df$key[1], mean = x, trend=res$f(x))
  })
df_trend_FANS = do.call(rbind, df_trend_FANS)

df_cv_FANS = lapply(res_cv_FANS, function(res){ 
  res$df
  })
df_cv_FANS = do.call(rbind, df_cv_FANS)

colorCodes['GABA.GLU'] = "orange"
colorCodes['GABA.GLU.Olig.MgAs'] = "dodgerblue"

# Take sqrt(CV^2)
ggplot(df_trend_FANS, aes(log2(mean), log2(sqrt(trend)), color=key, size=Dataset)) + geom_line() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position=c(.7, .9)) + xlab(bquote(log[2]~Mean~expression)) + ylab(bquote(log[2]~CV~trend)) + scale_size_manual(values=c(2, .1)) + scale_color_manual(values=colorCodes) + scale_x_continuous(expand=c(0.01, 0))

ggplot(df_cv_FANS[df_cv_FANS$key %in% names(colorCodes)[1:4],], aes(log2(sqrt(total)), color=key)) + geom_density(size=2) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="right")  + scale_color_manual(values=colorCodes) + xlab(bquote(log[2]~CV))

ggplot(df_cv_FANS, aes(log2(sqrt(total)), color=key)) + geom_density(size=2) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="right")  + scale_color_manual(values=colorCodes) + xlab(bquote(log[2]~CV))

#  GTEX
res_cv_GTEx = lapply( names(which(table(gtex_sample$SMTSD) > 30)), function(Tissue){

  id = gtex_sample$SAMPID[which( gtex_sample$SMTSD == Tissue)]

  isexpr = rowSums(edgeR::cpm(gtex_count[,id])>1) >= 0.1*ncol(gtex_count[,id])

  se <- SummarizedExperiment( assays=list(counts = gtex_count[isexpr,id]), colData = gtex_sample[id,])
  res = modelGeneCV2( se )
  fit = metadata(res)

  geneTab = select(ensdb, keys=names(isexpr[isexpr]), columns=c("GENEID", "GENEBIOTYPE"), keytype="GENEID")

  list(df = data.frame( key = Tissue, res ),
       f = fit$trend,
       nExpressed = sum(isexpr),
       nExpressed.CDS = sum(geneTab$GENEBIOTYPE == 'protein_coding'))
})

df_trend_GTEx = lapply(res_cv_GTEx, function(res){ 
  data.frame( Dataset="GTEx", key = res$df$key[1], mean = x, trend=res$f(x))
  })
df_trend_GTEx = do.call(rbind, df_trend_GTEx)
df_trend_GTEx$key = factor(df_trend_GTEx$key)

df_cv_GTEx = lapply(res_cv_GTEx, function(res){ 
  res$df
  })
df_cv_GTEx = do.call(rbind, df_cv_GTEx)
df_cv_GTEx$key = factor(df_cv_GTEx$key)

i = match(levels(df_trend_GTEx$key), gtexColors$description)
colsGTEx = as.hexmode(gtexColors$color)[i[!is.na(i)]]

ggplot(df_trend_GTEx, aes(log2(mean), log2(sqrt(trend)), color=key)) + geom_line() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + xlab(bquote(log[2]~Mean~expression)) + ylab(bquote(log[2]~CV~trend)) + scale_color_manual(values=colsGTEx) + scale_x_continuous(expand=c(.01,0))

ggplot(df_cv_GTEx, aes(log2(sqrt(total)), color=key)) + geom_density(size=2) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + xlab(bquote(log[2]~CV~trend)) + scale_color_manual(values=colsGTEx)

# combine
df_combined = rbind(df_trend_GTEx, df_trend_FANS)

ggplot(df_combined, aes(log2(mean), log2(trend), color=key, size=Dataset)) + geom_line() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + xlab(bquote(log[2]~Mean~expression)) + ylab(bquote(log[2]~CV~trend)) + scale_size_manual(values=c(2, .2)) + scale_color_manual(values=c(rep("grey60", length(colsGTEx)), colorCodes)) + scale_x_continuous(expand=c(.01,0))

df_cv_GTEx$Dataset = "GTEx"
df_cv_GTEx$Type = "Real"
df_cv_FANS$Dataset = "FANS4"
df_cv_FANS$Type = "Real"
df_cv_FANS$Type[df_cv_FANS$key %in% names(colorCodes)[5:6]] = "Synthetic"

df = data.table(rbind(df_cv_GTEx, df_cv_FANS))

ggplot(df, aes(log2(total), color=key, size=Dataset, linetype=Type)) + geom_density(size=2) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + xlab(bquote(log[2]~CV~trend)) + scale_color_manual(values=c(rep("grey", length(colsGTEx)), colorCodes)) + scale_size_manual(values=c(2, .2)) + scale_linetype_discrete(c("solid", "dashed"))

# number of expressed genes
 df[,sum(total > 0), by='key']

df2 = lapply(sort(unique(df$key)), function(KEY){
  data.frame(key = KEY, Dataset = df$Dataset[df$key == KEY][1],
    Type = df$Type[df$key == KEY][1],
    CV2=df[key==KEY,][order(mean,decreasing=TRUE)[1:15000],]$total)
  })
df2 = data.table(do.call(rbind, df2))
df2$key = factor(df2$key, (sort(unique(df2$key))))
levels(df2$key)

i = match(levels(df2$key), gtexColors$description)
colsGTEx = paste0("#", as.hexmode(gtexColors$color)[i[!is.na(i)]])
names(colsGTEx) = gtexColors$description[i[!is.na(i)]]

cols = c(colsGTEx, colorCodes)[levels(df2$key)]
cols[names(colsGTEx)] = "grey"

ggplot(df2, aes(log2(sqrt(CV2)), color=key, size=Dataset, linetype=Type)) + geom_density(size=2) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + xlab(bquote(log[2]~CV~trend)) + scale_color_manual(values=cols) + scale_size_manual(values=c(2, .1)) + scale_linetype_discrete(c("solid", "dashed"))

df3 = df2[,median(sqrt(CV2)), by='key']
df3$key = factor(df3$key, rev(levels(df3$key)))

ggplot( df3[!is.na(df3$V1)], aes(key, V1, fill=key)) + geom_bar(stat='identity') + theme_bw(10) + theme(aspect.ratio=5, plot.title = element_text(hjust = 0.5), legend.position="none")  + scale_fill_manual(values=c(colsGTEx, colorCodes)[levels(df3$key)]) + coord_flip() + ylab(bquote(Median~CV)) + scale_y_continuous(expand=c(0,0)) + xlab('') + ylim(0, max(df3$V1, na.rm=TRUE)*1.05)

# Compare to IPS and LCL

# genes expressed

df_count_fans4 = lapply(res_cv_FANS, function(res){  
  data.frame( Dataset="FANS4", key = res$df$key[1], 
    "Expressed genes"=res$nExpressed, "Expressed coding genes"=res$nExpressed.CDS)
 })
df_count_fans4 = do.call(rbind, df_count_fans4)

df_count_gtex = lapply(res_cv_GTEx, function(res){  
  data.frame( Dataset="GTEx", key = res$df$key[1], 
    "Expressed genes"=res$nExpressed, "Expressed coding genes"=res$nExpressed.CDS)
 })
df_count_gtex = do.call(rbind, df_count_gtex)

df_count = rbind(df_count_gtex, df_count_fans4)
df_count$key = factor(df_count$key, rev(levels(df2$key)))

i = match(levels(df_count$key), gtexColors$description)
colsGTEx = paste0("#", as.hexmode(gtexColors$color)[i[!is.na(i)]])
names(colsGTEx) = gtexColors$description[i[!is.na(i)]]

df_melt = reshape2::melt(df_count, id.vars=c("Dataset", "key"))

ggplot( df_melt, aes(key, value, fill=key)) + geom_bar(stat='identity') + theme_bw(10) + theme(aspect.ratio=5, plot.title = element_text(hjust = 0.5), legend.position="none")  + scale_fill_manual(values=c(colsGTEx, colorCodes)[levels(df3$key)]) + coord_flip() + ylab("# genes expressed") + scale_y_continuous(expand=c(0,0)) + xlab('') + facet_wrap(~variable)

```

```{r plot.coef.variation, fig.width=12, fig.height=12, eval=FALSE}
# use colors here
# gtexColors and colorCodes

df = rbind(df_cv_GTEx, df_cv_FANS)
df$key = factor(df$key, c(rev(sort(unique(df_cv_FANS$key))), rev(sort(unique(df_cv_GTEx$key)))))

i = match(levels(df$key), gtexColors$description)
cols = as.hexmode(gtexColors$color)[i[!is.na(i)]]
cols = paste0('#', cols)
names(cols) = gtexColors$description[i[!is.na(i)]]

cols = c(cols, colorCodes)[levels(df$key)]

cols[!names(cols) %in% unique(df_cv_FANS$key)] = "black"

# ggplot( df, aes(log2(BCV), color=key)) + geom_density() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="bottom") + facet_wrap(~Dataset, ncol=1) + scale_color_manual("Type", values=cols)

ggplot(df[,nrow(.SD), by="key"], aes(key, V1, fill=key)) + geom_bar(stat="identity") + theme_bw(10) + theme(aspect.ratio=5, plot.title = element_text(hjust = 0.5), legend.position="none") + scale_fill_manual("Cell Type", values=cols) + coord_flip() + xlab('') + ylab("# genes exprressed")

ggplot( df[CV.ratio>=1,], aes(CV.ratio, ..density.., color=key)) + geom_density(size=2) + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + scale_color_manual("Cell Type", values=cols) + xscale("log2", .format = TRUE) + facet_wrap(~Dataset, ncol=1)

ggplot( df[CV.ratio>=1,], aes(log2(CV.ratio), color=key)) + stat_ecdf() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position="none") + scale_color_manual("Cell Type", values=cols) + facet_wrap(~Dataset, ncol=2) + ylab("eCDF") + scale_x_continuous(expand=c(0,0), limits=c(0, 10))

df2 = df[,data.frame(Dataset=Dataset[1], Count = sum(CV.ratio > 2) / length(CV.ratio)),by="key"]

v = df2[Dataset=="GTEx", median(Count)]

ggplot(df2, aes(key, Count, fill=key))  + geom_bar(stat="identity") + theme_bw(10) + theme(aspect.ratio=5, plot.title = element_text(hjust = 0.5), legend.position="none") + scale_fill_manual("Cell Type", values=cols) + coord_flip() + xlab('') + ylab('# genes passing cutoff') + geom_hline(yintercept=v, linetype="dashed", color="grey70")

df2 = df[,data.frame(Dataset=Dataset[1], Count = sum(FDR < .05)/length(FDR)),by="key"]
v = df2[Dataset=="GTEx", median(Count)]

ggplot(df[,sum(FDR < .05),by='key'], aes(key, V1, fill=key))  + geom_bar(stat="identity") + theme_bw(8) + theme(aspect.ratio=5, plot.title = element_text(hjust = 0.5), legend.position="none") + scale_fill_manual("Cell Type", values=cols) + coord_flip() + xlab('') + ylab('# genes passing cutoff') + geom_hline(yintercept=v, linetype="dashed", color="grey70")

```

```{r cv.old}

# cv = function(x){
#   sd(x) / mean(x)
# }

# #   cv_est = apply(geneExpr[,id], 1, cv)

# #   data.frame( key = CT, Gene = names(cv_est), CV = cv_est, stringsAsFactors=FALSE)
# # })
# # df_cv_FANS = do.call(rbind, df_cv_FANS)

# # CV for GTEX
# df_cv_GTEx = lapply( unique(gtex_sample$SMTSD), function(Tissue){

#   id = gtex_sample$SAMPID[which( gtex_sample$SMTSD == Tissue)]

#   cv_est = apply(gtex_log2cpm[,id], 1, cv)

#   data.frame( key = Tissue, Gene = names(cv_est), CV = cv_est, stringsAsFactors=FALSE)
# })
# df_cv_GTEx = do.call(rbind, df_cv_GTEx)

# df = rbind(df_cv_FANS, df_cv_GTEx)

# ggplot( df_cv_FANS, aes(log10(CV), color=key)) + geom_density() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=colorCodes) + xlim(-2, 5) + ylim(0,1.4)

# ggplot( df_cv_GTEx, aes(log10(CV), color=key)) + geom_density() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + xlim(-2, 5) + ylim(0,1.4)

# ggplot( df, aes(log10(CV), color=key)) + geom_density() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + xlim(-2, 5) + ylim(0,1.4)

```

### PCA including GTEx
Computed on residuals

```{r PCA.gtex}

id = gtex_sample$SAMPID[which( gtex_sample$SMTSD == 'Brain - Frontal Cortex (BA9)')]
countsCombined = merge(geneCounts, gtex_count[,id], by="row.names")
rownames(countsCombined) = countsCombined$Row.names
countsCombined$Row.names = NULL

isexpr = rowSums(edgeR::cpm(countsCombined)>1) >= 0.3*ncol(countsCombined)

dge = DGEList(countsCombined[isexpr,])
dge = calcNormFactors(dge)

geneExprCombined = edgeR::cpm(dge, log=TRUE)

infoExpand = rbind(data.frame(info[,c(1,3)], dataset="FANS4", stringsAsFactors=FALSE), 
  data.frame(Name=colnames(gtex_count[,id]), CellType = "GTEx BA9", dataset="GTEx", stringsAsFactors=FALSE))
rownames(infoExpand) = infoExpand$Name

# the first PC separates FANS from bulk, so subtract that and use residuals
fit = dream(geneExprCombined, ~ dataset, infoExpand)
resid = residuals(fit, geneExprCombined)

dcmp = prcomp( t(resid), scale=TRUE)

df_pca = data.frame(dcmp$x)

df_pca = merge(df_pca, infoExpand, by.x="row.names", by.y="Name")

varFrac = dcmp$sdev^2 / sum(dcmp$sdev^2)

xlab = paste0('PC1 (', format(varFrac[1]*100, digits=3), '%)')
ylab = paste0('PC2 (', format(varFrac[2]*100, digits=3), '%)')
ggplot(df_pca, aes(PC1, PC2, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, "grey")) + xlab(xlab) + ylab(ylab)

# xlab = paste0('PC2 (', format(varFrac[2]*100, digits=3), '%)')
# ylab = paste0('PC3 (', format(varFrac[3]*100, digits=3), '%)')
# ggplot(df_pca, aes(PC2, PC3, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, "grey")) + xlab(xlab) + ylab(ylab)

```

### MDS
computed on log cpm

```{r GTEX.MDS}

d = as.dist(1 - cor(geneExprCombined)^2)
# d = as.dist(1 - cor(apply(geneExprCombined, 2, RNOmni::rankNorm), method="sp")^2)

df_mds = data.frame(cmdscale( d, k=3))

df_mds = merge(df_mds, infoExpand, by.x="row.names", by.y="Name")

ggplot(df_mds, aes(X1, X2, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, "grey")) + xlab('Coordinate 1') + ylab('Coordinate 2')

ggplot(df_mds, aes(X2, X3, color=CellType)) + geom_point() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + scale_color_manual("Cell Type", values=c(colorCodes, "grey")) + xlab('Coordinate 2') + ylab('Coordinate 3')
```

## Differential expression with Increasing sample size
```{r subset.analysis}

n_array =  c(3, 5, 10, 20, 28)
sampleRamp = lapply(n_array, function(N){

  df = sapply(levels(info$CellType), function(CT){
    info$Name[which(info$CellType == CT)[1:N]]
    })
  data.frame(df, stringsAsFactors=FALSE)
  })

index = combn(nlevels(info$CellType),2)

tabRampList = lapply( sampleRamp, function(sampleSet){

  infoSub = droplevels(info[info$Name %in% unlist(sampleSet),])
  geneCountSub = geneCounts[,infoSub$Name]

  # for all pairs of cell types
  tabList = lapply( 1:ncol(index), function(j){

    CT1 = levels(infoSub$CellType)[index[1,j]]
    CT2 = levels(infoSub$CellType)[index[2,j]]
    key = paste0(CT1, '.', CT2)

    idx = which(infoSub$CellType %in% c(CT1, CT2))

    message(key, " ", nrow(sampleSet) )
    infoSub2 = droplevels(infoSub[idx,])

    isexpr = rowSums(edgeR::cpm(geneCountSub[,idx])>1) >= 0.3*ncol(geneCountSub[,idx])
    table(isexpr)

    dge = DGEList( geneCountSub[isexpr,idx] )
    dge = calcNormFactors(dge)

    form = ~ CellType
    vobj = voomWithDreamWeights(dge, form, infoSub2, plot=FALSE, save.plot=TRUE, BPPARAM=SnowParam(6, progressbar=TRUE), quiet=TRUE)

    form = ~ CellType
    fit = dream( vobj, form, infoSub2, BPPARAM=SnowParam(6, progressbar=TRUE), quiet=TRUE)
    fit = eBayes(fit)

    coef = paste0('CellType',levels(infoSub2$CellType)[2])

    topTable(fit, coef=coef, number=Inf)
  })
  names(tabList) = apply(matrix(levels(infoSub$CellType)[index], nrow=2), 2, function(x) paste0(x, collapse='.'))
  tabList  
})
```

Just counting coding genes, but analysis was for everything

```{r subset.analysis.plots, fig.width=9}

df_de_summary = lapply( 1:length(tabRampList), function(j){

  tabList = tabRampList[[j]]

  df_de_summary = lapply( names(tabList), function(key){
    tab = tabList[[key]]

    geneTab = select(ensdb, keys=rownames(tab), columns=c("GENEID", "GENEBIOTYPE"), keytype="GENEID")

    tab = merge(tab, geneTab, by.x="row.names", by.y="GENEID")

    tab = data.table(tab)
    tab = tab[GENEBIOTYPE=='protein_coding',]

    df_de_summary = data.frame( Comparison  = key, 
                N           = n_array[j],
                nGenes      = nrow(tab), 
                nDE         = sum(tab$adj.P.Val < 0.05),
                nDE_gt_1    =  nrow(tab[adj.P.Val < 0.05 & abs(logFC) > 1,]), 
                nDE_gt_2    =  nrow(tab[adj.P.Val < 0.05 & abs(logFC) > 2,]), 
                nDE_gt_3    =  nrow(tab[adj.P.Val < 0.05 & abs(logFC) > 3,]), 
                nDE_gt_4    =  nrow(tab[adj.P.Val < 0.05 & abs(logFC) > 4,]), 
                pi1         = 1-qvalue(tab$P.Value)$pi0)
    })
   do.call(rbind, df_de_summary)
})
df_de_summary = do.call(rbind, df_de_summary)

ggplot(df_de_summary, aes(N, nDE_gt_1, color=Comparison)) + geom_point() + geom_line() + theme_bw(15) + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5), legend.position = "right") + ylab("# DE and |logFC| > 1") + xlab("Sample size") + ylim(0,NA)
```

# Disease
```{r DE.disease}
# Disease
fitDisease = dream( vobj, ~ 0 + CellType*Dx + (1|Individual.ID) + (1|Gender), info, BPPARAM=SnowParam(6, progressbar=TRUE))
```

```{r Disease.analysis}
df_disease = lapply(levels(info$CellType), function(CT){

  key = paste0('CellType', CT, ':DxSCZ')

  if( CT == levels(info$CellType)[1]){
    key = 'DxSCZ'
  }

  tabl = topTable( fitDisease, coef = key, number=Inf)

  data.frame(CellType = CT, key, count = sum(tabl$adj.P.Val < 0.05), pi1 = 1 - qvalue(tabl$P.Value)$pi0)
  })
df_disease = do.call(rbind, df_disease)

kable(df_disease, digits=2, row.names=FALSE) %>% kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

Volcano plot colored by cell type

DE from null model
Power vs sample size
Paired versus non-paired
Plot examples of marker genes
distribution of effect size
GTEX brain span.
Compare cell types
Homogeneity of variance for each cell type

For next week: deconvolution of GTEx, PEC, CMC, BrainSpan, BrainVar

bMIND can incorporate the prior of cell type variance/covariance. That is, the K x K covariance matrix for K cell types for each gene. I am wondering if Gabriel can also share this info, if sharing the sample-level data is not possible.

### Save mean expression per cell type
https://www.synapse.org/#!Synapse:syn22231629
```{r save.reference.panel}

# compute mean for each cell type, using precision weights
fitCT = dream( vobj, ~ 0 + CellType, info, BPPARAM=SnowParam(6, progressbar=TRUE))
fitCT = eBayes(fitCT)

df_RefPanel = lapply( colnames(fitCT), function(coef){

  # Extract mean for this cell type
  tabl = topTable(fitCT, coef=coef, number=Inf, sort.by='none')
  tabl$se = tabl$logFC / tabl$t # compute standard error given beta and t

  df = data.frame(tabl[,c('logFC', 'se')])
  rownames(df) = rownames(tabl)
  colnames(df) = paste0(gsub('CellType', '', coef), c('', '.se'))
  df

  })
df_RefPanel = do.call(cbind, df_RefPanel)
df_RefPanel = data.frame(EnsID = rownames(df_RefPanel), df_RefPanel)

exclude = grep("\\.se$", colnames(df_RefPanel))
```

```{r store, eval=TRUE}
# Code
CODE <- Folder(name = "FANS4 processing and testing", parentId = "syn22231629") # Make sure these match the folder name and parent Id in the first chunk.
CODE <- synStore(CODE)

# write mean expression
file = 'Roussos_MSSM_MolProf_FANS_4_log2CPM_reference_panel.csv.gz'
gz1 <- gzfile(file, "w")
write.table(df_RefPanel[,-exclude], gz1, quote=FALSE, sep=",", row.names=FALSE)
close(gz1)

nEXP_OBJ = File(file, 
                name = 'FANS4 reference panel - log2CPM', 
                parentId = CODE$properties$id)
nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
                    executed = thisFile)

# write standard error
file = 'Roussos_MSSM_MolProf_FANS_4_log2CPM_se_reference_panel.csv.gz'
gz1 <- gzfile(file, "w")
write.table(df_RefPanel[,c(1,exclude)], gz1, quote=FALSE, sep=",", row.names=FALSE)
close(gz1)

nEXP_OBJ = File(file, 
                name = 'FANS4 reference panel - log2CPM standard error', 
                parentId = CODE$properties$id)
nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
                    executed = thisFile)

# full counts
file = 'Roussos_MSSM_MolProf_FANS_4_geneCounts.csv.gz'
gz1 <- gzfile(file, "w")
write.table(data.table(EnsID = rownames(geneCounts), geneCounts), gz1, quote=FALSE, sep=",", row.names=FALSE)
close(gz1)

nEXP_OBJ = File(file, 
                name = 'FANS4 reference panel - full counts', 
                parentId = CODE$properties$id)
nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
                    executed = thisFile)

# full counts
file = 'Roussos_MSSM_MolProf_FANS_4_info.csv.gz'
gz1 <- gzfile(file, "w")
write.table(info, gz1, quote=FALSE, sep=",", row.names=FALSE)
close(gz1)

nEXP_OBJ = File(file, 
                name = 'FANS4 reference panel - metadata', 
                parentId = CODE$properties$id)
nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
                    executed = thisFile)

# Differential expression
file = "Roussos_MSSM_MolProf_FANS_4_diffExpression.RDS"

nEXP_OBJ = File(file, 
                name = 'FANS4 Differential Expression results', 
                parentId = CODE$properties$id)
nEXP_OBJ = synStore(nEXP_OBJ, used = Ids, 
                    executed = thisFile)

```

### Source
[markdown](`r thisFile`)